### 2021-01-17 research project/sections/Introduction.tex
update

\section{Introduction}

The value of $\pi$ has been researched for many years, although under different names,
and the amount of different approaches to reach the value is large.

The value has been found through many processes, be it geometrically, algebraically 
or through other means.

This paper seeks to examine the extent at which two historical methods of
approximation of the value $\pi$, namely the approaches suggested by the aforementioned 
mathematicians Madhava and Viète, differ in terms of computational efficiency and 
speed.

### 2021-01-18 chemistry/doc.tex
""

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{mhchem}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
  \usepackage{amssymb}
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Limiting reagents},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Limiting reagents}
\author{}
\date{}

\begin{document}
\maketitle

\hypertarget{limiting-reagents}{%
\section{Limiting reagents}\label{limiting-reagents}}

\hypertarget{lab-about-limiting-reagents}{%
\subsection{Lab about limiting
reagents}\label{lab-about-limiting-reagents}}

\(\ce{ZnO}\):\\
6.O1g before 4.10g after 1. Mass of ZnO used:
\[6.11 g - 4.10 g = 2.01 g\] 2. Molar mass of ZnO:
\[65.38 g/mol + 15.99 g/mol = 81.38 g/mol\] 3. How many moles of zinc
reacted with the HCl? \[2.01 g \div 81.38 g/mol = 0.0246 mol\] 4.
Balance the equation for this experiment
\[\ce{ZnO(aq) + 2 HCl(aq) -> ZnCl2(aq) + H2O(l)}\] 5. Find the number of
moles of \(\ce{HCl}\) in
\[\ce{HCl} = 1.01g/mol + 35.45g/mol = 36.46g/mol\]
\[\ce{HCl} = 20 \div 36.46g/mol = 0.549 mol\] 6. Find the number of
moles of \(\ce{HCl}\) in \[\ce{HCl} = 0.549 mol * 50 = 27.45 mol\]

\end{document}

### 2021-01-19 math/ia_2021/model/model.py
broke myhead doing this

@@ -1,0 +1,67 @@
+import numpy as np
+import pandas as pd
+from matplotlib import pyplot as plt
+
+
+def filter_time(df, threshold):
+    filter_df = df.loc[df.year > threshold].copy()
+    return filter_df
+
+
+def filter_distance(df, threshold):
+    filter_df = df.loc[df['distance from source (km)'] < threshold].copy()
+    return filter_df
+
+
+def filter_doubtful(df):
+    filter_df = df.loc[df['doubtful runup'] != 'y'].copy()
+    return filter_df
+
+
+def calculate_mean(x):
+    x['weight'] * x['maximum water height (m)']
+
+
+def weighted_mean_height(df):
+    return df.groupby('earthquake magnitude', as_index=True).apply(lambda x: (x['weight'] * x['maximum water height (m)']).mean())
+
+
+def get_droppable(df, whitelist):
+    to_drop = []
+    for (columnName, columnData) in df.iteritems():
+        if columnName not in whitelist:
+            to_drop.append(columnName)
+    return to_drop
+
+
+if __name__ == "__main__":
+    df = pd.read_table("dataset/runups.tsv", sep='\t')
+    df.columns = df.columns.str.lower()
+
+    # Filter data points before installation of
+    filter_df = filter_time(df, threshold=1960)
+
+    # No doubtful runups.
+    filter_df = filter_doubtful(filter_df)
+
+    # Drop all entries with no reported wave height
+    filter_df = filter_df.dropna(how='any', subset=['maximum water height (m)'])
+
+    # Filter waves originating from a certain distance away
+    # filter_df = filter_distance(filter_df, 30)
+
+    filter_df = filter_df.drop(get_droppable(filter_df,
+                                             ['year', 'mo', 'earthquake magnitude', 'distance from source (km)',
+                                              'tsunami event validity',
+                                              'maximum water height (m)']), axis=1)
+
+    filter_df['weight'] = 1 - filter_df['distance from source (km)'] / filter_df['distance from source (km)'].max()
+
+    filter_df = weighted_mean_height(filter_df)
+
+    filter_df.to_csv("test.csv")
+
+    print(filter_df)
+
+    plt.scatter(filter_df['earthquake magnitude'], filter_df['0'], marker="x")
+    plt.show()

### 2021-01-19 math/ia_2021/model/test.py

@@ -1,0 +1,7 @@
+import matplotlib.pyplot as plt
+import pandas as pd
+
+df = pd.read_table("test.csv", sep='\t')
+
+plt.scatter(df['aaa'], df['test'], marker="x")
+plt.show()

### 2021-01-20 research project/sections/Background.tex
""

\section{Theoretical approach}

\subsection{Focus on two methods}

For the sake of this paper, two different methods with similar 
convergence rates but different approaches have been chosen for 
comparison, the process for the original discovery of these methods
will be explained.

\subsubsection{Madhava's method}

Madhava, known by other famous mathematicians of his time, as having 
discovered

\subsubsection{Viète's method}

### 2021-01-20 research project/sections/Introduction.tex
""

@@
 The value of $\pi$ has been researched for many years, although under different names,
-and the amount of different approaches to reach the value is large.
-
-The value has been found through many processes, be it geometrically, algebraically 
+and the amount of different approaches to reach the value is large. The value has been 
+found through many processes, be it geometrically, algebraically 
 or through other means.
@@
 mathematicians Madhava and Viète, differ in terms of computational efficiency and 
-speed.
+speed, and explain these differences.

### 2021-01-21 research project/sections/Background.tex
worked on rp

@@
 comparison, the process for the original discovery of these methods
-will be explained.
+will be explained. The two mathematicians in question are Madhava 
+of Sangamagramma of Medieval India and the French mathematician 
+Viète.
 
@@
 
-Madhava, known by other famous mathematicians of his time, as having 
-discovered
+Madhava,
 
 \subsubsection{Viète's method}
+
+The French mathematician, François Viète, approached the value of 
+$\pi$ from a geometric standpoint, finding the following formula:
+
+$$\pi = 2 \frac{2}{\sqrt{2}} \frac{2}{\sqrt{2 + \sqrt{2}}} \frac{2}{\sqrt{2 + \sqrt{2 + \sqrt{2}}}} \dots$$
+
+% $$\pi = \frac{2}{\prod_{k = 0}^{\infty} U_{k}}$$    symbolic
+He was able to calculate $\pi$ to a place of 9 decimal points, in 
+the year 1593 \cite{Kreminski}, using his method. His method consists 
+of finding the area of a polygon of $n$ sides in a circle of constant 
+radius. As the value of $n$ is increased, the area of the $n$-gon 
+tends toward the area of a circle. The geometric origin of this formula 
+can be found using simple right-angle trigonometry, by first finding 
+the lengths $OH$ and subsequently $BD$ in Figure 1 (\ref{fig:image1}).
+
+With the radius $R = OB$,
+
+$OH = R \cos{\alpha}$
+
+and
+
+$BD = 2BH = 2 R \sin{\alpha}$
+
+Since the equation for the area of a polygon is defined as $A = \frac{p \cdot a}{2}$, 
+where $p$ is the perimeter of the polygon and $a$ is the apothem, in this case $BD \cdot n$ 
+and $OH$ respectively.
+
+Let $A_{n}$ equal the area of the polygon with $n$ sides as such: \\
+$A_{n} = \frac{OH \cdot BD \cdot n}{2}$ \\
+$A_{n} = \frac{R \cos{\alpha} \cdot 2 R \sin{\alpha} \cdot n}{2} = n R^2 \sin{\alpha} \cos{\alpha}$
+
+And if n is multiplied by 2, the angle $\angle \alpha$ is divided by 2, 
+and the new area becomes:
+
+$A_{2n} = 2n R^2 \sin{\frac{\alpha}{2}} \cos{\frac{\alpha}{2}}$
+
+So
+
+\begin{figure}[!h]
+\definecolor{qqwuqq}{rgb}{0,0.39215686274509803,0}
+\definecolor{uuuuuu}{rgb}{0.26666666666666666,0.26666666666666666,0.26666666666666666}
+\definecolor{xdxdff}{rgb}{0.49019607843137253,0.49019607843137253,1}
+\definecolor{qqqqzz}{rgb}{0,0,0.6}
+\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=1cm,y=1cm,scale=0.7]
+\clip(-8.372520930232552,-8.50470697674418) rectangle (7.69259534883721,7.028316279069767);
+\draw [shift={(-0.03355999206952884,0)},line width=1pt,color=qqwuqq,fill=qqwuqq,fill opacity=0.10000000149011612] (0,0) -- (0:1.3302325581395344) arc (0:44.77428433325932:1.3302325581395344) -- cycle;
+\draw [line width=0.8pt,color=qqqqzz] (0,0) circle (6cm);
+\draw [line width=1pt] (4.242640687119286,4.242640687119285)-- (6,0);
+\draw [line width=1pt] (6,0)-- (4.242640687119286,-4.242640687119285);
+\draw [line width=1pt] (4.242640687119286,-4.242640687119285)-- (4.242640687119286,4.242640687119285);
+\draw [line width=1pt] (-6,0)-- (6,0);
+\draw [line width=1pt] (-4.275936263984878,-4.209081736713964)-- (4.242640687119286,4.242640687119285);
+\draw [line width=1pt] (-4.242640687119285,4.242640687119286)-- (4.242640687119286,-4.242640687119285);
+\draw [line width=1pt] (-0.03355999206952884,0)-- (5.121320343559643,2.1213203435596424);
+\draw [line width=1pt] (-4.242640687119285,4.242640687119286)-- (-4.275936263984878,-4.209081736713964);
+\begin{scriptsize}
+\draw [fill=xdxdff] (4.242640687119286,4.242640687119285) circle (2.5pt);
+\draw[color=xdxdff] (4.4772465116279085,4.623665116279071) node {$B$};
+\draw [fill=xdxdff] (-4.242640687119285,4.242640687119286) circle (2.5pt);
+\draw[color=xdxdff] (-4.295311627906972,4.623665116279071) node {$A$};
+\draw [fill=xdxdff] (-4.275936263984878,-4.209081736713964) circle (2.5pt);
+\draw[color=xdxdff] (-4.336241860465112,-4.55) node {$C$};
+\draw [fill=xdxdff] (6,0) circle (2.5pt);
+\draw[color=xdxdff] (6.237246511627907,0.3873860465116302) node {$E$};
+\draw [fill=xdxdff] (4.242640687119286,-4.242640687119285) circle (2.5pt);
+\draw[color=xdxdff] (4.3772465116279085,-4.5) node {$D$};
+\draw[color=black] (5.3595720930232575,2.3111069767441874) node {$H'$};
+\draw [fill=xdxdff] (5.12,2.12) circle (2.5pt);
+\draw[color=black] (3.947479069767444,-0.615404651162788) node {$H$};
+\draw [fill=xdxdff] (4.242640687119286,0) circle (2.5pt);
+\draw [fill=xdxdff] (0,0) circle (2.5pt);
+\draw[color=uuuuuu] (0.00003720930232879,-0.3464558139534907) node {$O$};
+\draw[color=qqwuqq] (1.5325953488372117,0.3873860465116302) node {$\alpha$};
+\end{scriptsize}
+\end{tikzpicture}
+\label{fig:image1}
+\caption{Circle with 1 segment from a $n$-gon with point $H$ and 2 
+segments from an $2n$-gon, one of which on point $H'$, inscribed in 
+a circle of radius $OB$}
+\end{figure}

### 2021-01-21 research project/sections/Computational_approach.tex
worked on rp

\section{Computational approach}

\subsection{Implementing}

\subsection{The variables}

### 2021-01-21 math/ia_2021/model/test.py

@@ -1,7 +1,9 @@
 import matplotlib.pyplot as plt
+import numpy as np
 import pandas as pd
 
-df = pd.read_table("test.csv", sep='\t')
+df = pd.read_csv("test.csv")
 
+plt.plot(np.unique(df['aaa']), np.poly1d(np.polyfit(df['aaa'], df['test'], 1))(np.unique(df['aaa'])))
 plt.scatter(df['aaa'], df['test'], marker="x")
 plt.show()

### 2021-01-21 research project/sections/Background.tex
""

@@
 $A_{n} = \frac{OH \cdot BD \cdot n}{2}$ \\
-$A_{n} = \frac{R \cos{\alpha} \cdot 2 R \sin{\alpha} \cdot n}{2} = n R^2 \sin{\alpha} \cos{\alpha}$
+$= \frac{R \cos{\alpha} \cdot 2 R \sin{\alpha} \cdot n}{2} = n R^2 \sin{\alpha} \cos{\alpha}$
 
@@
 
-So
+So it can be written that, by definition,
+
+$\frac{A_{n}}{A_{2n}} = \frac{n R^2 \sin{\alpha} \cos{\alpha}}{2 n R^2 \sin{\frac{\alpha}{2}} \cos{\frac{\alpha}{2}}} = \frac{\sin{2 \alpha}}{2 \sin{\alpha}}$
+
+Which through the trigonometric identity... // todo
 
@@
 segments from an $2n$-gon, one of which on point $H'$, inscribed in 
-a circle of radius $OB$}
+a circle of radius $OB$, adapted from \cite{borisgourevitch2013}}
 \end{figure}

### 2021-01-22 research project/sections/Background.tex
""

@@
 He was able to calculate $\pi$ to a place of 9 decimal points, in 
-the year 1593 \cite{Kreminski}, using his method. His method consists 
-of finding the area of a polygon of $n$ sides in a circle of constant 
-radius. As the value of $n$ is increased, the area of the $n$-gon 
+the year 1593 \cite{Kreminski}, using his method. His method is reminiscent 
+of Archimedes' method, but differs in that it consists of finding 
+the area of a polygon of $n$ sides in a circle of constant 
+radius, rather than the circumference. As the value of $n$ is increased, the area of the $n$-gon 
 tends toward the area of a circle. The geometric origin of this formula 
 can be found using simple right-angle trigonometry, by first finding 
-the lengths $OH$ and subsequently $BD$ in Figure 1 (\ref{fig:image1}).
+the lengths $OH$ and subsequently $BD$ in \ref{fig:image1}.
 
-With the radius $R = OB$,
+With the radius of the circle with center $R = OB$,
 
@@
 where $p$ is the perimeter of the polygon and $a$ is the apothem, in this case $BD \cdot n$ 
-and $OH$ respectively.
-
-Let $A_{n}$ equal the area of the polygon with $n$ sides as such: \\
+and $OH$ respectively, let $A_{n}$ equal the area of 
+the polygon with $n$ sides such that: \\
 $A_{n} = \frac{OH \cdot BD \cdot n}{2}$ \\
@@
 
-So it can be written that, by definition,
+So it can be written that, by definition, the ratio of 
+the area of an $n$-gon
 
@@
 
-Which through the trigonometric identity... // todo
+Which through the trigonometric identity $\sin{2 \alpha} = 2\sin{\alpha} \cos{\alpha}$ can 
+be simplified to:
+
+$\frac{A_{n}}{A_{2n}} = \frac{2\sin{\alpha} \cos{\alpha} }{2\sin{\alpha}}  = \cos{\alpha}$
+
+It can be then written that, through a new variable $P$, \\
+$P = \frac{A_{n}}{A_{2n}} \frac{A_{2n}}{A_{4n}} \frac{A_{4n}}{A_{8n}} \dots \frac{A_{(k-2)n}}{A_{kn}} \frac{A_{kn}}{A}$ \\
+where $A$ is the area of the circle of radius $R$ 
+in \ref{fig:image1}.
+
+So $P = \frac{A_{n}}{A}$, since the values $A_{kn}$ cancel, and 
+followingly, it is true that $P = \frac{A_{n}}{A} \Leftrightarrow A = \frac{A_{n}}{P}$. Since 
+the area of a circle is defined by $A = \pi R^2$, the value of $R$ being constant in this case,
 

### 2021-01-22 math/ia_2021/model/model.py

@@ -62,6 +62,6 @@ if __name__ == "__main__":
     filter_df.to_csv("test.csv")
 
     print(filter_df)
-
-    plt.scatter(filter_df['earthquake magnitude'], filter_df['0'], marker="x")
+    
+    plt.scatter(filter_df['earthquake magnitude'], filter_df['maximum water height (m)'], marker="x")
     plt.show()

### 2021-01-22 math/ia_2021/model/test.py

@@ -2,8 +2,33 @@ import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
 
-df = pd.read_csv("test.csv")
 
-plt.plot(np.unique(df['aaa']), np.poly1d(np.polyfit(df['aaa'], df['test'], 1))(np.unique(df['aaa'])))
+def polyfit(x, y, degree):
+    results = {}
+
+    coeffs = np.polyfit(x, y, degree)
+
+     # Polynomial Coefficients
+    results['polynomial'] = coeffs.tolist()
+
+    # r-squared
+    p = np.poly1d(coeffs)
+    # fit values, and mean
+    yhat = p(x)                         # or [p(z) for z in x]
+    ybar = np.sum(y)/len(y)          # or sum(y)/len(y)
+    ssreg = np.sum((yhat-ybar)**2)   # or sum([ (yihat - ybar)**2 for yihat in yhat])
+    sstot = np.sum((y - ybar)**2)    # or sum([ (yi - ybar)**2 for yi in y])
+    results['determination'] = ssreg / sstot
+
+    return results
+
+
+df = pd.read_csv("test.csv")
+power = 1
+
+
+print(polyfit(df['aaa'], df['test'], power))
+
+plt.plot(np.unique(df['aaa']), np.poly1d(np.polyfit(df['aaa'], df['test'], power))(np.unique(df['aaa'])))
 plt.scatter(df['aaa'], df['test'], marker="x")
 plt.show()

### 2021-01-23 research project/sections/Background.tex
worked on rp

@@
 comparison, the process for the original discovery of these methods
-will be explained. The two mathematicians in question are Madhava 
-of Sangamagramma of Medieval India and the French mathematician 
-Viète.
+will be explained. The two methods in question are one based on 
+the infinite series definition of an inverse trigonometric function 
+and one where the value is derived using geometry, by mathematician 
+Viète. These two mathematical methods were chosen as they are both 
+represent a first occurence in mathematics: Madhava was the first to 
+find the series notation of the $\arctan$ function and Viète was one 
+of the first mathematicians to use an infinite series in calculation.
 
-\subsubsection{Madhava's method}
+\subsubsection{Madhava-Gregory-Leibniz method}
 
-Madhava,
+Madhava, having indentified the integral and series for the $\arctan$ function, 
+in the Kerala school during medieval India \cite{amermathmonthly}
+
+TODO... finish
 
@@
 $A_{n} = \frac{OH \cdot BD \cdot n}{2}$ \\
-$= \frac{R \cos{\alpha} \cdot 2 R \sin{\alpha} \cdot n}{2} = n R^2 \sin{\alpha} \cos{\alpha}$
+$A_{n} = \frac{R \cos{\alpha} \cdot 2 R \sin{\alpha} \cdot n}{2} = n R^2 \sin{\alpha} \cos{\alpha}$
 
@@
 So $P = \frac{A_{n}}{A}$, since the values $A_{kn}$ cancel, and 
-followingly, it is true that $P = \frac{A_{n}}{A} \Leftrightarrow A = \frac{A_{n}}{P}$. Since 
-the area of a circle is defined by $A = \pi R^2$, the value of $R$ being constant in this case,
+followingly, it is true that $P = \frac{A_{n}}{A} \Leftrightarrow A = \frac{A_{n}}{P}$. The 
+value of $R$ being constant in this case, and since the area of a circle is defined 
+by $A = \pi R^2$, therefore: \\
+$\pi = \frac{A_{n}}{P}$ \\
+By definition, we can say that as the value $k$ approaches infinity, the area of the $kn$-gon 
+aproaches that of a circle, and therefore, the value of $\pi$.
 
-\begin{figure}[!h]
+$\frac{A_{n}}{\cos{\alpha} \cos{\frac{\alpha}{2}} \cos{\frac{\alpha}{4}} \dots}
+\to_{k \to \infty} \pi$
+
+There is however, a neater expression for this same formula. 
+To simplify to it, firstly we let $U_{0} = \cos{\alpha}$. Then, we 
+define $U_{1} = cos{\frac{\alpha}{2}}$, following the formula stated 
+before.
+
+TODO... finish
+
+\begin{figure}[!ht]
 \definecolor{qqwuqq}{rgb}{0,0.39215686274509803,0}

### 2021-01-24 research project/sections/Analysis_of_results.tex
worked on rp

\section{Analysis of the results}

\subsection{Presentation of the data}

\subsection{Analysis of the data}

### 2021-01-24 research project/sections/Background.tex
worked on rp

@@
 represent a first occurence in mathematics: Madhava was the first to 
-find the series notation of the $\arctan$ function and Viète was one 
-of the first mathematicians to use an infinite series in calculation.
+find the series notation of the arctangent function and Viète was one 
+of the first mathematicians to use infinite series in his calculation.
 
@@
 So it can be written that, by definition, the ratio of 
-the area of an $n$-gon
+the area of an $n$-gon to one of a $2n$-gon is
 
@@
 
-Which through the trigonometric identity $\sin{2 \alpha} = 2\sin{\alpha} \cos{\alpha}$ can 
+Which through the trigonometric identity $\sin{2 \theta} = 2\sin{\theta} \cos{\theta}$ can 
 be simplified to:
@@
 
-There is however, a neater expression for this same formula. 
-To simplify to it, firstly we let $U_{0} = \cos{\alpha}$. Then, we 
-define $U_{1} = cos{\frac{\alpha}{2}}$, following the formula stated 
-before.
+Where the value of $A_{n}$ is the area of the first polygon, with $n=4$ sides, and as such 
+$A_{n} = 4 \sin{45} \cos{45} = 2$. We can define: \\
+$U_{0} = \cos{a} = \cos{45} = \frac{1}{\sqrt{2}}$ \\
+$U_{1} = cos{\frac{\alpha}{2}}$ \\
+Which we can, through the trigonometric identity 
+$\cos^2{\theta} = \frac{1}{2} + \frac{1}{2} \cos{2\theta}$, simplify as
+$U_{1} = \sqrt{\frac{1}{2} + \frac{1}{2} U_{0}}$ \\
+So it can be said that $U_{n} = \sqrt{\frac{1}{2} + \frac{1}{2} U_{n - 1}}$, which leads to 
+a fully defined expression for the value of pi:
 
-TODO... finish
+$\pi = \frac{2}{\prod\limits_{k=0}^\infty U_{k}}, U_{0} = \frac{1}{\sqrt{2}}, U_{n} = \sqrt{\frac{1}{2} + \frac{1}{2} U_{n - 1}}$ \footnotemark
+
+These expressions, when under a single expression result in the aforementioned formula with nested roots.
+
+\footnotetext{where $\prod$ signifies a product. Similar expression to $\sum$ }
 

### 2021-01-24 research project/sections/Background.tex
worked on rp

@@
-\section{Theoretical approach}
+\section{Background Information}
 
@@
 
-For the sake of this paper, two different methods with similar 
-convergence rates but different approaches have been chosen for 
+For the sake of this paper, one by French mathematician François Viète,
+and another supposedly discovered by Madhava of Sangamagrama, and rediscovered 
+by Swiss mathematician Leibniz. Two methods with different approaches have been chosen for 
 comparison, the process for the original discovery of these methods
@@
 find the series notation of the arctangent function and Viète was one 
-of the first mathematicians to use infinite series in his calculation.
+of the first mathematicians to use infinite series in his calculation. Furthermore, 
+a research based on computational speed of two different kinds approaches to 
+the constant $\pi$ has not been done to date.
 
@@
 
-The French mathematician, François Viète, approached the value of 
-$\pi$ from a geometric standpoint, finding the following formula:
+François Viète, having approached the value of 
+$\pi$ from a geometric standpoint, found the following formula:
 
@@
 the year 1593 \cite{Kreminski}, using his method. His method is reminiscent 
-of Archimedes' method, but differs in that it consists of finding 
-the area of a polygon of $n$ sides in a circle of constant 
+of Archimedes' method, where the length of a side is calculated \cite{archimedespi}, 
+but differs in that it consists of finding the area of a polygon of $n$ sides in a circle of constant 
 radius, rather than the circumference. As the value of $n$ is increased, the area of the $n$-gon 
@@
 followingly, it is true that $P = \frac{A_{n}}{A} \Leftrightarrow A = \frac{A_{n}}{P}$. The 
-value of $R$ being constant in this case, and since the area of a circle is defined 
+value of $R = 1$ in this case, and since the area of a circle is defined 
 by $A = \pi R^2$, therefore: \\
@@
 
-These expressions, when under a single expression result in the aforementioned formula with nested roots.
+These expressions, when under a single expression result in the aforementioned 
+formula with nested roots.
 
@@
 segments from an $2n$-gon, one of which on point $H'$, inscribed in 
-a circle of radius $OB$, adapted from \cite{borisgourevitch2013}}
+a circle of radius $OB$, adapted from Boris Gourévitch \cite{borisgourevitch2013}}
 \end{figure}

### 2021-01-24 research project/sections/Computational_approach.tex
worked on rp

@@
 
-\subsection{Implementing}
+\subsection{The variables}
+The dependent variable of this experiment is the time taken $t$ by the program 
+to approximate a given number $n$ of correct decimal value of the constant $\pi$.
 
-\subsection{The variables}
+The value $n$ will be altered in order to avoid possible similar convergence rates 
+at a small amount of decimal places, and multiple trials will be run to decrease 
+margin of error. Other variables of the experiment will be controlled. For example, the experiment 
+will be run on a same isolated system, a virtual machine, with a minimal amount of 
+processes running to avoid any possible variance in results.
+
+\subsection{Implementing in Python}
+
+The main source for data in this experiment is primary. A Python application 
+was programmed (see appendix) in order to run the two methods aforementioned, and manage 
+the collection of data.
+
+The program made assigns the time before the execution of the method to a variable \verb|t1| 
+with the \verb|time.time()| function. At each iteration of the method, the number of valid decimal 
+places of the resultant approximation are counted and once a specified threshold is reached, 
+a new \verb|t2| time variable is assigned and the time taken, defined by the difference between 
+\verb|t2| and \verb|t1| is stored. This process is repeated for all specified decimal accuracies and 
+for both methods.

### 2021-01-24 research project/madhavaviete.py

@@ -1,0 +1,2 @@
+def test(): 
+    sqdqdsqsd

### 2021-01-28 research project/sections/Background.tex
worked on research project; added python script

@@
-\section{Background Information}
+\section{Background}
 
@@
 
-For the sake of this paper, one by French mathematician François Viète,
-and another supposedly discovered by Madhava of Sangamagrama, and rediscovered 
+In this paper, one by French mathematician François Viète,and another 
+supposedly discovered by Madhava of Sangamagrama, and rediscovered 
 by Swiss mathematician Leibniz. Two methods with different approaches have been chosen for 
@@
 
-\subsubsection{Madhava-Gregory-Leibniz method}
+\subsubsection{Madhava-Leibniz method}
 
-Madhava, having indentified the integral and series for the $\arctan$ function, 
-in the Kerala school during medieval India \cite{amermathmonthly}
-
-TODO... finish
+Madhava is the first mathematician known to have found the series notation for 
+the inverse tangent function, in the Kerala school of mathematics of Medieval India 
+\cite{amermathmonthly}.
 
@@
 So it can be said that $U_{n} = \sqrt{\frac{1}{2} + \frac{1}{2} U_{n - 1}}$, which leads to 
-a fully defined expression for the value of pi:
+a fully defined expression for the value of pi, using an infinite product:
 

### 2021-01-28 research project/madhavaviete.py

@@ -1,2 +1,76 @@
-def test(): 
-    sqdqdsqsd
+import time
+import math
+from mpmath import *
+
+
+RED = "\x1B[31m"
+GREEN = "\x1B[32m"
+RESET = "\x1B[0m"
+mp.dps = 100
+PI_CONST = str(mp.pi)
+
+
+def print_as_text(pi):
+    pi_string = str(pi)
+    print("Constant:       " + PI_CONST)
+    print("Approximation:  ", end="")
+    for i in range(0, len(pi_string)):
+        if pi_string[i] == PI_CONST[i]:
+            print(GREEN +  pi_string[i] + RESET, end="")
+        else:
+            print(RED +  pi_string[i] + RESET, end="")
+    print("\n")
+
+def decimal_is_correct(pi, decvalue):
+    pi_string = str(pi)
+    print(pi_string)
+    if pi_string[decvalue] == PI_CONST[decvalue]:
+        return True
+    else: 
+        return False
+
+def madhavaleibniz(iter):
+    piapprox = 0
+    i = 0 
+
+    t1 = time.time()
+    while not decimal_is_correct(piapprox * mp.sqrt(12), 10):
+        # This is a direct mirror of the summation from the formula
+        piapprox += mp.power(-3, -i) / (2*i+1)
+        i += 1
+    piapprox *= mp.sqrt(12)
+    t2 = time.time()
+
+    print_as_text(piapprox)
+
+    # Return the time spent (t2-t1) and number of iterations
+    return t2-t1
+
+
+
+
+def viete(iter):
+    piapprox = 1
+    numer = 0
+
+    t1 = time.time()
+
+    # Viete's method
+    for i in range(1, iter + 1):
+        numer = mp.sqrt(2.0 + numer)
+        piapprox *= (numer / 2.0)
+    piapprox = (1.0 / piapprox) * 2.0
+
+    t2 = time.time()
+
+    print_as_text(piapprox)
+
+    # Return the time spent (t2-t1) getting d value of decimal places
+    return t2-t1
+
+if __name__ == "__main__":
+    iter = 100
+    trials = 3 
+
+    print("time: ", madhavaleibniz(iter))
+    print("time: ", viete(iter))

### 2021-01-28 research project/madhavaviete.py
rp work

@@ -21,13 +21,17 @@ def print_as_text(pi):
             print(RED +  pi_string[i] + RESET, end="")
     print("\n")
 
+
 def decimal_is_correct(pi, decvalue):
     pi_string = str(pi)
     print(pi_string)
     if pi_string[decvalue] == PI_CONST[decvalue]:
         return True
-    else: 
+    elif pi_string == "0.0":
         return False
+    else:
+        return False
+
 
 def madhavaleibniz(iter):
     piapprox = 0
@@ -35,6 +39,7 @@ def madhavaleibniz(iter):
 
     t1 = time.time()
     while not decimal_is_correct(piapprox * mp.sqrt(12), 10):
+        #for i in range(iter):
         # This is a direct mirror of the summation from the formula
         piapprox += mp.power(-3, -i) / (2*i+1)
         i += 1
@@ -45,8 +50,6 @@ def madhavaleibniz(iter):
 
     # Return the time spent (t2-t1) and number of iterations
     return t2-t1
-
-
 
 
 def viete(iter):
@@ -69,7 +72,7 @@ def viete(iter):
     return t2-t1
 
 if __name__ == "__main__":
-    iter = 100
+    iter = 10
     trials = 3 
 
     print("time: ", madhavaleibniz(iter))

### 2021-01-29 research project/sections/Introduction.tex
""

@@
 and the amount of different approaches to reach the value is large. The value has been 
-found through many processes, be it geometrically, algebraically 
-or through other means.
+found through many processes, be it analytically, the most popular, geometrically, or through 
+more obscure or convoluted methods, such as the possibility to approximate the value using 
+physics from billiards. \cite{galperin_2003}
 
@@
 mathematicians Madhava and Viète, differ in terms of computational efficiency and 
-speed, and explain these differences.
+speed, and explain these differences. This paper does not however, 
+suggest a method to use for computation but rather seeks to compare the efficiency of
+a geometrically derived formula and a algebraically derived one.

### 2021-01-29 research project/sections/Background.tex
""

@@
 represent a first occurence in mathematics: Madhava was the first to 
-find the series notation of the arctangent function and Viète was one 
-of the first mathematicians to use infinite series in his calculation. Furthermore, 
+find the series notation of the arctangent function and Viète was 
+the first mathematician to use and infinite product in calculation. Furthermore, 
 a research based on computational speed of two different kinds approaches to 
@@
 the inverse tangent function, in the Kerala school of mathematics of Medieval India 
-\cite{amermathmonthly}.
+\cite{amermathmonthly}. One of
 
@@
 
-François Viète, having approached the value of 
-$\pi$ from a geometric standpoint, found the following formula:
-
-$$\pi = 2 \frac{2}{\sqrt{2}} \frac{2}{\sqrt{2 + \sqrt{2}}} \frac{2}{\sqrt{2 + \sqrt{2 + \sqrt{2}}}} \dots$$
-
-% $$\pi = \frac{2}{\prod_{k = 0}^{\infty} U_{k}}$$    symbolic
+François Viète approached the value of $\pi$ from a 
+geometric standpoint, and found an infinite product. 
 He was able to calculate $\pi$ to a place of 9 decimal points, in 
@@
 the lengths $OH$ and subsequently $BD$ in \ref{fig:image1}.
-
+% $$\pi = \frac{2}{\prod_{k = 0}^{\infty} U_{k}}$$    symbolic
 With the radius of the circle with center $R = OB$,
@@
 segments from an $2n$-gon, one of which on point $H'$, inscribed in 
-a circle of radius $OB$, adapted from Boris Gourévitch \cite{borisgourevitch2013}}
+a circle of radius $OB$, adapted from (Boris Gourévitch) \cite{borisgourevitch2013}}
 \end{figure}

### 2021-01-29 research project/sections/Computational_approach.tex
worked on research project; added python script
""

@@
 
-The main source for data in this experiment is primary. A Python application 
-was programmed (see appendix) in order to run the two methods aforementioned, and manage 
-the collection of data.
+A Python application was programmed (see appendix) in order to run the two 
+methods aforementioned, and manage the collection of data.
 
 The program made assigns the time before the execution of the method to a variable \verb|t1| 
-with the \verb|time.time()| function. At each iteration of the method, the number of valid decimal 
+with the \verb|time.time()| Python function. At each iteration of the method, the number of valid decimal 
 places of the resultant approximation are counted and once a specified threshold is reached, 
 a new \verb|t2| time variable is assigned and the time taken, defined by the difference between 
-\verb|t2| and \verb|t1| is stored. This process is repeated for all specified decimal accuracies and 
+\verb|t2| and \verb|t1|, is stored. This process is repeated for all specified decimal accuracies and 
 for both methods.
+
+The \verb|mpmath| library was used for the floating point operations required 
+for comparaison between approximated values and the constant $\pi$ that wouldn't 
+have been possible using standard Python libraries \cite{mpmath}. The n

### 2021-01-29 research project/sections/Introduction.tex
rp work

@@
 more obscure or convoluted methods, such as the possibility to approximate the value using 
-physics from billiards. \cite{galperin_2003}
+physics akin to those from a simple game of billiards. \cite{galperin_2003}
 

### 2021-01-29 research project/madhavaviete.py

@@ -1,79 +1,86 @@
 import time
-import math
 from mpmath import *
+import csv 
+import pandas as pd
 
 
-RED = "\x1B[31m"
-GREEN = "\x1B[32m"
-RESET = "\x1B[0m"
 mp.dps = 100
-PI_CONST = str(mp.pi)
+PI_CONST = mp.pi
 
 
-def print_as_text(pi):
-    pi_string = str(pi)
-    print("Constant:       " + PI_CONST)
-    print("Approximation:  ", end="")
-    for i in range(0, len(pi_string)):
-        if pi_string[i] == PI_CONST[i]:
-            print(GREEN +  pi_string[i] + RESET, end="")
-        else:
-            print(RED +  pi_string[i] + RESET, end="")
-    print("\n")
-
-
+# Function that determines if the approximated value of
+# pi is correct to a specified decimal
 def decimal_is_correct(pi, decvalue):
-    pi_string = str(pi)
-    print(pi_string)
-    if pi_string[decvalue] == PI_CONST[decvalue]:
+    #pi_diff = str(abs(pi))
+    zeros = 0
+    pistr = str(pi)[2:]
+    piconst = str(PI_CONST)[2:]
+    for i in range(len(pistr)):
+        if pistr[i] != piconst[i]: break
+        else: 
+            zeros += 1
+    if zeros == decvalue:
         return True
-    elif pi_string == "0.0":
-        return False
     else:
         return False
 
 
-def madhavaleibniz(iter):
+# Function that approximates pi using the Madhava-Leibniz 
+# method
+def madhavaleibniz(decimals):
     piapprox = 0
     i = 0 
 
     t1 = time.time()
-    while not decimal_is_correct(piapprox * mp.sqrt(12), 10):
-        #for i in range(iter):
+    
+    while not decimal_is_correct(piapprox * mp.sqrt(12), decimals):
         # This is a direct mirror of the summation from the formula
         piapprox += mp.power(-3, -i) / (2*i+1)
         i += 1
     piapprox *= mp.sqrt(12)
     t2 = time.time()
 
-    print_as_text(piapprox)
-
-    # Return the time spent (t2-t1) and number of iterations
+    # Return the time spent (t2-t1) getting d value of decimal places
     return t2-t1
 
 
-def viete(iter):
+# Function that approximates pi using Viete's method
+def viete(decimals):
     piapprox = 1
     numer = 0
 
[…]

### 2021-01-29 research project/sections/Analysis_of_results.tex
""
workedo n rp. added graphs and tables

@@
 \section{Analysis of the results}
+\newcolumntype{C}{>{\Centering\arraybackslash\hspace{0pt}}X}
 
@@
 
-\subsection{Analysis of the data}
+\subsubsection{Tabular presentation}
+
+The arithmetic mean of the 100 trials was caluclated and the data points 
+in the table shown below were found (rounded to 3 s.f.):
+\begin{table}[h]
+    \noindent% <-- important
+    \setlength\tabcolsep{3pt} % default: 6pt
+    \begin{tabularx}{\textwidth}{@{} l *{6}{C} @{}}
+    \toprule
+    Decimal place & Average time for approximation using Viete's method, in milliseconds (ms) & Average time for approximation using Madhava's method method, in milliseconds (ms) \\ 
+    \midrule
+    5  & 0.760  & 0.592   \\
+    10 & 1.29  & 1.31    \\
+    15 & 2.05  & 1.9     \\
+    20 & 2.64  & 2.68    \\
+    25 & 3.09  & 3.21    \\
+    30 & 3.82  & 3.9     \\
+    35 & 4.44  & 4.55    \\
+    40 & 5.08  & 5.26    \\
+    45 & 5.71  & 5.84    \\
+    50 & 6.37  & 6.58    \\
+    55 & 7.02  & 7.27    \\
+    60 & 7.72  & 8.00   
+    \end{tabularx}
+\end{table}
+
+\pgfplotstableread{
+    X Y
+    5 0.7601666451
+    10 1.288158894
+    15 2.050116062 
+    20 2.63982296 
+    25 3.091802597
+    30 3.819538355 
+    35 4.442822933
+    40 5.081979036 
+    45 5.706284046 
+    50 6.370962858 
+    55 7.020516396 
+    60 7.723451853   
+}\vietetable
+
+\pgfplotstableread{
+    X Y
+    5 0.5919837952 
+    10 1.307063103  
+    15 1.902475357  
+    20 2.679748535  
+    25 3.212277889  
+    30 3.904948235  
+    35 4.547913074  
+    40 5.255579948  
+    45 5.837199688  
+    50 6.578524113  
+    55 7.270073891  
+    60 8.004565239
+
+}\madhavatable
+
+\subsubsection{Graphical presentation}
+
+To better show the trend in the data collected, the data 
+has been presented in a chart below.
+
+\begin{tikzpicture}[scale=1.2]
+    \begin{axis}[
+        title={Comparing the time taken for approximating the value of $\pi$},
+        xlabel={Decimal places},
+        ylabel={Time taken [ms]},
+        xmin=0, xmax=60,
+        ymin=0, ymax=10,
+        ytick={0,2,4,6,8,10},
+        xtick={0,20,40,60},
+        legend pos=north west,
+        ymajorgrids=true,
+        grid style=dashed,
+    ]
+
+    \addplot[
+        only marks,
+        color=blue,
+        mark=*,
+        ] table {\vietetable};
+        \legend{Viète's method}
+
+    \addplot [thin, blue] table[
+        y={create col/linear regression={y=Y}}
+    ] % compute a linear regression from the input table
+    {\vietetable};
+    \addlegendentry{best fit}
+
+    \addplot[
+        only marks,
+        color=red,
+        mark=*,
+        ] table {\madhavatable};
+        \legend{Madhava's method}
+
+    \addplot [thin, red] table[
+        y={create col/linear regression={y=Y}}
+    ] % compute a linear regression from the input table
+    {\madhavatable};
+    \addlegendentry{best fit}
+
+    \end{axis}
+\end{tikzpicture}
+
+\subsection{Analysis}

### 2021-01-29 research project/sections/Computational_approach.tex
rp work
workedo n rp. added graphs and tables

@@
 \verb|t2| and \verb|t1|, is stored. This process is repeated for all specified decimal accuracies and 
-for both methods.
+for both methods. The times recorded were stored in a \verb|.csv| file for further analysis.
 
 The \verb|mpmath| library was used for the floating point operations required 
-for comparaison between approximated values and the constant $\pi$ that wouldn't 
-have been possible using standard Python libraries \cite{mpmath}. The n
+for comparison between approximated values and the constant $\pi$ that wouldn't 
+have been possible using standard Python libraries \cite{mpmath}.

### 2021-01-29 research project/madhavaviete.py

@@ -62,25 +62,18 @@ def viete(decimals):
 
 
 if __name__ == "__main__":
-    decimals = [1, 10, 20, 30, 40, 50, 60]
+    decimals = [i for i in range(0, 65, 5)]
     trials = 100
+
+    # Exporting data for plotting and analysis
     f = open(r'out.csv', 'w')
     fieldnames = ['decimals', 'viete', 'madhava']
     writer = csv.DictWriter(f, fieldnames=fieldnames)
-    writer.writerow({'decimals':'decimals', 'viete':'viete', 'madhava':'madhava'})
+    writer.writerow({'decimals':'decimals', 'viete':'viete', 
+    'madhava':'madhava'})
 
     for dec in decimals:
-        print("decimal ", dec)
+        for t in range(trials):
+            writer.writerow({'decimals':dec, 'viete':viete(dec), 
+            'madhava':madhavaleibniz(dec)})
 
-        writer.writerow({'decimals':dec, 'viete':viete(dec), 'madhava':madhavaleibniz(dec)})
-
-        print()
-
-    
-
-    '''
-    with open('flip_file.csv', 'w') as f:
-        writer = csv.writer(f)
-        writer.writerows(m)
-        writer.writerows(v)
-    '''

### 2021-01-30 research project/sections/Analysis_of_results.tex
rp

@@
 To better show the trend in the data collected, the data 
-has been presented in a chart below.
+has been presented in a chart below. The blue markers and line of best-fit 
+respresent the results received from Viète's method while the red represent
 
-\begin{tikzpicture}[scale=1.2]
+\begin{tikzpicture}[scale=1.1]
     \begin{axis}[
@@
         grid style=dashed,
+        scale=1.2
     ]
@@
     {\vietetable};
-    \addlegendentry{best fit}
+    \addlegendentry{best fit line}
 
@@
         ] table {\madhavatable};
-        \legend{Madhava's method}
+        \addlegendentry{Madhava's method}
 
@@
     {\madhavatable};
-    \addlegendentry{best fit}
+    \addlegendentry{best fit line}
 

### 2021-01-30 research project/sections/Background.tex
rp

@@
 
-In this paper, one by French mathematician François Viète,and another 
-supposedly discovered by Madhava of Sangamagrama, and rediscovered 
-by Swiss mathematician Leibniz. Two methods with different approaches have been chosen for 
+In this paper, we focus on two methods for approximating the value of $\pi$. 
+The first by French mathematician François Viète, and another discovered by Madhava of Sangamagrama and 
+rediscovered by Swiss mathematician Leibniz. Two methods with different approaches have been chosen for 
 comparison, the process for the original discovery of these methods
@@
 represent a first occurence in mathematics: Madhava was the first to 
-find the series notation of the arctangent function and Viète was 
-the first mathematician to use and infinite product in calculation. Furthermore, 
-a research based on computational speed of two different kinds approaches to 
-the constant $\pi$ has not been done to date.
+find the series notation of the inverse tangent function and Viète was 
+the first mathematician to use and infinite product in calculation.
 

### 2021-01-30 research project/sections/Computational_approach.tex
rp

@@
 will be run on a same isolated system, a virtual machine, with a minimal amount of 
-processes running to avoid any possible variance in results.
+processes running simultaneously to avoid any possible variance in results.
 
@@
 
-The program made assigns the time before the execution of the method to a variable \verb|t1| 
+The program assigns the time before the execution of the method to a variable \verb|t1| 
 with the \verb|time.time()| Python function. At each iteration of the method, the number of valid decimal 

### 2021-01-30 research project/sections/Introduction.tex
rp

@@
 approximation of the value $\pi$, namely the approaches suggested by the aforementioned 
-mathematicians Madhava and Viète, differ in terms of computational efficiency and 
+mathematicians Madhava and Viète, differ in terms of computational speed and 
 speed, and explain these differences. This paper does not however, 
-suggest a method to use for computation but rather seeks to compare the efficiency of
-a geometrically derived formula and a algebraically derived one.
+suggest a method to use for computation but rather seeks to only compare 
+the speed of a geometrically derived formula and a analytically derived one.
+
+This research could prove useful in the field of computer science, as 
+there is always a demand for faster and more efficient programs in an
+everchanging society.
+
+Furthermore, a research based on computational speed of two different 
+kinds approaches to the constant $\pi$ has not been done to date.

### 2021-01-31 research project/sections/Background.tex
rp work

@@
 rediscovered by Swiss mathematician Leibniz. Two methods with different approaches have been chosen for 
-comparison, the process for the original discovery of these methods
-will be explained. The two methods in question are one based on 
+comparison, the process for original discovery of these methods
+will be explained. The two in question are one based on 
 the infinite series definition of an inverse trigonometric function 
@@
 Viète. These two mathematical methods were chosen as they are both 
-represent a first occurence in mathematics: Madhava was the first to 
-find the series notation of the inverse tangent function and Viète was 
+represent a first occurrence in mathematics: Madhava was the first to 
+find the infinite series expansion of the inverse tangent function and Viète was 
 the first mathematician to use and infinite product in calculation.
@@
 the inverse tangent function, in the Kerala school of mathematics of Medieval India 
-\cite{amermathmonthly}. One of
+\cite{amermathmonthly}. The research made at the Kerala school was documented by 
+many mathematicias of their time, namely the astronomer-mathematician Jyesthadeva 
+in his treatise \textit{Yukti-Bhasa} written in the Malayam language \cite{oconnor_robertson_2000}. 
+Madhava's infinite series expansion of the inverse tangent function can also be 
+found in this treatise in this translation, adapted from (R. C. Gupta 67-70) 
+\cite{rc_gupta_mgseries}:
+
+\textit{The first term is the product of the given sine and radius of the desired arc divided 
+by the cosine of the arc. The succeeding terms are obtained by a process of 
+iteration when the first term is repeatedly multiplied by the square of the 
+sine and divided by the square of the cosine. All the terms are then divided by 
+the odd numbers 1, 3, 5, ... The arc is obtained by adding and subtracting 
+respectively the terms of odd rank and those of even rank. It is laid down that 
+the sine of the arc or that of its complement whichever is the smaller should be 
+taken here as the given sine. Otherwise the terms obtained by this above iteration 
+will not tend to the vanishing magnitude.}
 
@@
 \label{fig:image1}
-\caption{Circle with 1 segment from a $n$-gon with point $H$ and 2 
+\caption[]{Circle with 1 segment from a $n$-gon with point $H$ and 2 
 segments from an $2n$-gon, one of which on point $H'$, inscribed in 

### 2021-01-31 research project/sections/Introduction.tex
rp work

@@
 there is always a demand for faster and more efficient programs in an
-everchanging society.
+ever-changing society.
 

### 2021-01-31 research project/sections/Analysis_of_results.tex
rp work
rp

@@
 
-The arithmetic mean of the 100 trials was caluclated and the data points 
+The arithmetic mean of a decimal place was caluclated for 100 trials and the data points 
 in the table shown below were found (rounded to 3 s.f.):
@@
 has been presented in a chart below. The blue markers and line of best-fit 
-respresent the results received from Viète's method while the red represent
+respresent the results received from Viète's method while the red represent those 
+from Madhava's method.
 
@@
 \subsection{Analysis}
+
+The results gathered from this experiment are descisive.

### 2021-01-31 research project/sections/Background.tex
rp

--- 2021-01-30 research project/sections/Background.tex
@@
 The first by French mathematician François Viète, and another discovered by Madhava of Sangamagrama and 
-rediscovered by Swiss mathematician Leibniz. Two methods with different approaches have been chosen for 
-comparison, the process for the original discovery of these methods
-will be explained. The two methods in question are one based on 
-the infinite series definition of an inverse trigonometric function 
+rediscovered more than a hundred years later by James Gregory. Two methods with different 
+approaches have been chosen for comparison and the process for original discovery of these methods
+will be explained. The two in question are one based on 
+the infinite series definition of an inverse trigonometric function discovered by Madhava
 and one where the value is derived using geometry, by mathematician 
 Viète. These two mathematical methods were chosen as they are both 
-represent a first occurence in mathematics: Madhava was the first to 
-find the series notation of the inverse tangent function and Viète was 
-the first mathematician to use and infinite product in calculation.
+represent a first occurrence in mathematics: Madhava was the first to 
+find the infinite series expansion of the inverse tangent function and Viète was 
+the first mathematician to use an infinite product in calculation.
 
-\subsubsection{Madhava-Leibniz method}
+\subsubsection{An analytical method: Madhava's method}
 
@@
 the inverse tangent function, in the Kerala school of mathematics of Medieval India 
-\cite{amermathmonthly}. One of
+\cite{amermathmonthly}. The research made at the Kerala school was documented by 
+many mathematicians of their time, namely the astronomer-mathematician Jyesthadeva 
+in his treatise \textit{Yukti-Bhasa} written in the Malayalam language \cite{oconnor_robertson_2000}. 
+Madhava's infinite series expansion of the inverse tangent function can also be 
+found in this treatise. R. C. Gupta, in his translation, describes this expansion as one 
+where the arc $\theta$ is equal to the sum of a first term, "the product of the given sine 
+and radius of the desired arc divided by the cosine of the arc" \footnotemark, followed by terms that "are 
+obtained by a process of iteration": . \cite{rc_gupta_mgseries}
 
-\subsubsection{Viète's method}
+\footnotetext{Note: the old Indian meaning for the sine of $\theta$ is $r \sin{\theta}$, where $r$ is the radius. The same applies for the cosine.}
+
+This text explains what can be written in modern mathematical terms as such:
+
+$r \theta = \frac{r (r \sin{\theta})}{1 (r \cos{\theta})} - \frac{r (r \sin{\theta})^3}{3 (r \cos{\theta})^3} + \frac{r (r \sin{\theta})^5}{5 (r \cos{\theta})^5} - \frac{r (r \sin{\theta})^7}{7 (r \cos{\theta})^7} + \dots$
+
+And for a circle of radius $r = 1$, we can cancel all $r$ terms:
+
+$\theta = \frac{\sin{\theta}}{\cos{\theta}} - \frac{\sin^3{\theta}}{3 \cos^3{\theta}} + \frac{\sin^5{\theta}}{5 \cos^5{\theta}} - \frac{\sin^7{\theta}}{7 \cos^7{\theta}} + \dots$
+
+And because $\frac{\sin{\theta}}{\cos{\theta}} = \tan{\theta}$, the aforementioned expression 
+can be simplified to:
+
+$\theta = \tan{\theta} - \frac{\tan^3{\theta}}{3} + \frac{\tan^5{\theta}}{5} - \frac{\tan^7{\theta}}{7} + \dots$
+
+So then, if we let $\tan{\theta} = \alpha$, we can find the infinite series expansion 
+of the arctangent function.
+
+$\arctan{\alpha} = \alpha - \frac{\alpha^3}{3} + \frac{\alpha^5}{5} - \frac{\alpha^7}{7} + \dots = \theta$
+
+Moreover, it is said in a further commentary describing the findings of Madhava that a  
+method for the calculation of the circumference $c$ of a circle of diameter $d$ exists. 
+The passage in question from the commentary \textit{Tantrasamgraha-vyakhya} of 
+anonymous authorship states that, by following the same argument as stated before, one 
+can find the circumference of a circle through a similar infinite sum. The first 
+term of this sum would be "the square root of the square of the diameter multiplied by 
+twelve", followed by the first term "divided by three in each successive case", and 
+when these "are divided in order by the odd numbers, beginning with 1", and after the 
+even terms are subtracted from sum of the odd", one is left with the circumference of the 
+circle (translation from C. K. Raju). \cite{raju_2007}
+
+This can be expressed in mathematical terms as such:
+
+\subsubsection{A geometric method: Viète's method}
 
@@
 can be found using simple right-angle trigonometry, by first finding 
-the lengths $OH$ and subsequently $BD$ in \ref{fig:image1}.
+the lengths $OH$ and subsequently $BD$ in (Figure 1 ) \ref{fig:image1}.
 % $$\pi = \frac{2}{\prod_{k = 0}^{\infty} U_{k}}$$    symbolic
@@
 So $P = \frac{A_{n}}{A}$, since the values $A_{kn}$ cancel, and 
-followingly, it is true that $P = \frac{A_{n}}{A} \Leftrightarrow A = \frac{A_{n}}{P}$. The 
+so it is true that $P = \frac{A_{n}}{A} \Leftrightarrow A = \frac{A_{n}}{P}$. The 
 value of $R = 1$ in this case, and since the area of a circle is defined 
@@
 By definition, we can say that as the value $k$ approaches infinity, the area of the $kn$-gon 
-aproaches that of a circle, and therefore, the value of $\pi$.
+approaches that of a circle, and therefore, the value of $\pi$.
 

### 2021-01-31 research project/madhavaviete.py

@@ -25,9 +25,8 @@ def decimal_is_correct(pi, decvalue):
         return False
 
 
-# Function that approximates pi using the Madhava-Leibniz 
-# method
-def madhavaleibniz(decimals):
+# Function that approximates pi using Madhava's method
+def madhava(decimals):
     piapprox = 0
     i = 0 
 
@@ -75,5 +74,5 @@ if __name__ == "__main__":
     for dec in decimals:
         for t in range(trials):
             writer.writerow({'decimals':dec, 'viete':viete(dec), 
-            'madhava':madhavaleibniz(dec)})
+            'madhava':madhava(dec)})
 

### 2021-01-31 research project/sections/Background.tex
rp

@@
 In this paper, we focus on two methods for approximating the value of $\pi$. 
-The first by French mathematician François Viète, and another discovered by Madhava of Sangamagrama and 
-rediscovered more than a hundred years later by James Gregory. Two methods with different 
+The first by French mathematician François Viète, and another discovered by Madhava of Sangamagrama. Two methods with different 
 approaches have been chosen for comparison and the process for original discovery of these methods
@@
 
-Moreover, it is said in a further commentary describing the findings of Madhava that a  
-method for the calculation of the circumference $c$ of a circle of diameter $d$ exists. 
+\textit{Yukti-Bhasa} also describes two methods for the calculation of $\pi$. Firstly, 
+Madhava's infinite series expansion for $\pi$ is described, which he obtained through 
+the previous expansion for the arctangent function. And since $\arctan{1} = \frac{\pi}{4}$,
+it can be said that:
+
+$\frac{\pi}{4} = \arctan{1} = 1 - \frac{1}{3} + \frac{1}{5} - \frac{1}{7} + \dots + \frac{(-1)^n}{2n +1}$
+
+Which in form of an infinite sum can be expressed as:
+
+$\frac{\pi}{4} = \sum\limits_{n=0}^\infty \frac{(-1)^n}{2n +1}$
+
+Which has come to be known as the Leibniz formula for $\pi$, after the German mathematician 
+who discovered the same formula two decades later \cite{edwards_1994}
+
+However, the method with a faster convergence can be found in a further commentary describing the findings of Madhava that states  
+a method for the calculation of the circumference $c$ of a circle of diameter $d$ exists. 
 The passage in question from the commentary \textit{Tantrasamgraha-vyakhya} of 

### 2021-01-31 research project/sections/Background.tex
rp

@@
 Which has come to be known as the Leibniz formula for $\pi$, after the German mathematician 
-who discovered the same formula two decades later \cite{edwards_1994}
+who discovered the same formula two decades later. \cite{edwards_1994}
 
@@
 
-This can be expressed in mathematical terms as such:
+This can be expressed in mathematical terms as such, where we let $c$ be the circumference of 
+a circle of diameter $d$:
+
+$c = \sqrt{12 d^2} - \frac{\sqrt{12 d^2}}{3 \cdot 3} + \frac{\sqrt{12 d^2}}{3^2 \cdot 5} - \frac{\sqrt{12 d^2}}{3^3 \cdot 7} + \dots$
+
+And since $c = \pi d$, for a circle of diameter 1, $c = \pi$, all $d$ terms cancel and 
+the expression can be factorized as:
+
+$c = \pi = \sqrt{12} ( 1 - \frac{1}{3 \cdot 3} + \frac{1}{3^2 \cdot 5} - \frac{1}{3^3 \cdot 7} + \dots )$
+
+Or as an infinite sum:
+
+$\pi = \sqrt{12} \sum\limits_{n=0}^\infty \frac{(-3)^{-n}}{2n+1}$
 
@@
 
-These expressions, when under a single expression result in the aforementioned 
-formula with nested roots.
-
 \footnotetext{where $\prod$ signifies a product. Similar expression to $\sum$ }

### 2021-01-31 research project/sections/Computational_approach.tex
rp

@@
 for comparison between approximated values and the constant $\pi$ that wouldn't 
-have been possible using standard Python libraries \cite{mpmath}.
+have been possible using standard Python libraries. \cite{mpmath}

### 2021-01-31 research project/sections/Further_research.tex
rp

\section{Further research opportunities}

\subsection{}

### 2021-02-01 research project/sections/Analysis_of_results.tex
research project

@@
         ytick={0,2,4,6,8,10},
-        xtick={0,20,40,60},
+        xtick={0,10,20,30,40,50,60},
         legend pos=north west,
@@
 
-The results gathered from this experiment are descisive.
+The results gathered from this experiment show linear relationship between the amount of decimal 
+places approximated and the time required for this approximation. The two lines of best fit for 
+the two methods show similarities but also differences. Firstly, from 5 to 20 decimal places, 
+Madhava's

### 2021-02-01 research project/sections/Background.tex
rp
research project

@@
 where the arc $\theta$ is equal to the sum of a first term, "the product of the given sine 
-and radius of the desired arc divided by the cosine of the arc" \footnotemark, followed by terms that "are 
-obtained by a process of iteration": . \cite{rc_gupta_mgseries}
+and radius of the desired arc divided by the cosine of the arc"\footnotemark, followed by terms that "are 
+obtained by a process of iteration": in which the original term is multiplied by the 
+square of the sine and divided by the square of the cosine. Thereafter, each term is 
+divided by an odd number in order (1, 3, 5, 7, \dots). It is then said that the arc can 
+be found by "adding and subtracting respectively the terms of odd rank and those of even rank",  
+definition of an alternating series. \cite{rc_gupta_mgseries}
 
@@
 Which has come to be known as the Leibniz formula for $\pi$, after the German mathematician 
-who discovered the same formula two decades later. \cite{edwards_1994}
+who discovered the same formula two decades later. \cite{edwards_1994} This series converges to $\pi$, as seen in (Figure 1) \ref{fig:numberline}
 
@@
 can be found using simple right-angle trigonometry, by first finding 
-the lengths $OH$ and subsequently $BD$ in (Figure 1 ) \ref{fig:image1}.
+the lengths $OH$ and subsequently $BD$ in (Figure 2) \ref{fig:image1}.
 % $$\pi = \frac{2}{\prod_{k = 0}^{\infty} U_{k}}$$    symbolic
@@
 
+\begin{figure}[h]
+
+    \begin{tikzpicture}
+        \draw[latex-latex] (-0.5,0) -- (4.5,0);% the x-axis
+        \foreach \x in  {0,...,4} % tick marks
+          \draw (\x,0) -- (\x,-3pt);
+        \foreach \x in {0,...,4} % the numbers
+          \node [below] at (\x,-0.1) {$\x$};
+        \node [below] at (3.14,0.5) {$\pi$};
+        \draw (3.14,0) -- (3.14,3pt);
+
+        \draw[->, dotted] (3.14,0) -- (3.14,-3);% the x-axis
+
+        \draw[xshift=-1cm, dotted] (1,-0.7) node[circle,fill,inner sep=1pt](a){} -- (5,-0.7) 
+node[circle,fill,inner sep=1pt,label=right:$+\frac{4}{1}$](b){};
+
+\draw[xshift=-1cm, dotted] (3.67,-1.1) node[circle,fill,inner sep=1pt, label=left:$-\frac{4}{3}$](c){} -- (5,-1.1) 
+node[circle,fill,inner sep=1pt](d){};
+
+\draw[xshift=-1cm, dotted] (3.67,-1.5) node[circle,fill,inner sep=1pt](e){} -- (4.46,-1.5) 
+node[circle,fill,inner sep=1pt, label=right:$+\frac{4}{5}$](f){};
+
+\draw[xshift=-1cm, dotted] (3.89,-1.9) node[circle,fill,inner sep=1pt, label=left:$-\frac{4}{7}$](g){} -- (4.46,-1.9) 
+node[circle,fill,inner sep=1pt](h){};
+
+\draw[xshift=-1cm, dotted] (3.89,-2.3) node[circle,fill,inner sep=1pt](i){} -- (4.33,-2.3) 
+node[circle,fill,inner sep=1pt, label=right:$+\frac{4}{9}$](j){};
+
+\draw[->, to path={-| (\tikztotarget)}, dotted]
+  (b) edge (d) (c) edge (e) (f) edge (h) (g) edge (i);
+     \end{tikzpicture}
+
+\label{fig:numberline}
+\caption[]{A number line demonstrating the convergence of Madhava's first method for the 
+calculation of $\pi$. Adapted from (V. M. Jamkar) \cite{jamkar_2018}}
+\end{figure}
+
 \begin{figure}[!ht]

### 2021-02-01 research project/sections/Analysis_of_results.tex
rp

@@
 in the table shown below were found (rounded to 3 s.f.):
+
 \begin{table}[h]
@@
 To better show the trend in the data collected, the data 
-has been presented in a chart below. The blue markers and line of best-fit 
+has been presented in a scatter chart below. The blue markers and line of best-fit 
 respresent the results received from Viète's method while the red represent those 
@@
 
-\subsection{Analysis}
+\subsection{Observations and analysis}
 
@@
 places approximated and the time required for this approximation. The two lines of best fit for 
-the two methods show similarities but also differences. Firstly, from 5 to 20 decimal places, 
-Madhava's
+the two methods show similarities but also differences. It can be firstly noted that Madhava's 
+method is faster than the other until the decimal 20, when the two best fit lines diverge. This 
+can be explained as due to the convergence rates of the two methods, which can be seen in (Figure 3) 
+\ref{fig:comparaison}. In the beginning of this curve representing the convergence of Viète's method, 
+it cna be seen that it does not follow an \textit{alternating} method, like the one formulated by 
+Madhava. Madhava's method is what is called an alternating series, in this case one that converges 
+to the value $\pi$ \footnote{Note: not all alternating series converge}
+
+\begin{figure}
+    \includegraphics[width=\linewidth]{image.png}
+    \caption{Comparison of historical methods of approximating the value of $\pi$. This image demonstrates
+    their convergence rates. The line labelled Madhava in dark blue is the method used in this paper. 
+    From (Wikimedia Commons) \cite{infinite_series_comparaison}}
+    \label{fig:comparaison}
+  \end{figure}

### 2021-02-01 research project/test.py

@@ -1,0 +1,65 @@
+import time
+from mpmath import *
+import csv 
+import pandas as pd
+
+
+mp.dps = 100
+PI_CONST = mp.pi
+
+
+# Function that determines if the approximated value of
+# pi is correct to a specified decimal
+def decimal_is_correct(pi, decvalue):
+    #pi_diff = str(abs(pi))
+    zeros = 0
+    pistr = str(pi)[2:]
+    piconst = str(PI_CONST)[2:]
+    for i in range(len(pistr)):
+        if pistr[i] != piconst[i]: break
+        else: 
+            zeros += 1
+    if zeros == decvalue:
+        return True
+    else:
+        return False
+
+def madhava(decimals):
+    piapprox = 0
+    i = 0 
+
+    t1 = time.time()
+    
+    while not decimal_is_correct(piapprox * mp.sqrt(12), decimals):
+        # This is a direct mirror of the summation from the formula
+        piapprox += mp.power(-3, -i) / (2*i+1)
+        i += 1
+    piapprox *= mp.sqrt(12)
+    t2 = time.time()
+
+    # Return the time spent (t2-t1) getting d value of decimal places
+    return i
+
+
+# Function that approximates pi using Viete's method
+def viete(decimals):
+    piapprox = 1
+    numer = 0
+    i= 0
+
+    while not decimal_is_correct((1.0 / piapprox) * 2.0, decimals):
+        numer = mp.sqrt(2.0 + numer)
+        piapprox *= (numer / 2.0)
+        i += 1
+    piapprox = (1.0 / piapprox) * 2.0
+
+
+    # Return the time spent (t2-t1) getting d value of decimal places
+    return i
+
+if __name__ == "__main__":
+    decimals = [68]
+
+    for dec in decimals:
+        print(viete(dec))
+        print(madhava(dec))

### 2021-02-02 research project/sections/Analysis_of_results.tex
""

@@
 \newcolumntype{C}{>{\Centering\arraybackslash\hspace{0pt}}X}
-
-\subsection{Presentation of the data}
-
-\subsubsection{Tabular presentation}
-
-The arithmetic mean of a decimal place was caluclated for 100 trials and the data points 
-in the table shown below were found (rounded to 3 s.f.):
-
-\begin{table}[h]
-    \noindent% <-- important
-    \setlength\tabcolsep{3pt} % default: 6pt
-    \begin{tabularx}{\textwidth}{@{} l *{6}{C} @{}}
-    \toprule
-    Decimal place & Average time for approximation using Viete's method, in milliseconds (ms) & Average time for approximation using Madhava's method method, in milliseconds (ms) \\ 
-    \midrule
-    5  & 0.760  & 0.592   \\
-    10 & 1.29  & 1.31    \\
-    15 & 2.05  & 1.9     \\
-    20 & 2.64  & 2.68    \\
-    25 & 3.09  & 3.21    \\
-    30 & 3.82  & 3.9     \\
-    35 & 4.44  & 4.55    \\
-    40 & 5.08  & 5.26    \\
-    45 & 5.71  & 5.84    \\
-    50 & 6.37  & 6.58    \\
-    55 & 7.02  & 7.27    \\
-    60 & 7.72  & 8.00   
-    \end{tabularx}
-\end{table}
 
@@
 
+\subsection{Presentation of the data}
+
+\subsubsection{Tabular presentation}
+
+The arithmetic mean of a decimal place was caluclated for 100 trials and the data points 
+in the table shown below were found (rounded to 3 s.f.):
+
+\begin{table}[h]
+    \noindent% <-- important
+    \setlength\tabcolsep{3pt} % default: 6pt
+    \begin{tabularx}{\textwidth}{@{} l *{6}{C} @{}}
+    \toprule
+    Decimal place & Average time for approximation using Viete's method, in milliseconds (ms) & Average time for approximation using Madhava's method method, in milliseconds (ms) \\ 
+    \midrule
+    5  & 0.760  & 0.592   \\
+    10 & 1.29  & 1.31    \\
+    15 & 2.05  & 1.9     \\
+    20 & 2.64  & 2.68    \\
+    25 & 3.09  & 3.21    \\
+    30 & 3.82  & 3.9     \\
+    35 & 4.44  & 4.55    \\
+    40 & 5.08  & 5.26    \\
+    45 & 5.71  & 5.84    \\
+    50 & 6.37  & 6.58    \\
+    55 & 7.02  & 7.27    \\
+    60 & 7.72  & 8.00   
+    \end{tabularx}
+\end{table}
+
+\begin{table}[h]
+    \noindent% <-- important
+    \setlength\tabcolsep{3pt} % default: 6pt
+    \begin{tabularx}{\textwidth}{@{} l *{6}{C} @{}}
+    \toprule
+    Decimal place & Iterations needed, Viète's method (no unit) & Iterations needed, Madhava's method (no unit)\\
+    \midrule
+    5 & 10 & 9 \\
+    10 & 17 & 20 \\
+    15 & 27 & 29 \\
+    20 & 35 & 41 \\
+    25 & 42 & 50 \\
+    30 & 51 & 61 \\
+    35 & 59 & 71 \\
+    40 & 67 & 82 \\
+    45 & 76 & 91 \\
+    50 & 84 & 102 \\
+    55 & 92 & 112 \\
+    60 & 101 & 123
+    \end{tabularx}
+\end{table}
+
+The iterations needed for both methods were also measured, using the Python program 
+(see appendix), slightly modified to increment a variable \verb|i| and subsequently 
+return it, writing it in a similar manner to a \verb|.csv| file.
+
 \subsubsection{Graphical presentation}
 
-To better show the trend in the data collected, the data 
-has been presented in a scatter chart below. The blue markers and line of best-fit 
+To better show the trend in the data collected, see below the decimal places compared 
+to the time taken per method. The blue markers and line of best-fit 
 respresent the results received from Viète's method while the red represent those 
-from Madhava's method.
+from Madhava's method. \\
 
@@
 
+Include Decimal places vs Iterations
+
 \subsection{Observations and analysis}
@@
 \ref{fig:comparaison}. In the beginning of this curve representing the convergence of Viète's method, 
-it cna be seen that it does not follow an \textit{alternating} method, like the one formulated by 
+it can be seen that it does not follow an \textit{alternating} method, like the one formulated by 
 Madhava. Madhava's method is what is called an alternating series, in this case one that converges 
-to the value $\pi$ \footnote{Note: not all alternating series converge}
+to the value $\pi$, and as such it is closer to the correct value of $\pi$ at lower decimal values 
+and gradually takes longer and longer than Viète's method, which is an infinite product. This difference 
+in speeds can be attributed to the fact that
 
-\begin{figure}
+\begin{figure}[h]
     \includegraphics[width=\linewidth]{image.png}

### 2021-02-02 research project/madhavaviete.py

@@ -1,7 +1,6 @@
 import time
 from mpmath import *
 import csv 
-import pandas as pd
 
 
 mp.dps = 100

### 2021-02-02 research project/test.py

@@ -1,7 +1,6 @@
 import time
 from mpmath import *
 import csv 
-import pandas as pd
 
 
 mp.dps = 100
@@ -58,8 +57,13 @@ def viete(decimals):
     return i
 
 if __name__ == "__main__":
-    decimals = [68]
+    decimals = [i for i in range(0, 65, 5)]
+    f = open(r'out.csv', 'w')
+    fieldnames = ['decimals', 'viete', 'madhava']
+    writer = csv.DictWriter(f, fieldnames=fieldnames)
+    writer.writerow({'decimals':'decimals', 'viete':'viete', 
+    'madhava':'madhava'})
 
     for dec in decimals:
-        print(viete(dec))
-        print(madhava(dec))
+        writer.writerow({'decimals':dec, 'viete':viete(dec), 
+            'madhava':madhava(dec)})

### 2021-02-02 research project/sections/Analysis_of_results.tex
""

@@
 The arithmetic mean of a decimal place was caluclated for 100 trials and the data points 
-in the table shown below were found (rounded to 3 s.f.):
+in the table shown below were found (rounded to 3 significant figures). The values for the decimal place 
+0 have been dropped as it would not be logical to include them.
 

### 2021-02-02 research project/sections/Analysis_of_results.tex
rp

@@
 }\madhavatable
+
+\pgfplotstableread{
+    X Y
+    5 10
+    10 17
+    15 27
+    20 35
+    25 42
+    30 51
+    35 59
+    40 67
+    45 76
+    50 84
+    55 92
+    60 101
+}\vietetableiter
+
+\pgfplotstableread{
+    X Y
+    5 9
+    10 20
+    15 29
+    20 41
+    25 50
+    30 61
+    35 71
+    40 82
+    45 91
+    50 102
+    55 112
+    60 123
+}\madhavatableiter
 
@@
 
-To better show the trend in the data collected, see below the decimal places compared 
-to the time taken per method. The blue markers and line of best-fit 
-respresent the results received from Viète's method while the red represent those 
+Shown below is a graph demonstrating the correlation between the decimal places approximated 
+and the amount of iterations needed for this specific decimal milestone. The blue markers and 
+line of best-fit respresent the results received from Viète's method while the red represent those 
 from Madhava's method. \\
+
+\begin{tikzpicture}[scale=1.1]
+    \begin{axis}[
+        title={Comparing the iterations needed for approximating the value of $\pi$},
+        xlabel={Decimal places},
+        ylabel={Iterations},
+        xmin=0, xmax=60,
+        ymin=0, ymax=130,
+        ytick={0,10,20,30,40,50,60,70,80,90,100,110,120,130},
+        xtick={0,10,20,30,40,50,60},
+        legend pos=north west,
+        ymajorgrids=true,
+        grid style=dashed,
+        scale=1.2
+    ]
+
+    \addplot[
+        only marks,
+        color=blue,
+        mark=*,
+        ] table {\vietetableiter};
+        \legend{Viète's method}
+
+    \addplot [thin, blue] table[
+        y={create col/linear regression={y=Y}}
+    ] % compute a linear regression from the input table
+    {\vietetableiter};
+    \addlegendentry{best fit line}
+
+    \addplot[
+        only marks,
+        color=red,
+        mark=*,
+        ] table {\madhavatableiter};
+        \addlegendentry{Madhava's method}
+
+    \addplot [thin, red] table[
+        y={create col/linear regression={y=Y}}
+    ] % compute a linear regression from the input table
+    {\madhavatableiter};
+    \addlegendentry{best fit line}
+
+    \end{axis}
+\end{tikzpicture}
+
+To show the trend in the time data collected, see below the decimal places compared 
+to the time taken per method. The color coding here is relevant to the one aforementioned. \\
 
@@
 
-Include Decimal places vs Iterations
-
 \subsection{Observations and analysis}
@@
 The results gathered from this experiment show linear relationship between the amount of decimal 
-places approximated and the time required for this approximation. The two lines of best fit for 
-the two methods show similarities but also differences. It can be firstly noted that Madhava's 
-method is faster than the other until the decimal 20, when the two best fit lines diverge. This 
-can be explained as due to the convergence rates of the two methods, which can be seen in (Figure 3) 
-\ref{fig:comparaison}. In the beginning of this curve representing the convergence of Viète's method, 
-it can be seen that it does not follow an \textit{alternating} method, like the one formulated by 
-Madhava. Madhava's method is what is called an alternating series, in this case one that converges 
-to the value $\pi$, and as such it is closer to the correct value of $\pi$ at lower decimal values 
-and gradually takes longer and longer than Viète's method, which is an infinite product. This difference 
-in speeds can be attributed to the fact that
+places approximated and the time required for this approximation, as well as between the decimal places 
+and the amount of iterations needed. The two lines of best fit for the two methods show 
+similarities but also differences.
+
+Where decimal places and iterations are compared, there is a clear difference between the two 
+methods: Madhava's method indeed takes more iterations to converge to a specific decimal of 
+the value $\pi$. However, in the second graph comparing time, it can be noted that Madhava's 
+method is marginally faster until the decimal 20, when the two methods diverge. It could be 
+assumed that this occurrence is due to differences between the types of the two methods. Madhava's 
+method is what is called an alternating series, due to its nature of alternating between values in order 
+to converge to a specific value, in this case $\pi$, while Viète's method is an infinite product. The 
+difference between these two methods is underlined in (Figure 3) \ref{fig:comparaison}. It can be 
+deduced that at lower decimal values Madhava's method is faster as Viète's method's approximation 
+accuracy increases in what looks like an exponential fashion.
+
+\textit{More info to be added}
 

### 2021-02-02 research project/sections/Conclusion.tex
rp

\section{Conclusion}

To conclude, even though Viète's method can seem more convoluted in practice, due to its geometric origins, 
it can be seen through the experiment conducted in this paper that this is clearly not the case 
in a computational context. His method is simplified to a point where its geometric origins are no 
longer findable from its computational implementation. And despite the assumption that could be 
made based on the types of operations needed for each method, Viète's requiring arguably more 
computationally challenging operations. These operations, such as square-root function or the product function 
used on high-precision floating point numbers, can be said to be more demanding than what could be said 
is simple summation, in the case of Madhava. This essay could be considered in defense of 
what is known as geometric algebra --- which can clearly produce results more capable than ones 
originating from an analytical method.

\subsection{Evaluation of the experimental method}

It must be noted that the experimental procedure in this paper is far from perfect. Firstly, 
there are many variables that would be difficult to control.

\textit{More info to be added}

### 2021-02-03 research project/sections/Analysis_of_results.tex
rp

@@
 deduced that at lower decimal values Madhava's method is faster as Viète's method's approximation 
-accuracy increases in what looks like an exponential fashion.
+accuracy firstly increases in what seems similar to a curve from an exponential function $f : x \to -2x^2$. 
+It can also be said that while Madhava's method alternates between values such that $m_{n} > \pi, m_{n+1} < \pi$, 
+Viète's method approaches pi in a way that the approximation is always $v < \pi$.
 

### 2021-02-17 math/ia_2021/paper/paper.tex
math ia prliminary edits

% !TEX root = paper.tex

\documentclass[12pt,a4paper]{article}

% Packages
\usepackage{packages}

% Document Settings
\geometry{margin=1in}
\setlength{\parskip}{1ex}
\setlength{\parindent}{0pt}

% =========================================
%             DOCUMENT
% =========================================

\begin{document}
\doublespacing % Double spacing throughout the document

% =========================
%      TITLE PAGE
% =========================

% Suppresses headers, footers, and page numbers on title page
\begin{titlepage}
    \begin{center}
    \vspace*{4cm}
        Comparing the speed of two methods of approximating the value of $\pi$ ---
        a computational approach\\
    \vspace{1cm}
    To what extent can a method of approximation of the value $\pi$ be computationally more efficient than another? \\
    \vspace{4cm}
    Word count: 2122
    \vfill
    \vspace{0.1cm}
    \end{center}
    \end{titlepage}

% =========================
%      DOCUMENT BODY
% =========================

\pagenumbering{roman}

\begin{center}
\pdfbookmark{\contentsname}{Contents}
\tableofcontents
\vspace{1in}

\end{center}

% TODO: Add hyper-references to links
% TODO: Improve listings' styling
% TODO: Add links to Contents entries

\newpage

\pagenumbering{arabic}

\include{sections/Introduction}

% Just testing sections.. 
\newpage
\include{sections/Implementation}

\end{document}

### 2021-02-19 math/ia_2021/paper/paper.tex
math ia

@@
     \vspace*{4cm}
-        Comparing the speed of two methods of approximating the value of $\pi$ ---
-        a computational approach\\
+    A statistical approach to predicting a tsunami’s wave height \\
     \vspace{1cm}
-    To what extent can a method of approximation of the value $\pi$ be computationally more efficient than another? \\
+    Can the height of a tsunami's wave be predicted based on the distance from and 
+    the magnitude of a preceding earthquake? \\
     \vspace{4cm}
-    Word count: 2122
+    Word count: TBD
     \vfill
@@
 
-% Just testing sections.. 
-\newpage
-\include{sections/Implementation}
-
 \end{document}

### 2021-02-19 math/ia_2021/model/test.py

@@ -24,11 +24,11 @@ def polyfit(x, y, degree):
 
 
 df = pd.read_csv("test.csv")
-power = 1
+power = 2
 
 
 print(polyfit(df['aaa'], df['test'], power))
 
-plt.plot(np.unique(df['aaa']), np.poly1d(np.polyfit(df['aaa'], df['test'], power))(np.unique(df['aaa'])))
+plt.plot(np.unique(df['aaa']), np.poly1d(np.polyfit(df[ 'aaa'], df['test'], power))(np.unique(df['aaa'])))
 plt.scatter(df['aaa'], df['test'], marker="x")
 plt.show()

### 2021-02-19 math/ia_2021/paper/paper.tex
info
math ia

@@
 
-\documentclass[12pt,a4paper]{article}
+\documentclass[11pt,letterpaper]{article}
 
@@
 \usepackage{packages}
+\addbibresource{sources.bib}
 
@@
 
-\include{sections/Introduction}
+\section{Introduction}
+
+\subsection{Background}
+
+Naturally occurring disasters such as tsunamis are devastating for 
+people and resources residing in risk areas. These tsunamis are caused 
+by many different types of natural phenomena such as landslides, collapse of 
+seamount, or more rarely, the impact of a meteorite. This paper however will focus 
+on one of the more common cases of tsunami; one caused by the earthquakes. This 
+is due to earthquakes causing around 72\% of earthquakes \cite{pacifictsunamimuseum}, and as the time between an 
+earthquake event and its tsunami is relatively high compared to events such as landslides \cite{sue_nokes_walters}, 
+leaving adequate time for prediction and correct preparation of the events properties, such as 
+the wave height, as discussed in this paper.
+
+\subsection{Personal Engagement}
+
+I have chosen to write about this subject as I have recently become interested by the 
+statistics and its implications in real-life matters. This research has proved very 
+interesting for me as it has abstracted the field of statistics for me. Furthermore, 
+as I enjoy programming, the research I have done has helped me learn about the tools 
+and technologies used by data analysts and other professionals of the matter.
+
+\section{}
+
+\printbibliography[heading=bibintoc, title=Works Cited]
 

### 2021-02-20 math/ia_2021/paper.tex
cleanup
update

--- 2021-02-19 math/ia_2021/paper/paper.tex
@@
 
+\section{Statistical Model}
+
+\subsection{Cleaning the dataset}
+
+The dataset, \cite{noaa}
+
+\subsection{Regression}
+
+\subsection{Ease of access of data}
+For an easy access to the predictions generated by the program, I implemented 
+a simple function to fetch new earthquake events from the \url{https://earthquake.usgs.gov/earthquakes/feed/v1.0/geojson.php} 
+live earthquake feed. These earthquake events are quantified into a magnitude \verb|mag| 
+and a distance \verb|dist| variable, which are supplied to the statistical model. 
+This information is then easily accessible from a human-readable report exported 
+to a plaintext file.
+
+\section{Uses of this research}
+
+The research included in this paper could be useful for many life-saving scenarios, 
+notably ones in which, for example, high quality and large-scale tsunami alert 
+systems are not available.
+
+\section{Conclusion}
+
+\subsection{Evaluation of the method used}
+
+This method, like any other statistical approximation, is by no means perfect 
+and for that reason should not be used in any real-world application. The results 
+procured by the program
+
 \printbibliography[heading=bibintoc, title=Works Cited]
 
+\appendix
+\section{Appendix}
+\label{app}
+\subsection{Python program}
+\label{app:scripts}
+qsd
+\lstinputlisting[language=Python]{model/multregr.py}
+
 \end{document}

### 2021-02-20 math/ia_2021/model/multregr.py

@@ -1,0 +1,83 @@
+import numpy as np
+import pandas as pd
+from matplotlib import pyplot as plt
+from mpl_toolkits.mplot3d import Axes3D
+from sklearn.preprocessing import PolynomialFeatures
+from sklearn.linear_model import LinearRegression
+
+
+def filter_time(df, threshold):
+    filter_df = df.loc[df.year > threshold].copy()
+    return filter_df
+
+
+def filter_distance(df, threshold):
+    filter_df = df.loc[df['distance from source (km)'] < threshold].copy()
+    return filter_df
+
+
+def filter_doubtful(df):
+    filter_df = df.loc[df['doubtful runup'] != 'y'].copy()
+    return filter_df
+
+
+def calculate_mean(x):
+    x['weight'] * x['maximum water height (m)']
+
+
+def weighted_mean_height(df):
+    return df.groupby('earthquake magnitude', as_index=True).apply(lambda x: (x['weight'] * x['maximum water height (m)']).mean())
+
+
+def get_droppable(df, whitelist):
+    to_drop = []
+    for (columnName, columnData) in df.iteritems():
+        if columnName not in whitelist:
+            to_drop.append(columnName)
+    return to_drop
+
+
+if __name__ == "__main__":
+    df = pd.read_table("dataset/runups.tsv", sep='\t')
+    df.columns = df.columns.str.lower()
+
+    # Filter data points before installation of
+    filter_df = filter_time(df, threshold=1960)
+
+    # Doubtful runups to be removed
+    filter_df = filter_doubtful(filter_df)
+
+    # Drop all columns except the ones needed for X1, X2 and the Y variables.
+    filter_df = filter_df.drop(
+        get_droppable(filter_df,
+                      ['earthquake magnitude',
+                       'distance from source (km)',
+                       'maximum water height (m)']), axis=1)
+
+    # Drop all rows with a NaN value  
+    filter_df = filter_df.dropna(how='any')
+
+    # Set X to the independent variables (equivalent to dropping the dependent)
+    X = filter_df.drop('maximum water height (m)', 1)
+
+    # Set y to our output column: maximum water height
+    y = filter_df['maximum water height (m)']
+
+    # The values to predict y with
+    pred = [[5.2, 225]]
+
+    # Fit the values 
+    poly = PolynomialFeatures(degree=2)
+    X_ = poly.fit_transform(X)
[…]

### 2021-02-21 math/ia_2021/paper.tex
python fixes, features, tex

@@
 
-\subsection{Cleaning the dataset}
+\subsection{Processing the dataset for use}
 
-The dataset, \cite{noaa}
+The dataset, procured from the NOAA website \cite{noaa}, includes many tsunami 
+runups, going back before the year 0. Due to this, the NOAA states uncertainties 
+that can occur in its data. For example, reports that were recorded before the 
+20th century tend to be based on written accounts, and not recorded seismic activity. 
+Furthermore, the installation of a system called the Worldwide Standardized Seismograph 
+Network in the 1960s allowed for more accurate data and for that reason any recorded 
+activity before this date will be discarded from the dataset.
+
+\dots
+
+The data must be correctly split into a training and an test dataset. The training 
+dataset will be used to generate a regression model while the test dataset will 
+be used to calculate the general accuracy of the model, by predicting values and 
+comparing those to the known value as found in the dataset. The dataset is split based on 
+a decimal number using the \verb|df.sample| method found in the Pandas library. A split 
+of 70\% for training and 30\% for test has been chosen. A \verb|random_state| argument is 
+set to an arbitrary value 42, as if it weren't, the dataset would be split in a different 
+manner at each execution of the program.
 
@@
 
-\subsection{Ease of access of data}
+Regression analysis is a very large portion of statistics, and regression models can 
+be found in numerous use cases of the field. Regression models consist of estimating the 
+relationship between one (or more) independent variables and a dependent variable. A 
+simple example of regression can be seen in the linear relationship between for example
+
+\dots
+
+The model's accuracy is then calculated, with the mentioned before "test" dataset. 
+The \verb|accuracy_score| method from the \verb|sklearn.metrics| class is used for this 
+task.
+
+\subsection{Ease of access}
 For an easy access to the predictions generated by the program, I implemented 
@@
 This information is then easily accessible from a human-readable report exported 
-to a plaintext file.
+to a plaintext file. Following is an example report generated by the tool:
 
@@
 and for that reason should not be used in any real-world application. The results 
-procured by the program
+procured by the program are subject to many
 

### 2021-02-21 math/ia_2021/model/multregr2.py

@@ -4,6 +4,10 @@ from matplotlib import pyplot as plt
 from mpl_toolkits.mplot3d import Axes3D
 from sklearn.preprocessing import PolynomialFeatures
 from sklearn.linear_model import LinearRegression
+from sklearn.model_selection import train_test_split
+from sklearn.metrics import r2_score
+
+
 
 
 def filter_time(df, threshold):
@@ -57,27 +61,37 @@ if __name__ == "__main__":
     # Drop all rows with a NaN value  
     filter_df = filter_df.dropna(how='any')
 
+    # Dataset split:
+    # The dataset is split into two parts, a "training" df_train dataset and 
+    # an "test" df_eval dataset. The "training" will be used for the 
+    # generation of a regression model and the "test" will 
+    # be used for the testing of this model, to calculate its accuracy. 
+    df_train, df_test = train_test_split(filter_df, test_size=0.3, random_state=42)
+
+    # For the train dataset, 
     # Set X to the independent variables (equivalent to dropping the dependent)
-    X = filter_df.drop('maximum water height (m)', 1)
-
     # Set y to our output column: maximum water height
-    y = filter_df['maximum water height (m)']
+    X_train = df_train.drop('maximum water height (m)', 1)
+    y_train = df_train['maximum water height (m)']
 
-    # The values to predict y with
-    pred = [[5.2, 225]]
+    # The test dataset, idem to the train dataset
+    X_test = df_test.drop('maximum water height (m)', 1)
+    y_test = df_test['maximum water height (m)']
 
-    # Fit the values 
+
+    # Fit the values to 
     poly = PolynomialFeatures(degree=2)
-    X_ = poly.fit_transform(X)
-    predict_ = poly.fit_transform(pred)
+    X_ = poly.fit_transform(X_train)
+    predict_ = poly.fit_transform(X_test)
 
     # Predict 
     mlr_model = LinearRegression()
-    mlr_model.fit(X_, y)
+    mlr_model.fit(X_, y_train)
     y_pred = mlr_model.predict(predict_)
-    print(y_pred)
+
+    print(r2_score(y_test, y_pred))
 
     # Print the coefficients and intercept
     theta0 = mlr_model.intercept_
     theta1, theta2, theta3, theta4, theta5, theta6 = mlr_model.coef_
-    print(theta0, theta1, theta2, theta3, theta4, theta5, theta6)
+    print("Intercept: ", theta0, "Coefficients: ", theta1, theta2, theta3, theta4, theta5, theta6)

### 2021-02-21 math/ia_2021/paper.tex
modif

@@
 relationship between one (or more) independent variables and a dependent variable. A 
-simple example of regression can be seen in the linear relationship between for example
+simple example of regression can be seen in the linear relationship between an independent 
+variable $x_{i}$, and the dependent variable $y$. The following expression shows this relationship in what is 
+called linear regression, as it models a straight line.
+
+$y = a + \beta x + e $
+
+Where the term $a$ represents the y-intercept of the line and $\beta$ represents the 
+coefficient of $x_i$. The final term $e$ is what is known as the error or residual 
+term, it is an estimate of the amount by which the predictions differ from the real 
+value.
+
+However, the relationship between the distance from an event and the wave height 
+is not necessarily linear, especially if it is considered that the travel of 
+sound in a fluid obeys the inverse square law. This already tells us a simple 
+linear regression model cannot be used for this research. If we want a second degree 
+polynomial regression formula, a new $x^2$ term can be added to the previous formula 
+and we get:
+
+$y = a + \beta_1 x + \beta_2 x^2 + e$
+
+Even though the right side of this equation is quadratic in $x$, this is still 
+linear regression as the right side is linear in $\beta$ (if we were to let 
+$x = 1$, we would be left with $y = a + \beta_1 + \beta_2 + e$, which is linear).
+
+In addition, our dataset has two independent variables: the distance from a particular 
+earthquake event and the magnitude of that earthquake.
+
+\dots
+
+So the relationship between the two independent variables and the wave height can 
+be estimated with a multivariate, polynomial regression model. We will use the Python 
+scikit-learn library to aid in creating this model.
 
@@
 task.
+
+\dots
 

### 2021-02-21 math/ia_2021/model/multregr2.py

@@ -66,7 +66,7 @@ if __name__ == "__main__":
     # an "test" df_eval dataset. The "training" will be used for the 
     # generation of a regression model and the "test" will 
     # be used for the testing of this model, to calculate its accuracy. 
-    df_train, df_test = train_test_split(filter_df, test_size=0.3, random_state=42)
+    df_train, df_test = train_test_split(filter_df, test_size=0.30, random_state=42)
 
     # For the train dataset, 
     # Set X to the independent variables (equivalent to dropping the dependent)
@@ -83,15 +83,17 @@ if __name__ == "__main__":
     poly = PolynomialFeatures(degree=2)
     X_ = poly.fit_transform(X_train)
     predict_ = poly.fit_transform(X_test)
+    #pre = poly.fit_transform([[9.0, 100]])
 
     # Predict 
     mlr_model = LinearRegression()
     mlr_model.fit(X_, y_train)
     y_pred = mlr_model.predict(predict_)
 
-    print(r2_score(y_test, y_pred))
+    print(y_test, y_pred)
+
+    print(mlr_model.score(predict_, y_test))
 
     # Print the coefficients and intercept
     theta0 = mlr_model.intercept_
-    theta1, theta2, theta3, theta4, theta5, theta6 = mlr_model.coef_
-    print("Intercept: ", theta0, "Coefficients: ", theta1, theta2, theta3, theta4, theta5, theta6)
+    print("Intercept: ", theta0, "Coefficients: ", mlr_model.coef_)

### 2021-02-21 math/ia_2021/paper.tex
wrote some text

@@
 Network in the 1960s allowed for more accurate data and for that reason any recorded 
-activity before this date will be discarded from the dataset.
+activity before this date will be discarded from the dataset. The dataset itself can 
+be accessed through two different methods; searching the database using the online 
+tool or downloading it in a somewhat obscure Google Earth format. The online tool 
+supports the download of search results in a much more human-readable \verb|.csv| format, 
+so I search for all entries after the year 1900, and downloaded the result.
 
-\dots
+The data is cleaned of all "dubious" reports, reports before the year 1960 (see above), and 
+of all irrelevant fields such as "Injuries" or "Damage \$Mil". This preliminary data cleanup 
+is done through the use of the pandas Python library \cite{reback2020pandas}\cite{mckinney-proc-scipy-2010},
 
-The data must be correctly split into a training and an test dataset. The training 
+The data must be then correctly split into a training and an test dataset. The training 
 dataset will be used to generate a regression model while the test dataset will 
@@
 comparing those to the known value as found in the dataset. The dataset is split based on 
-a decimal number using the \verb|df.sample| method found in the Pandas library. A split 
+a decimal number using the \verb|df.sample| method found in the pandas library. A split 
 of 70\% for training and 30\% for test has been chosen. A \verb|random_state| argument is 
@@
 $x = 1$, we would be left with $y = a + \beta_1 + \beta_2 + e$, which is linear). 
+A second degree polynomial regression would best fit our model for both independent 
+variables, as seen in the graphs below, where a theoretical regression line is drawn 
+by hand:
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\textwidth]{modelshowcase.jpeg}
+    \label{fig:boat1}
+\end{figure}
 
@@
 and for that reason should not be used in any real-world application. The results 
-procured by the program are subject to many
+procured by the program are subject to many uncertainties, one of which is described 
+below.
+
+When attempting to predict the maximum wave height of a tsunami such as the one 
+caused by the 2011 earthquake off the Pacific coast of Tōhoku, approximately 70km 
+away from shore, which produced a 38.9m wave \cite{yomiuri_2011}. The model in this 
+paper approximates this same event as only a 8m wave, when supplied the 9.0 magnitude 
+of the earthquake and the 70km distance. These types of errors can be attributed to the 
+fact that there are many more independent variables, such as the shape of the seafloor for 
+example, that can amplify or potentially change the height of a tsunami's wave. 
+%todo add
 

### 2021-02-21 math/ia_2021/model/multregr2.py

@@ -8,29 +8,14 @@ from sklearn.model_selection import train_test_split
 from sklearn.metrics import r2_score
 
 
-
-
 def filter_time(df, threshold):
     filter_df = df.loc[df.year > threshold].copy()
-    return filter_df
-
-
-def filter_distance(df, threshold):
-    filter_df = df.loc[df['distance from source (km)'] < threshold].copy()
     return filter_df
 
 
 def filter_doubtful(df):
     filter_df = df.loc[df['doubtful runup'] != 'y'].copy()
     return filter_df
-
-
-def calculate_mean(x):
-    x['weight'] * x['maximum water height (m)']
-
-
-def weighted_mean_height(df):
-    return df.groupby('earthquake magnitude', as_index=True).apply(lambda x: (x['weight'] * x['maximum water height (m)']).mean())
 
 
 def get_droppable(df, whitelist):
@@ -58,17 +43,26 @@ if __name__ == "__main__":
                        'distance from source (km)',
                        'maximum water height (m)']), axis=1)
 
-    # Drop all rows with a NaN value  
+    # Drop all rows with a NaN value
     filter_df = filter_df.dropna(how='any')
 
-    # Dataset split:
-    # The dataset is split into two parts, a "training" df_train dataset and 
-    # an "test" df_eval dataset. The "training" will be used for the 
-    # generation of a regression model and the "test" will 
-    # be used for the testing of this model, to calculate its accuracy. 
-    df_train, df_test = train_test_split(filter_df, test_size=0.30, random_state=42)
+    fig, (ax1, ax2) = plt.subplots(2)
+    ax1.scatter(filter_df['earthquake magnitude'], filter_df['maximum water height (m)'])
+    ax2.scatter(filter_df['distance from source (km)'], filter_df['maximum water height (m)'])
 
-    # For the train dataset, 
+    ax1.set(xlabel='earthquake magnitude', ylabel='maximum water height (m)')
+    ax2.set(xlabel='distance from source (km)', ylabel='maximum water height (m)')
+
+    plt.show()
+    # Dataset split:
+    # The dataset is split into two parts, a "training" df_train dataset and
+    # an "test" df_eval dataset. The "training" will be used for the
+    # generation of a regression model and the "test" will
+    # be used for the testing of this model, to calculate its accuracy.
+    df_train, df_test = train_test_split(
+        filter_df, test_size=0.30, random_state=42)
+
[…]

### 2021-02-24 math/ia_2021/model/multregr2.py
ia work

@@ -7,7 +7,6 @@ from sklearn.linear_model import LinearRegression
 from sklearn.model_selection import train_test_split
 from sklearn.metrics import r2_score
 
-
 def filter_time(df, threshold):
     filter_df = df.loc[df.year > threshold].copy()
     return filter_df
@@ -46,14 +45,14 @@ if __name__ == "__main__":
     # Drop all rows with a NaN value
     filter_df = filter_df.dropna(how='any')
 
+    """
     fig, (ax1, ax2) = plt.subplots(2)
     ax1.scatter(filter_df['earthquake magnitude'], filter_df['maximum water height (m)'])
     ax2.scatter(filter_df['distance from source (km)'], filter_df['maximum water height (m)'])
 
     ax1.set(xlabel='earthquake magnitude', ylabel='maximum water height (m)')
     ax2.set(xlabel='distance from source (km)', ylabel='maximum water height (m)')
-
-    plt.show()
+    """
     # Dataset split:
     # The dataset is split into two parts, a "training" df_train dataset and
     # an "test" df_eval dataset. The "training" will be used for the
@@ -73,10 +72,9 @@ if __name__ == "__main__":
     y_test = df_test['maximum water height (m)']
 
     # Fit the values to
-    poly = PolynomialFeatures(degree=-2)
+    poly = PolynomialFeatures(degree=2)
     X_ = poly.fit_transform(X_train)
     predict_ = poly.fit_transform(X_test)
-    #pre = poly.fit_transform([[9.0, 100]])
 
     # Predict
     mlr_model = LinearRegression()
@@ -84,9 +82,26 @@ if __name__ == "__main__":
     y_pred = mlr_model.predict(predict_)
 
     print(y_test, y_pred)
-
     print(mlr_model.score(predict_, y_test))
 
     # Print the coefficients and intercept
     theta0 = mlr_model.intercept_
     print("Intercept: ", theta0, "Coefficients: ", mlr_model.coef_)
+
+    # Code for generating 3d plot with prediction scatter
+    
+    fig = plt.figure()
+    plt.clf()
+    ax = Axes3D(fig)
+    ax = fig.gca(projection='3d')
+
+    ax.scatter(X_test['distance from source (km)'],
+               X_test['earthquake magnitude'], y_test)
+
+    ax.scatter(X_test['distance from source (km)'],
+               X_test['earthquake magnitude'], y_pred)
+
+    plt.show()
+    
+    
+    print(mlr_model.predict(poly.fit_transform([[4.3, 41]])))

### 2021-02-24 math/ia_2021/model/multregr2.py
added

@@ -45,14 +45,6 @@ if __name__ == "__main__":
     # Drop all rows with a NaN value
     filter_df = filter_df.dropna(how='any')
 
-    """
-    fig, (ax1, ax2) = plt.subplots(2)
-    ax1.scatter(filter_df['earthquake magnitude'], filter_df['maximum water height (m)'])
-    ax2.scatter(filter_df['distance from source (km)'], filter_df['maximum water height (m)'])
-
-    ax1.set(xlabel='earthquake magnitude', ylabel='maximum water height (m)')
-    ax2.set(xlabel='distance from source (km)', ylabel='maximum water height (m)')
-    """
     # Dataset split:
     # The dataset is split into two parts, a "training" df_train dataset and
     # an "test" df_eval dataset. The "training" will be used for the
@@ -89,7 +81,7 @@ if __name__ == "__main__":
     print("Intercept: ", theta0, "Coefficients: ", mlr_model.coef_)
 
     # Code for generating 3d plot with prediction scatter
-    
+    """
     fig = plt.figure()
     plt.clf()
     ax = Axes3D(fig)
@@ -102,6 +94,7 @@ if __name__ == "__main__":
                X_test['earthquake magnitude'], y_pred)
 
     plt.show()
-    
-    
+    """
+
+    #TODO: add GeoJSON report thingy https://earthquake.usgs.gov/earthquakes/feed/v1.0/summary/2.5_hour.geojson
     print(mlr_model.predict(poly.fit_transform([[4.3, 41]])))

### 2021-02-24 math/ia_2021/paper.tex
added
ia work

@@
 \setlength{\parindent}{0pt}
+
+\pagestyle{fancy}
+\lhead{Väinö-verneri Kauppila} % controls the left corner of the header
+\chead{} % controls the center of the header
+\rhead{} % controls the right corner of the header
+\lfoot{} % controls the left corner of the footer
+\cfoot{} % controls the center of the footer
+\rfoot{Page~\thepage} % controls the right corner of the footer
+\renewcommand{\headrulewidth}{0.4pt}
+\renewcommand{\footrulewidth}{0.4pt}
 
@@
         the magnitude of a preceding earthquake? \\
+        \vspace{1cm}
+        \textit{Väinö-verneri Kauppila} \\
+        February 2021 \\
         \vspace{4cm}
@@
 
-Naturally occurring disasters such as tsunamis are devastating for 
-people and resources residing in risk areas. These tsunamis are caused 
-by many different types of natural phenomena such as landslides, collapse of 
-seamount, or more rarely, the impact of a meteorite. This paper however will focus 
-on one of the more common cases of tsunami; one caused by the earthquakes. This 
-is due to earthquakes causing around 72\% of earthquakes \cite{pacifictsunamimuseum}, and as the time between an 
-earthquake event and its tsunami is relatively high compared to events such as landslides \cite{sue_nokes_walters}, 
-leaving adequate time for prediction and correct preparation of the events properties, such as 
-the wave height, as discussed in this paper.
+Naturally occurring disasters, such as tsunamis, are devastating for people, human settlements, 
+buildings and other resources residing in risk areas. Tsunamis are caused
+by many different types of natural phenomena, such as landslides, collapse of
+seamount, or more rarely, the impact of a meteorite. This paper, however, focuses
+on one of the more common causes of tsunamis; earthquakes. Earthquakes cause around 72\% 
+of tsunamis \cite{pacifictsunamimuseum}. The time between an earthquake and its tsunami is 
+relatively long compared to events such as landslides \cite{sue_nokes_walters},
+leaving adequate time to prepare for impact. Predicting wave height, as discussed in this 
+paper, can further help to assess the seriousness of the tsunami and provide help to 
+prepare for the event.
 
@@
 statistics and its implications in real-life matters. This research has proved very
-interesting for me as it has abstracted the field of statistics for me. Furthermore, 
-as I enjoy programming, the research I have done has helped me learn about the tools 
-and technologies used by data analysts and other professionals of the matter.
-
-\section{}
+interesting for me as it has made the field of statistics more concrete, because it provides 
+a real world use case. Furthermore, as I enjoy programming, the research I have done 
+has helped me learn about the tools and techniques used by data analysts and other professionals.
 
@@
 \subsection{Processing the dataset for use}
-
+%TODO: #2 Ask for help 
 The dataset, procured from the NOAA website \cite{noaa}, includes many tsunami
@@
 20th century tend to be based on written accounts, and not recorded seismic activity.
-Furthermore, the installation of a system called the Worldwide Standardized Seismograph 
-Network in the 1960s allowed for more accurate data and for that reason any recorded 
-activity before this date will be discarded from the dataset. The dataset itself can 
+Only since the 1960s, the installation of the Worldwide Standardized Seismograph
+Network system allowed for more accurate data in this paper. For that reason, any recorded
+activity before this date was discarded from the dataset for this paper. The dataset itself can
 be accessed through two different methods; searching the database using the online
-tool or downloading it in a somewhat obscure Google Earth format. The online tool 
-supports the download of search results in a much more human-readable \verb|.csv| format, 
-so I search for all entries after the year 1900, and downloaded the result.
+tool or downloading it in a somewhat obscure compressed Google Earth \verb|.kmz| format. 
+The online tool supports the download of search results in a much more human-readable 
+\verb|.csv| format. For this paper, all entries after the year 1900 were searched and downloaded.
 
@@
 
-The data must be then correctly split into a training and an test dataset. The training 
-dataset will be used to generate a regression model while the test dataset will 
-be used to calculate the general accuracy of the model, by predicting values and 
-comparing those to the known value as found in the dataset. The dataset is split based on 
-a decimal number using the \verb|df.sample| method found in the pandas library. A split 
-of 70\% for training and 30\% for test has been chosen. A \verb|random_state| argument is 
+The data was split into a training and a test dataset. The training
+dataset was used to generate a regression model while the test dataset was used 
+to calculate the general accuracy of the model, by predicting values and
+comparing those to the known value as found in the dataset. The dataset was split based on
+a fraction using the \verb|df.sample| method found in the pandas library. A split
+of 70\% for training and 30\% for test was chosen. A \verb|random_state| argument is
 set to an arbitrary value 42, as if it weren't, the dataset would be split in a different
@@
 
-Regression analysis is a very large portion of statistics, and regression models can 
-be found in numerous use cases of the field. Regression models consist of estimating the 
+Regression analysis is a commonly used tool in statistics, and regression models can
+be found in numerous use cases. Regression models consist of estimating the
 relationship between one (or more) independent variables and a dependent variable. A
@@
 
-$y = a + \beta x + e $
+$y = a + \beta x + e $,
 
-Where the term $a$ represents the y-intercept of the line and $\beta$ represents the 
+where the term $a$ represents the y-intercept of the line and $\beta$ represents the
 coefficient of $x_i$. The final term $e$ is what is known as the error or residual
-term, it is an estimate of the amount by which the predictions differ from the real 
+term. It is an estimate of the amount by which the predictions differ from the real
 value.
 
-However, the relationship between the distance from an event and the wave height 
-is not necessarily linear, especially if it is considered that the travel of 
-sound in a fluid obeys the inverse square law. This already tells us a simple 
-linear regression model cannot be used for this research. If we want a second degree 
-polynomial regression formula, a new $x^2$ term can be added to the previous formula 
-and we get:
+However, the relationship between the distance between an earthquake and its tsunami 
+and the wave height is not necessarily linear, especially if it is considered that the travel of
+sound (which is a wave, or the displacement of a fluid that can travel) in a fluid obeys the 
+inverse square law. Therefore, a simple linear regression model cannot be 
+used for this research. A second degree polynomial regression formula would be more 
+appropriate, and a new $x^2$ term can be added to the previous formula. We get:
 
@@
 A second degree polynomial regression would best fit our model for both independent
-variables, as seen in the graphs below, where a theoretical regression line is drawn 
+variables, as seen in the graphs below, where a theoretical curve of regression is drawn
 by hand:
@@
 In addition, our dataset has two independent variables: the distance from a particular
-earthquake event and the magnitude of that earthquake.
+earthquake event and the magnitude of that earthquake. Models such as the one explored 
+in this paper, where multiple independent variables are present, are called multivariate 
+and can be modeled by adding the terms of a new independent variable, as such 
+for a linear model with two independent variables:
 
-\dots
+$a_1 + \beta_1 x_1 + a_2 + \beta_2 x_2 + e$,
+
+where the terms $a_n$ are the y-intercepts of the $n$th independent variable,
 
@@
 
-\dots
-
-The model's accuracy is then calculated, with the mentioned before "test" dataset. 
-The \verb|accuracy_score| method from the \verb|sklearn.metrics| class is used for this 
-task.
+\begin{figure}[h]
+    \centering
+    \subfloat[\centering Linear model]{{\includegraphics[width=5cm]{linear.png} }}
+    \qquad
+    \subfloat[\centering Polynomial model]{{\includegraphics[width=5cm]{poly.png} }}
+    \caption{2 multivariate (hence the 3 dimensions) regression models}
+    \label{fig:example}
+\end{figure}
 
@@
 
-\subsection{Ease of access}
-For an easy access to the predictions generated by the program, I implemented 
-a simple function to fetch new earthquake events from the \url{https://earthquake.usgs.gov/earthquakes/feed/v1.0/geojson.php} 
-live earthquake feed. These earthquake events are quantified into a magnitude \verb|mag| 
-and a distance \verb|dist| variable, which are supplied to the statistical model. 
-This information is then easily accessible from a human-readable report exported 
-to a plaintext file. Following is an example report generated by the tool:
+The model's accuracy is then calculated, with the test dataset. At first, it was considered 
+to use the \verb|accuracy_score| method from the \verb|sklearn.metrics| class was used for 
+this task. This function calculates what is called the $R^2$ value, which is a measure used to 
+evaluate the fit a model on the dataset.
+
+\dots
 
@@
 example, that can amplify or potentially change the height of a tsunami's wave.
-%todo add
 
@@
 \label{app:scripts}
-qsd
-\lstinputlisting[language=Python]{model/multregr.py}
+
+Listed here is the Python program through which the dataset was cleaned and
+predictions based on the statistical model were made. Run using 64-bit Python
+version 3.9.1.
+
+\lstinputlisting[language=Python]{model/multregr2.py}
 

### 2021-02-24 math/ia_2021/model/multregr2.py

@@ -7,6 +7,7 @@ from sklearn.linear_model import LinearRegression
 from sklearn.model_selection import train_test_split
 from sklearn.metrics import r2_score
 
+
 def filter_time(df, threshold):
     filter_df = df.loc[df.year > threshold].copy()
     return filter_df
@@ -69,32 +70,36 @@ if __name__ == "__main__":
     predict_ = poly.fit_transform(X_test)
 
     # Predict
-    mlr_model = LinearRegression()
-    mlr_model.fit(X_, y_train)
-    y_pred = mlr_model.predict(predict_)
+    regressor = LinearRegression()
+    regressor.fit(X_, y_train)
+    y_pred = regressor.predict(predict_)
 
-    print(y_test, y_pred)
-    print(mlr_model.score(predict_, y_test))
+    #print(regressor.score(predict_, y_test))
 
     # Print the coefficients and intercept
-    theta0 = mlr_model.intercept_
-    print("Intercept: ", theta0, "Coefficients: ", mlr_model.coef_)
+    print('Coefficients: \n', regressor.coef_)
+    # The mean squared error
+    print("Mean squared error: %.2f" % np.mean((y_pred - y_test) ** 2))
+    # Explained variance score: 1 is perfect prediction
+    print('Variance score: %.2f' % regressor.score(predict_, y_test))
 
     # Code for generating 3d plot with prediction scatter
-    """
+    
     fig = plt.figure()
     plt.clf()
     ax = Axes3D(fig)
     ax = fig.gca(projection='3d')
+    ax.view_init(elev=15, azim=-122)
 
-    ax.scatter(X_test['distance from source (km)'],
-               X_test['earthquake magnitude'], y_test)
+    ax.set_title("2nd degree polynomial model")
+    ax.set_xlabel('distance from source (km)')
+    ax.set_ylabel('earthquake magnitude')
+    ax.set_zlabel('wave height (m)')
+
+    ax.scatter(X_test['distance from source (km)'].sample(frac=0.2, random_state=42),
+               X_test['earthquake magnitude'].sample(frac=0.2, random_state=42), y_test.sample(frac=0.2, random_state=42))
 
     ax.scatter(X_test['distance from source (km)'],
                X_test['earthquake magnitude'], y_pred)
 
     plt.show()
-    """
-
-    #TODO: add GeoJSON report thingy https://earthquake.usgs.gov/earthquakes/feed/v1.0/summary/2.5_hour.geojson
-    print(mlr_model.predict(poly.fit_transform([[4.3, 41]])))

### 2021-02-24 math/ia_2021/model/multregr2.py
ia work

@@ -83,8 +83,9 @@ if __name__ == "__main__":
     # Explained variance score: 1 is perfect prediction
     print('Variance score: %.2f' % regressor.score(predict_, y_test))
 
-    # Code for generating 3d plot with prediction scatter
-    
+    # Code for generating 3d plot with prediction scatter 
+    # only generates one graph, rerun with PolynomialFeatures(degree=n)
+    """
     fig = plt.figure()
     plt.clf()
     ax = Axes3D(fig)
@@ -103,3 +104,4 @@ if __name__ == "__main__":
                X_test['earthquake magnitude'], y_pred)
 
     plt.show()
+    """

### 2021-02-27 math/ia_2021/paper.tex
ia work

@@
 
-The data is cleaned of all "dubious" reports, reports before the year 1960 (see above), and
-of all irrelevant fields such as "Injuries" or "Damage \$Mil". This preliminary data cleanup
+The data is cleaned of data points marked as "dubious" (there is an column for this parameter), 
+reports before the year 1960 (see above), and of all irrelevant fields such as "Injuries" 
+or "Damage \$Mil". This preliminary data cleanup
 is done through the use of the pandas Python library \cite{reback2020pandas}\cite{mckinney-proc-scipy-2010}, 
+which is commonly used for the handling of large datasets for its efficiency.
 
@@
 variable $x_{i}$, and the dependent variable $y$. The following expression shows this relationship in what is
-called linear regression, as it models a straight line.
+called linear regression, as it models a straight line that can be used for prediction of 
+values:
 
-$y = a + \beta x + e $,
+$y =  \beta_0 a + \beta_1 x + e $,
 
-where the term $a$ represents the y-intercept of the line and $\beta$ represents the
-coefficient of $x_i$. The final term $e$ is what is known as the error or residual
+where the term $\beta_0$ represents the y-intercept of the line and $\beta$ represents the
+coefficient of $x$. The final term $e$ is what is known as the error or residual
 term. It is an estimate of the amount by which the predictions differ from the real
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{modelshowcase.jpeg}
+    \includegraphics[width=0.6\textwidth]{modelshowcase.jpeg}
     \label{fig:boat1}
@@
 
-The model's accuracy is then calculated, with the test dataset. At first, it was considered 
-to use the \verb|accuracy_score| method from the \verb|sklearn.metrics| class was used for 
-this task. This function calculates what is called the $R^2$ value, which is a measure used to 
-evaluate the fit a model on the dataset.
+The model's fit is then calculated, with the test dataset. The \verb|accuracy_score| method 
+from the \verb|sklearn.metrics| class was used for this task. This function calculates what 
+is called the $R^2$ value, which explains the extent at which the variance (the amount by which numbers 
+are spread from the average value) in the dependent variable can be explained by the independent 
+variable(s). An $R^2$ value of 0.6 indicates that 60\% of the model's results match those of the 
+dataset. This value is defined as:
 
-\dots
+$R^2 = 1 - \frac{SS_{res}}{SS_{total}}$,
 
-\section{Uses of this research}
+where $SS_{res}$ is the sum of the error terms, or the residual sum of squares, in a dataset of $n$ 
+values $y_1,\dots, y_i$, with their corresponding predicted values $p_1,\dots, p_i$ defined by
+
+$SS_{res} = \sum_{i} (y_i - p_i)^2 = \sum_{i} e_i^2$, \footnote{$\sum_i i$ would mean that all values i 
+are summed} where $e_i$ is the error term, equal to the difference of a data point $y_i$ 
+and its corresponding predicted value $p_i$.
+
+and $SS_{tot}$ is the total sum of squares, which is proportional to the variance in the data, 
+defined as:
+
+$SS_{tot} = \sum_{i} (y_i - \bar y)^2$, where $\bar y$ is the mean of the data $y_1,\dots, y_i$, 
+or $\bar y = \frac{1}{n} \sum_{i=1}^n y_i$
+
+\section{Conclusion}
 
@@
 systems are not available.
-
-\section{Conclusion}
 

### 2021-02-27 math/ia_2021/model/multregr.py

@@ -1,83 +1,0 @@
-import numpy as np
-import pandas as pd
-from matplotlib import pyplot as plt
-from mpl_toolkits.mplot3d import Axes3D
-from sklearn.preprocessing import PolynomialFeatures
-from sklearn.linear_model import LinearRegression
-
-
-def filter_time(df, threshold):
-    filter_df = df.loc[df.year > threshold].copy()
-    return filter_df
-
-
-def filter_distance(df, threshold):
-    filter_df = df.loc[df['distance from source (km)'] < threshold].copy()
-    return filter_df
-
-
-def filter_doubtful(df):
-    filter_df = df.loc[df['doubtful runup'] != 'y'].copy()
-    return filter_df
-
-
-def calculate_mean(x):
-    x['weight'] * x['maximum water height (m)']
-
-
-def weighted_mean_height(df):
-    return df.groupby('earthquake magnitude', as_index=True).apply(lambda x: (x['weight'] * x['maximum water height (m)']).mean())
-
-
-def get_droppable(df, whitelist):
-    to_drop = []
-    for (columnName, columnData) in df.iteritems():
-        if columnName not in whitelist:
-            to_drop.append(columnName)
-    return to_drop
-
-
-if __name__ == "__main__":
-    df = pd.read_table("dataset/runups.tsv", sep='\t')
-    df.columns = df.columns.str.lower()
-
-    # Filter data points before installation of WWSSN
-    filter_df = filter_time(df, threshold=1960)
-
-    # Doubtful runups to be removed
-    filter_df = filter_doubtful(filter_df)
-
-    # Drop all columns except the ones needed for X1, X2 and the Y variables.
-    filter_df = filter_df.drop(
-        get_droppable(filter_df,
-                      ['earthquake magnitude',
-                       'distance from source (km)',
-                       'maximum water height (m)']), axis=1)
-
-    # Drop all rows with a NaN value  
-    filter_df = filter_df.dropna(how='any')
-
-    # Set X to the independent variables (equivalent to dropping the dependent)
-    X = filter_df.drop('maximum water height (m)', 1)
-
-    # Set y to our output column: maximum water height
-    y = filter_df['maximum water height (m)']
-
-    # The values to predict y with
-    pred = [[5.2, 225]]
-
-    # Fit the values 
-    poly = PolynomialFeatures(degree=2)
-    X_ = poly.fit_transform(X)
[…]

### 2021-02-27 math/ia_2021/paper.tex
ia work

@@
 The online tool supports the download of search results in a much more human-readable 
-\verb|.csv| format. For this paper, all entries after the year 1900 were searched and downloaded.
+\verb|.tsv| (tab-separated) format. For this paper, all entries after the year 1900 were searched and downloaded.
 
@@
     \centering
-    \includegraphics[width=0.6\textwidth]{modelshowcase.jpeg}
+    \includegraphics[width=0.8\textwidth]{modelshowcase.jpeg}
     \label{fig:boat1}
@@
     \centering
-    \subfloat[\centering Linear model]{{\includegraphics[width=5cm]{linear.png} }}
+    \subfloat[\centering Linear model]{{\includegraphics[width=7.5cm]{linear.png} }}
     \qquad
-    \subfloat[\centering Polynomial model]{{\includegraphics[width=5cm]{poly.png} }}
+    \subfloat[\centering Polynomial model]{{\includegraphics[width=7.5cm]{poly.png} }}
     \caption{2 multivariate (hence the 3 dimensions) regression models}

### 2021-02-28 math/ia_2021/paper.tex
ia work

@@
 
-$SS_{res} = \sum_{i} (y_i - p_i)^2 = \sum_{i} e_i^2$, \footnote{$\sum_i i$ would mean that all values i 
-are summed} where $e_i$ is the error term, equal to the difference of a data point $y_i$ 
+$SS_{res} = \sum_{i} (y_i - p_i)^2 = \sum_{i} e_i^2$, \footnote{$\sum_{i=1}^n x_i$ means that all values of 
+x are summed, beginning from $x_1$ and ending at $x_n$} where $e_i$ is the error term, equal to the difference of a data point $y_i$ 
 and its corresponding predicted value $p_i$.

### 2021-03-02 math/ia_2021/paper.tex
ia work

--- 2021-02-27 math/ia_2021/paper.tex
@@
         \vspace{4cm}
-        Word count: TBD
+        Word count: TBD \\
+        \textit{Note: diagrams reserve a large amount of space, and cannot be placed in 
+        the appendix for the sake of clarity}
         \vfill
@@
 %TODO: #2 Ask for help 
-The dataset, procured from the NOAA website \cite{noaa}, includes many tsunami
+The dataset, procured from the National Oceanic and Atmospheric Administration website \cite{noaa}, includes many tsunami
 runups, going back before the year 0. Due to this, the NOAA states uncertainties
@@
 set to an arbitrary value 42, as if it weren't, the dataset would be split in a different
-manner at each execution of the program.
+manner at each execution of the program. The program can be found in Appendix \ref{app:scripts}.
 
 \subsection{Regression}
+
+\subsubsection{Linear model}
 
@@
 term. It is an estimate of the amount by which the predictions differ from the real
-value.
+value. An example of linear regression can be found in Appendix \ref{app:regression}, where 
+a set of data points is presented on a scatter plot, and the regression line is found 
+by finding the individual coefficients $\beta_n$ of $x$ (see equation above), like such:
+
+$\hat \beta_1 = \frac{\sum (x_i - \bar x) (y_i - \bar y)}{\sum (x_i - \bar x)^2}$, where 
+$\hat \beta_n$ represents an estimated value and $\bar x$ represents the mean of the set $x$
+
+and the $\hat \beta_0$ value can be approximated as $\hat \beta_0 = \bar y - \hat \beta_1 \bar x$
+
+For the dataset in \ref{app:regression}, the results to these equations are:
+
+$\hat \beta_1 = \frac{1806757.14}{322280.86} = 5.60$ and $\hat \beta_0 = 2257.14 - 5.61 \cdot 512.86 = -618.01$
+
+So the regression line can be expressed as $\hat y = 5.60 x -618.01$
+
+\subsubsection{Special cases}
 
@@
 Even though the right side of this equation is quadratic in $x$, this is still
-linear regression as the right side is linear in $\beta$ (if we were to let
+linear regression as the right side is linear in the $\beta$ terms (if we were to let
 $x = 1$, we would be left with $y = a + \beta_1 + \beta_2 + e$, which is linear).
@@
 
-$SS_{res} = \sum_{i} (y_i - p_i)^2 = \sum_{i} e_i^2$, \footnote{$\sum_i i$ would mean that all values i 
-are summed} where $e_i$ is the error term, equal to the difference of a data point $y_i$ 
+$SS_{res} = \sum_{i} e_i^2 = \sum_{i} (y_i - p_i)^2$, \footnote{$\sum_{i=1}^n x_i$ means that all values of
+    x are summed, beginning from $x_1$ and ending at $x_n$} where $e_i$ is the error term, equal to the difference of a data point $y_i$
 and its corresponding predicted value $p_i$.
@@
 \label{app}
+\subsection{Example linear regression}
+\label{app:regression}
+
+Below is an example of a dataset with $x$ values and their corresponding
+Y values, the data points chosen are purely arbitrary.
+
+\begin{center}
+    \begin{tabular}{||c c||}
+        \hline
+        X values & Y values \\ [0.5ex]
+        \hline\hline
+        368      & 1700     \\
+        \hline
+        340      & 1500     \\
+        \hline
+        665      & 2800     \\
+        \hline
+        954      & 5000     \\
+        \hline
+        331      & 1300     \\
+        \hline
+        556      & 2200     \\
+        \hline
+        376      & 1300     \\ [1ex]
+        \hline
+    \end{tabular}
+\end{center}
+
+\pgfplotstableread{
+    X Y
+    368 1700
+    340 1500
+    665 2800
+    954 5000
+    331 1300
+    556 2200
+    376 1300
+}\table
+
+Below is the dataset plotted, along with a line of regression.
+
+\begin{tikzpicture}[scale=1.1]
+    \begin{axis}[
+            title={Comparing X and Y},
+            xlabel={X axis},
+            ylabel={Y axis},
+            xmin=300, xmax=1000,
+            ymin=0, ymax=5500,
+            ytick={0,1000,2000,3000,4000,5000},
+            xtick={0,400,500,600,700,800,900},
+            legend pos=north west,
+            ymajorgrids=true,
+            grid style=dashed,
+            scale=1.2
+        ]
+
+        \addplot[
+            only marks,
+            color=blue,
+            mark=*,
+        ] table {\table};
+        \legend{Datapoint}
+
+        \addplot [thin, blue] table[
+                y={create col/linear regression={y=Y}}
+            ] % compute a linear regression from the input table
+            {\table};
+        \addlegendentry{
+            Linear regression: $ x =
+                \pgfmathprintnumber{\pgfplotstableregressiona}
+                \cdot t
+                \pgfmathprintnumber[print sign]{\pgfplotstableregressionb}$
+        };
+
+    \end{axis}
+\end{tikzpicture}
+
 \subsection{Python program}

### 2021-03-02 math/ia_2021/paper.tex
mu-update
ia work

@@
 \begin{document}
-\doublespacing % Double spacing throughout the document
+%\doublespacing % Double spacing throughout the document
 
@@
         \vspace{4cm}
-        Word count: TBD \\
-        \textit{Note: diagrams reserve a large amount of space, and cannot be placed in 
-        the appendix for the sake of clarity}
+        Word count: TBD
         \vfill
@@
 
-So the regression line can be expressed as $\hat y = 5.60 x -618.01$
+So the regression line can be expressed as $\boxed{\hat y = 5.60 x -618.01}$
 
@@
 
+\section*{Notes}
+\label{sec:notes}
+\addcontentsline{toc}{section}{\nameref{sec:notes}}
+
+All figures are generated through the use of TikZ (Tantau) and the Python library 
+\verb|matplotlib|. This paper is written with the aid of the typesetting software \LaTeX. 
+On a PDF viewer, the sections in the table of contents can be clicked to access that 
+section.
+
 \printbibliography[heading=bibintoc, title=Works Cited]
@@
 
+\begin{center}
     \begin{tikzpicture}[scale=1.1]
@@
     \end{tikzpicture}
+\end{center}
 

### 2021-03-02 math/ia_2021/paper.tex
ia work, submitted rough draft

@@
         \vspace{4cm}
-        Word count: TBD
+        Word count: TBD \\
+        \textit{First draft, details are missing}
         \vfill
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
     \centering
@@
 be estimated with a multivariate, polynomial regression model. We will use the Python
-scikit-learn library to aid in creating this model.
+scikit-learn library to aid in creating this model. \\
 
-\begin{figure}[h]
+(\textit{Section not done, need to do more research on how the} \verb|PolynomialFeatures| 
+\textit{function transforms a linear array to one fit for polynomial regression})
+
+\begin{figure}[H]
     \centering
@@
 
-\dots
+\subsubsection{Calculating fit of the model}
 
@@
 
+\textit{Conclusion and evaluation of the method are contradictory, to be corrected}
+
 \subsection{Evaluation of the method used}
@@
 
+\newpage
+\pagenumbering{roman}
+
 \section*{Notes}
@@
 On a PDF viewer, the sections in the table of contents can be clicked to access that 
-section.
+section. Dataset not included due to large size (\textasciitilde 28000 columns).
 

### 2021-03-06 research project/sections/Background.tex
rp work

@@
 the lengths $OH$ and subsequently $BD$ in (Figure 2) \ref{fig:image1}.
+
+\begin{figure}[H]
+  \definecolor{qqwuqq}{rgb}{0,0.39215686274509803,0}
+  \definecolor{aazzee}{rgb}{0.898039215686, 0.364705882353, 0.294117647059}
+  \definecolor{uuuuuu}{rgb}{0.26666666666666666,0.26666666666666666,0.26666666666666666}
+  \definecolor{xdxdff}{rgb}{0.49019607843137253,0.49019607843137253,1}
+  \definecolor{qqqqzz}{rgb}{0,0,0.6}
+  \begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=1cm,y=1cm,scale=0.7]
+  \clip(-8.372520930232552,-8.50470697674418) rectangle (7.69259534883721,7.028316279069767);
+  \draw [shift={(-0.03355999206952884,0)},line width=1pt,color=qqwuqq,fill=qqwuqq,fill opacity=0.10000000149011612] (0,0) -- (0:1.3302325581395344) arc (0:44.77428433325932:1.3302325581395344) -- cycle;
+  \draw [line width=0.8pt,color=qqqqzz] (0,0) circle (6cm);
+  \draw [line width=1pt, color=aazzee] (4.242640687119286,4.242640687119285)-- (6,0);
+  \draw [line width=1pt, color=aazzee] (6,0)-- (4.242640687119286,-4.242640687119285);
+  \draw [line width=1pt, color=qqwuqq] (4.242640687119286,-4.242640687119285)-- (4.242640687119286,4.242640687119285);
+  \draw [line width=1pt] (-6,0)-- (6,0);
+  \draw [line width=1pt] (-4.275936263984878,-4.209081736713964)-- (4.242640687119286,4.242640687119285);
+  \draw [line width=1pt] (-4.242640687119285,4.242640687119286)-- (4.242640687119286,-4.242640687119285);
+  \draw [line width=1pt] (-0.03355999206952884,0)-- (5.121320343559643,2.1213203435596424);
+  \draw [line width=1pt, color=qqwuqq] (-4.242640687119285,4.242640687119286)-- (-4.275936263984878,-4.209081736713964);
+
+  \draw [line width=1pt, dotted, color=aazzee] (0,6)-- (4.242640687119286,4.242640687119285);
+  \draw [line width=1pt, dotted, color=aazzee] (0,6)-- (-4.242640687119286,4.242640687119285);
+
+  \draw [line width=1pt, dotted, color=aazzee] (-6,0)-- (-4.242640687119286,4.242640687119285);
+  \draw [line width=1pt, dotted, color=aazzee] (-6,0)-- (-4.242640687119286,-4.242640687119285);
+
+  \draw [line width=1pt, dotted, color=aazzee] (0,-6)-- (-4.242640687119286,-4.242640687119285);
+  \draw [line width=1pt, dotted, color=aazzee] (0,-6)-- (4.242640687119286,-4.242640687119285);
+
+  \draw [line width=1pt, dotted, color=qqwuqq] (-4.242640687119286,-4.242640687119285)-- (4.242640687119286,-4.242640687119285); 
+  \draw [line width=1pt, dotted, color=qqwuqq] (-4.242640687119286,4.242640687119285)-- (4.242640687119286,4.242640687119285);
+
+  \begin{scriptsize}
+  \draw [fill=xdxdff] (4.242640687119286,4.242640687119285) circle (2.5pt);
+  \draw[color=xdxdff] (4.4772465116279085,4.623665116279071) node {$B$};
+  \draw [fill=xdxdff] (-4.242640687119285,4.242640687119286) circle (2.5pt);
+  \draw[color=xdxdff] (-4.295311627906972,4.623665116279071) node {$A$};
+  \draw [fill=xdxdff] (-4.275936263984878,-4.209081736713964) circle (2.5pt);
+  \draw[color=xdxdff] (-4.336241860465112,-4.55) node {$C$};
+  \draw [fill=xdxdff] (6,0) circle (2.5pt);
+  \draw[color=xdxdff] (6.237246511627907,0.3873860465116302) node {$E$};
+  \draw [fill=xdxdff] (4.242640687119286,-4.242640687119285) circle (2.5pt);
+  \draw[color=xdxdff] (4.3772465116279085,-4.5) node {$D$};
+  \draw[color=black] (5.3595720930232575,2.3111069767441874) node {$H'$};
+  \draw [fill=xdxdff] (5.12,2.12) circle (2.5pt);
+  \draw[color=black] (3.947479069767444,-0.615404651162788) node {$H$};
+  \draw [fill=xdxdff] (4.242640687119286,0) circle (2.5pt);
+  \draw [fill=xdxdff] (0,0) circle (2.5pt);
+  \draw[color=uuuuuu] (0.00003720930232879,-0.3464558139534907) node {$O$};
+  \draw[color=qqwuqq] (1.5325953488372117,0.3873860465116302) node {$\alpha$};
+  \end{scriptsize}
+  \end{tikzpicture}
+  \label{fig:image1}
+  \caption[]{Circle with 1 segment from a $n$-gon with point $H$ and 2 
+  segments from an $2n$-gon, one of which on point $H'$, inscribed in 
+  a circle of radius $OB$, adapted from (Boris Gourévitch) \cite{borisgourevitch2013}}
+  \end{figure}
+
 % $$\pi = \frac{2}{\prod_{k = 0}^{\infty} U_{k}}$$    symbolic
@@
 
-        \draw[xshift=-1cm, dotted] (1,-0.7) node[circle,fill,inner sep=1pt](a){} -- (5,-0.7) 
-node[circle,fill,inner sep=1pt,label=right:$+\frac{4}{1}$](b){};
+        \draw[xshift=-1cm] (1,-0.7) node[circle,fill,inner sep=1pt](a){} -- (5,-0.7) 
+node[circle,fill,inner sep=1pt,label=right:$+\frac{4}{1} \boxed{\text{ iter. 1}}$](b){};
 
-\draw[xshift=-1cm, dotted] (3.67,-1.1) node[circle,fill,inner sep=1pt, label=left:$-\frac{4}{3}$](c){} -- (5,-1.1) 
+\draw[xshift=-1cm] (3.67,-1.1) node[circle,fill,inner sep=1pt, label=left:$\boxed{\text{iter. 2 }} -\frac{4}{3}$](c){} -- (5,-1.1) 
 node[circle,fill,inner sep=1pt](d){};
 
-\draw[xshift=-1cm, dotted] (3.67,-1.5) node[circle,fill,inner sep=1pt](e){} -- (4.46,-1.5) 
-node[circle,fill,inner sep=1pt, label=right:$+\frac{4}{5}$](f){};
+\draw[xshift=-1cm] (3.67,-1.5) node[circle,fill,inner sep=1pt](e){} -- (4.46,-1.5) 
+node[circle,fill,inner sep=1pt, label=right:$+\frac{4}{5} \boxed{\text{ iter. 3}}$](f){};
 
-\draw[xshift=-1cm, dotted] (3.89,-1.9) node[circle,fill,inner sep=1pt, label=left:$-\frac{4}{7}$](g){} -- (4.46,-1.9) 
+\draw[xshift=-1cm] (3.89,-1.9) node[circle,fill,inner sep=1pt, label=left:$\boxed{\text{iter. 4 }} -\frac{4}{7}$](g){} -- (4.46,-1.9) 
 node[circle,fill,inner sep=1pt](h){};
 
-\draw[xshift=-1cm, dotted] (3.89,-2.3) node[circle,fill,inner sep=1pt](i){} -- (4.33,-2.3) 
-node[circle,fill,inner sep=1pt, label=right:$+\frac{4}{9}$](j){};
+\draw[xshift=-1cm] (3.89,-2.3) node[circle,fill,inner sep=1pt](i){} -- (4.33,-2.3) 
+node[circle,fill,inner sep=1pt, label=right:$+\frac{4}{9} \boxed{\text{ iter. 5}}$](j){};
 
-\draw[->, to path={-| (\tikztotarget)}, dotted]
+\draw[->, to path={-| (\tikztotarget)}]
   (b) edge (d) (c) edge (e) (f) edge (h) (g) edge (i);
@@
 \end{figure}
-
-\begin{figure}[!ht]
-\definecolor{qqwuqq}{rgb}{0,0.39215686274509803,0}
-\definecolor{uuuuuu}{rgb}{0.26666666666666666,0.26666666666666666,0.26666666666666666}
-\definecolor{xdxdff}{rgb}{0.49019607843137253,0.49019607843137253,1}
-\definecolor{qqqqzz}{rgb}{0,0,0.6}
-\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=1cm,y=1cm,scale=0.7]
-\clip(-8.372520930232552,-8.50470697674418) rectangle (7.69259534883721,7.028316279069767);
-\draw [shift={(-0.03355999206952884,0)},line width=1pt,color=qqwuqq,fill=qqwuqq,fill opacity=0.10000000149011612] (0,0) -- (0:1.3302325581395344) arc (0:44.77428433325932:1.3302325581395344) -- cycle;
-\draw [line width=0.8pt,color=qqqqzz] (0,0) circle (6cm);
-\draw [line width=1pt] (4.242640687119286,4.242640687119285)-- (6,0);
-\draw [line width=1pt] (6,0)-- (4.242640687119286,-4.242640687119285);
-\draw [line width=1pt] (4.242640687119286,-4.242640687119285)-- (4.242640687119286,4.242640687119285);
-\draw [line width=1pt] (-6,0)-- (6,0);
-\draw [line width=1pt] (-4.275936263984878,-4.209081736713964)-- (4.242640687119286,4.242640687119285);
-\draw [line width=1pt] (-4.242640687119285,4.242640687119286)-- (4.242640687119286,-4.242640687119285);
-\draw [line width=1pt] (-0.03355999206952884,0)-- (5.121320343559643,2.1213203435596424);
-\draw [line width=1pt] (-4.242640687119285,4.242640687119286)-- (-4.275936263984878,-4.209081736713964);
-\begin{scriptsize}
-\draw [fill=xdxdff] (4.242640687119286,4.242640687119285) circle (2.5pt);
-\draw[color=xdxdff] (4.4772465116279085,4.623665116279071) node {$B$};
-\draw [fill=xdxdff] (-4.242640687119285,4.242640687119286) circle (2.5pt);
-\draw[color=xdxdff] (-4.295311627906972,4.623665116279071) node {$A$};
-\draw [fill=xdxdff] (-4.275936263984878,-4.209081736713964) circle (2.5pt);
-\draw[color=xdxdff] (-4.336241860465112,-4.55) node {$C$};
-\draw [fill=xdxdff] (6,0) circle (2.5pt);
-\draw[color=xdxdff] (6.237246511627907,0.3873860465116302) node {$E$};
-\draw [fill=xdxdff] (4.242640687119286,-4.242640687119285) circle (2.5pt);
-\draw[color=xdxdff] (4.3772465116279085,-4.5) node {$D$};
-\draw[color=black] (5.3595720930232575,2.3111069767441874) node {$H'$};
-\draw [fill=xdxdff] (5.12,2.12) circle (2.5pt);
-\draw[color=black] (3.947479069767444,-0.615404651162788) node {$H$};
-\draw [fill=xdxdff] (4.242640687119286,0) circle (2.5pt);
-\draw [fill=xdxdff] (0,0) circle (2.5pt);
-\draw[color=uuuuuu] (0.00003720930232879,-0.3464558139534907) node {$O$};
-\draw[color=qqwuqq] (1.5325953488372117,0.3873860465116302) node {$\alpha$};
-\end{scriptsize}
-\end{tikzpicture}
-\label{fig:image1}
-\caption[]{Circle with 1 segment from a $n$-gon with point $H$ and 2 
-segments from an $2n$-gon, one of which on point $H'$, inscribed in 
-a circle of radius $OB$, adapted from (Boris Gourévitch) \cite{borisgourevitch2013}}
-\end{figure}

### 2021-03-07 math/ia_2021/paper.tex
final rp draft

@@
 
-$SS_{res} = \sum_{i} e_i^2 = \sum_{i} (y_i - p_i)^2$, \footnote{$\sum_{i=1}^n x_i$ means that all values of
+$SS_{res} = \sum\limits_{i} e_i^2 = \sum\limits_{i} (y_i - p_i)^2$, \footnote{$\sum\limits_{i=1}^n x_i$ means that all values of
     x are summed, beginning from $x_1$ and ending at $x_n$} where $e_i$ is the error term, equal to the difference of a data point $y_i$
@@
 
-$SS_{tot} = \sum_{i} (y_i - \bar y)^2$, where $\bar y$ is the mean of the data $y_1,\dots, y_i$,
-or $\bar y = \frac{1}{n} \sum_{i=1}^n y_i$
+$SS_{tot} = \sum\limits_{i} (y_i - \bar y)^2$, where $\bar y$ is the mean of the data $y_1,\dots, y_i$,
+or $\bar y = \frac{1}{n} \sum\limits_{i=1}^n y_i$
 

### 2021-03-07 research project/sections/Analysis_of_results.tex
rp work
final rp draft

--- 2021-02-02 research project/sections/Analysis_of_results.tex
@@
     X Y
-    5 0.7601666451
-    10 1.288158894
-    15 2.050116062 
-    20 2.63982296 
-    25 3.091802597
-    30 3.819538355 
-    35 4.442822933
-    40 5.081979036 
-    45 5.706284046 
-    50 6.370962858 
-    55 7.020516396 
-    60 7.723451853   
+    0.7601666451 5
+    1.288158894 10
+    2.050116062 15
+    2.63982296 20
+    3.091802597 25
+    3.819538355 30
+    4.442822933 35
+    5.081979036 40
+    5.706284046 45
+    6.370962858 50
+    7.020516396 55
+    7.723451853 60
 }\vietetable
@@
     X Y
-    5 0.5919837952 
-    10 1.307063103  
-    15 1.902475357  
-    20 2.679748535  
-    25 3.212277889  
-    30 3.904948235  
-    35 4.547913074  
-    40 5.255579948  
-    45 5.837199688  
-    50 6.578524113  
-    55 7.270073891  
-    60 8.004565239
+    0.5919837952 5
+    1.307063103 10
+    1.902475357 15
+    2.679748535 20
+    3.212277889 25
+    3.904948235 30
+    4.547913074 35
+    5.255579948 40
+    5.837199688 45
+    6.578524113 50
+    7.270073891 55
+    8.004565239 60
 
@@
     X Y
-    5 10
-    10 17
-    15 27
-    20 35
-    25 42
-    30 51
-    35 59
-    40 67
-    45 76
-    50 84
-    55 92
-    60 101
+    10  5
+    17  10
+    27  15
+    35  20
+    42  25
+    51  30
+    59  35
+    67  40
+    76  45
+    84  50
+    92  55
+    101 60
 }\vietetableiter
@@
     X Y
-    5 9
-    10 20
-    15 29
-    20 41
-    25 50
-    30 61
-    35 71
-    40 82
-    45 91
-    50 102
-    55 112
-    60 123
+    9 5
+    20 10
+    29 15
+    41 20
+    50 25
+    61 30
+    71 35
+    82 40
+    91 45
+    102 50
+    112 55
+    123 60
 }\madhavatableiter
@@
 
-The arithmetic mean of a decimal place was caluclated for 100 trials and the data points 
+The arithmetic mean of a decimal place was calculated for 100 trials and the data points
 in the table shown below were found (rounded to 3 significant figures). The values for the decimal place
@@
         \toprule
-    Decimal place & Iterations needed, Viète's method (no unit) & Iterations needed, Madhava's method (no unit)\\
+        Decimal place & Iterations needed, Viète's method (no unit) & Iterations needed, Madhava's method (no unit) \\
         \midrule
@@
             title={Comparing the iterations needed for approximating the value of $\pi$},
-        xlabel={Decimal places},
-        ylabel={Iterations},
-        xmin=0, xmax=60,
-        ymin=0, ymax=130,
-        ytick={0,10,20,30,40,50,60,70,80,90,100,110,120,130},
-        xtick={0,10,20,30,40,50,60},
-        legend pos=north west,
+            xlabel={Iterations},
+            ylabel={Decimal places},
+            xmin=0, xmax=130,
+            ymin=0, ymax=60,
+            ytick={0,10,20,30,40,50,60},
+            xtick={0,10,20,30,40,50,60,70,80,90,100,110,120,130},
+            legend pos=south east,
             ymajorgrids=true,
@@
             title={Comparing the time taken for approximating the value of $\pi$},
-        xlabel={Decimal places},
-        ylabel={Time taken [ms]},
-        xmin=0, xmax=60,
-        ymin=0, ymax=10,
-        ytick={0,2,4,6,8,10},
-        xtick={0,10,20,30,40,50,60},
-        legend pos=north west,
+            xlabel={Time taken [ms]},
+            ylabel={Decimal places},
+            xmin=0, xmax=10,
+            ymin=0, ymax=60,
+            ytick={0,10,20,30,40,50,60},
+            xtick={0,2,4,6,8,10},
+            legend pos=south east,
             ymajorgrids=true,
@@
 the value $\pi$. However, in the second graph comparing time, it can be noted that Madhava's
-method is marginally faster until the decimal 20, when the two methods diverge. It could be 
-assumed that this occurrence is due to differences between the types of the two methods. Madhava's 
+method is marginally faster until the decimal 20, when the lines of the two methods stray apart. It could be
+assumed that this occurrence is due to differences between the types of the two methods. Although it 
+could be argued that this difference is only due to random error, the results in this graph 
+are those collected from the arithmetic mean over 100 trials, and so it can be assessed that this is not the case. Madhava's
 method is what is called an alternating series, due to its nature of alternating between values in order
@@
 deduced that at lower decimal values Madhava's method is faster as Viète's method's approximation
-accuracy increases in what looks like an exponential fashion.
-
-\textit{More info to be added}
+accuracy firstly increases in what seems similar to a curve from an exponential function $f(x) = k^x$.
+It can also be said that while Madhava's method alternates between values such that $m_{n} > \pi, m_{n+1} < \pi$,
+Viète's method $\pi_v$ approximates pi in a manner such that $\pi_v < \pi$.
 

### 2021-03-07 research project/sections/Background.tex
rp work
final rp draft

@@
 Which has come to be known as the Leibniz formula for $\pi$, after the German mathematician
-who discovered the same formula two decades later. \cite{edwards_1994} This series converges to $\pi$, as seen in (Figure 1) \ref{fig:numberline}
+who discovered the same formula two decades later. \cite{edwards_1994} This series converges 
+to $\pi$, as seen in (Figure 1) \ref{fig:numberline}
+
+\begin{figure}[H]
+
+  \begin{tikzpicture}
+    \draw[latex-latex] (-0.5,0) -- (4.5,0);% the x-axis
+    \foreach \x in  {0,...,4} % tick marks
+    \draw (\x,0) -- (\x,-3pt);
+    \foreach \x in {0,...,4} % the numbers
+    \node [below] at (\x,-0.1) {$\x$};
+    \node [below] at (3.14,0.5) {$\pi$};
+    \draw (3.14,0) -- (3.14,3pt);
+
+    \draw[->, dotted] (3.14,0) -- (3.14,-3);% the x-axis
+
+    \draw[xshift=-1cm] (1,-0.7) node[circle,fill,inner sep=1pt,label=left:$\boxed{\text{iter. 0}}$](a){} -- (5,-0.7)
+    node[circle,fill,inner sep=1pt,label=right:$+\frac{4}{1} \boxed{\text{ iter. 1}}$](b){};
+
+    \draw[xshift=-1cm] (3.67,-1.1) node[circle,fill,inner sep=1pt, label=left:$\boxed{\text{iter. 2 }} -\frac{4}{3}$](c){} -- (5,-1.1)
+    node[circle,fill,inner sep=1pt](d){};
+
+    \draw[xshift=-1cm] (3.67,-1.5) node[circle,fill,inner sep=1pt](e){} -- (4.46,-1.5)
+    node[circle,fill,inner sep=1pt, label=right:$+\frac{4}{5} \boxed{\text{ iter. 3}}$](f){};
+
+    \draw[xshift=-1cm] (3.89,-1.9) node[circle,fill,inner sep=1pt, label=left:$\boxed{\text{iter. 4 }} -\frac{4}{7}$](g){} -- (4.46,-1.9)
+    node[circle,fill,inner sep=1pt](h){};
+
+    \draw[xshift=-1cm] (3.89,-2.3) node[circle,fill,inner sep=1pt](i){} -- (4.33,-2.3)
+    node[circle,fill,inner sep=1pt, label=right:$+\frac{4}{9} \boxed{\text{ iter. 5}}$](j){};
+
+    \draw[->, to path={-| (\tikztotarget)}]
+    (b) edge (d) (c) edge (e) (f) edge (h) (g) edge (i);
+  \end{tikzpicture}
+
+  \label{fig:numberline}
+  \caption[]{A number line demonstrating the convergence of Madhava's first method for the
+    calculation of $\pi$. Adapted from (V. M. Jamkar) \cite{jamkar_2018}}
+\end{figure}
 
@@
   \label{fig:image1}
-  \caption[]{Circle with 1 segment from a $n$-gon with point $H$ and 2 
-  segments from an $2n$-gon, one of which on point $H'$, inscribed in 
+  \caption[]{Circle with 1 segment from a $n$-gon (shown by the \textcolor{green}{green} dotted line) with point $H$ and 2
+    segments from an $2n$-gon (shown by the \textcolor{red}{red} dotted line), one of which on point $H'$, inscribed in
     a circle of radius $OB$, adapted from (Boris Gourévitch) \cite{borisgourevitch2013}}
@@
 % $$\pi = \frac{2}{\prod_{k = 0}^{\infty} U_{k}}$$    symbolic
-With the radius of the circle with center $R = OB$,
+With the radius $R$ of the circle with center $O$ defined as $R = OB$,
 
@@
 Since the equation for the area of a polygon is defined as $A = \frac{p \cdot a}{2}$,
-where $p$ is the perimeter of the polygon and $a$ is the apothem, in this case $BD \cdot n$ 
+where $p$ is the perimeter of the polygon and $a$ is the apothem\footnote{apothem: the segment 
+from the center of a polygon to the midpoint of one of its sides, here $OH$ for the $n$-gon and 
+$OH'$ for the $2n$-gon}, in this case $BD \cdot n$
 and $OH$ respectively, let $A_{n}$ equal the area of
-the polygon with $n$ sides such that: \\
+the polygon with $n$ sides (displayed in \textcolor{green}{green} in Figure 2) such that: \\
 $A_{n} = \frac{OH \cdot BD \cdot n}{2}$ \\
@@
 
-\footnotetext{where $\prod$ signifies a product. Similar expression to $\sum$ }
-
-\begin{figure}[h]
-
-    \begin{tikzpicture}
-        \draw[latex-latex] (-0.5,0) -- (4.5,0);% the x-axis
-        \foreach \x in  {0,...,4} % tick marks
-          \draw (\x,0) -- (\x,-3pt);
-        \foreach \x in {0,...,4} % the numbers
-          \node [below] at (\x,-0.1) {$\x$};
-        \node [below] at (3.14,0.5) {$\pi$};
-        \draw (3.14,0) -- (3.14,3pt);
-
-        \draw[->, dotted] (3.14,0) -- (3.14,-3);% the x-axis
-
-        \draw[xshift=-1cm] (1,-0.7) node[circle,fill,inner sep=1pt](a){} -- (5,-0.7) 
-node[circle,fill,inner sep=1pt,label=right:$+\frac{4}{1} \boxed{\text{ iter. 1}}$](b){};
-
-\draw[xshift=-1cm] (3.67,-1.1) node[circle,fill,inner sep=1pt, label=left:$\boxed{\text{iter. 2 }} -\frac{4}{3}$](c){} -- (5,-1.1) 
-node[circle,fill,inner sep=1pt](d){};
-
-\draw[xshift=-1cm] (3.67,-1.5) node[circle,fill,inner sep=1pt](e){} -- (4.46,-1.5) 
-node[circle,fill,inner sep=1pt, label=right:$+\frac{4}{5} \boxed{\text{ iter. 3}}$](f){};
-
-\draw[xshift=-1cm] (3.89,-1.9) node[circle,fill,inner sep=1pt, label=left:$\boxed{\text{iter. 4 }} -\frac{4}{7}$](g){} -- (4.46,-1.9) 
-node[circle,fill,inner sep=1pt](h){};
-
-\draw[xshift=-1cm] (3.89,-2.3) node[circle,fill,inner sep=1pt](i){} -- (4.33,-2.3) 
-node[circle,fill,inner sep=1pt, label=right:$+\frac{4}{9} \boxed{\text{ iter. 5}}$](j){};
-
-\draw[->, to path={-| (\tikztotarget)}]
-  (b) edge (d) (c) edge (e) (f) edge (h) (g) edge (i);
-     \end{tikzpicture}
-
-\label{fig:numberline}
-\caption[]{A number line demonstrating the convergence of Madhava's first method for the 
-calculation of $\pi$. Adapted from (V. M. Jamkar) \cite{jamkar_2018}}
-\end{figure}
+\footnotetext{where $\prod x$ signifies a product. Similar expression to $\sum x$ as in the product of 
+the terms $x$ are calculated, as opposed to the sum}

### 2021-03-07 research project/sections/Computational_approach.tex
rp work
final rp draft

@@
 with the \verb|time.time()| Python function. At each iteration of the method, the number of valid decimal
-places of the resultant approximation are counted and once a specified threshold is reached, 
-a new \verb|t2| time variable is assigned and the time taken, defined by the difference between 
-\verb|t2| and \verb|t1|, is stored. This process is repeated for all specified decimal accuracies and 
+places of the resultant approximation are counted. Once a specified decimal place is reached,
+determined by a programming \verb|if| statement, that can be mathematically expressed as:
+
+$| \pi - \pi_{approx} | < 10^{-n}$, where $\pi_{approx}$ is the approximation by the method and 
+$n$ defines the number of correct decimal places desired.
+
+A new \verb|t2| time variable is assigned and the time taken, defined by $\verb|t2| - \verb|t1|$, is stored. This process is repeated for all specified decimal accuracies and
 for both methods. The times recorded were stored in a \verb|.csv| file for further analysis.
@@
 for comparison between approximated values and the constant $\pi$ that wouldn't
-have been possible using standard Python libraries. \cite{mpmath}
+have been possible using standard Python libraries, due to floating-point limitations (computing 
+term for decimal numbers). \cite{mpmath}

### 2021-03-07 research project/sections/Conclusion.tex
rp
rp work
final rp draft

@@
 used on high-precision floating point numbers, can be said to be more demanding than what could be said
-is simple summation, in the case of Madhava. This essay could be considered in defense of 
-what is known as geometric algebra --- which can clearly produce results more capable than ones 
-originating from an analytical method.
+is simple summation, in the case of Madhava.
+
+An example of this kind of difference in computing power 
+needed for different mathematical operations can be seen in the video-game \textit{Quake III Arena}, published in 1999. 
+The inverse square root of a number needed to be calculated, however due to performance issues, the standard 
+function could not be used. And as such, a new function for the approximation of the inverse square root of a number 
+was programmed \cite{id-software_2012}, highly more efficient than the old one, which shows that a mathematical function 
+implemented in computing can be more efficient than another.
+
+Ultimately, a method 
+of approximating the constant $\pi$ computationally can be more efficient than another, and to the 
+somewhat surprising extent that this method is a geometric one --- many would think an analytical one, 
+that uses simple summation as opposed to square roots, would be more efficient.
 
@@
 
-It must be noted that the experimental procedure in this paper is far from perfect. Firstly, 
-there are many variables that would be difficult to control.
+The method used in this paper is not perfect, and at closer inspection and in 
+retrospect, many issues can be found. Firstly, in the computational implementation of 
+Madhava's method, the value $\sqrt{12}$ is calculated at every iteration, in order to 
+determine the number of correct decimal places, which can 
+lead to efficiency losses. Instead of this, the value could have been calculated once 
+and stored in a variable, to be referenced at every iteration. However, after running 
+the program again with the modified code, the results stayed similar.
 
-\textit{More info to be added}
+Another weakness in the method used in this paper is that a higher number of trials 
+could have been used, as well as the value of $\pi$ approximated to a higher 
+number of decimal places. A higher number of decimal places was disregarded during the 
+implementation of the methods as it was noticed that values over 60 would take an indeterminately 
+long time. This should not have been overlooked, as a different kind of relationship 
+could have been made apparent that way.

### 2021-03-07 research project/sections/Further_research.tex
research project
rp
rp work
final rp draft

@@
 
-\subsection{}
+\subsection{Investigating newer, more efficient algorithms}
+
+As the computation of $\pi$ moved from paper to machines, as all did from accounting to
+mathematics, newer and more efficient methods were developed, notably by Ramanujan and
+the Chudnovsky brothers. These methods should also be investigated, as they 
+are arguably more pertinent in this field.
+
+\subsection{Investigating the uses of efficient code}
+
+This paper does not discuss the many advantages of efficient code in detail, 
+and a research on the speed of the mathematical functions in programming could 
+be conducted. Computer program efficiency is always a pertinent topic because 
+of a faster and faster world. An example of this could be the company Amazon, 
+whose production speed has skyrocketed during the last decade. An increasingly 
+fast demand also requires an increasingly fast computing infrastructure; to handle 
+shipments, manage customers or even conduct job interviews (an experimental system that received 
+major backlash in 2018). \cite{murad_2021}

### 2021-03-08 research project/sections/Background.tex
rp work

@@
 In this paper, we focus on two methods for approximating the value of $\pi$.
-The first by French mathematician François Viète, and another discovered by Madhava of Sangamagrama. Two methods with different
-approaches have been chosen for comparison and the process for original discovery of these methods
-will be explained. The two in question are one based on
-the infinite series definition of an inverse trigonometric function discovered by Madhava
-and one where the value is derived using geometry, by mathematician
-Viète. These two mathematical methods were chosen as they are both
-represent a first occurrence in mathematics: Madhava was the first to
-find the infinite series expansion of the inverse tangent function and Viète was
-the first mathematician to use an infinite product in calculation.
+The first by French mathematician François Viète, and another discovered by Madhava of Sangamagrama.
+Two methods with different approaches have been chosen for comparison and the process for original discovery of these methods will be explained.
+The two in question are one based on the infinite series definition of an inverse trigonometric function discovered by Madhava and one where the value is derived using geometry, by mathematician Viète.
+These two mathematical methods were chosen as they are both represent a first occurrence in mathematics: Madhava was the first to find the infinite series expansion of the inverse tangent function and Viète was the first mathematician to use an infinite product in calculation.
 
@@
 
-Madhava is the first mathematician known to have found the series notation for
-the inverse tangent function, in the Kerala school of mathematics of Medieval India
-\cite{amermathmonthly}. The research made at the Kerala school was documented by
-many mathematicians of their time, namely the astronomer-mathematician Jyesthadeva
-in his treatise \textit{Yukti-Bhasa} written in the Malayalam language \cite{oconnor_robertson_2000}.
-Madhava's infinite series expansion of the inverse tangent function can also be
-found in this treatise. R. C. Gupta, in his translation, describes this expansion as one
-where the arc $\theta$ is equal to the sum of a first term, "the product of the given sine
-and radius of the desired arc divided by the cosine of the arc"\footnotemark, followed by terms that "are
-obtained by a process of iteration": in which the original term is multiplied by the
-square of the sine and divided by the square of the cosine. Thereafter, each term is
-divided by an odd number in order (1, 3, 5, 7, \dots). It is then said that the arc can
-be found by "adding and subtracting respectively the terms of odd rank and those of even rank",
-definition of an alternating series. \cite{rc_gupta_mgseries}
+Madhava is the first mathematician known to have found the series notation for the inverse tangent function, in the Kerala school of mathematics of Medieval India \cite{amermathmonthly}.
+The research made at the Kerala school was documented by many mathematicians of their time, namely the astronomer-mathematician Jyesthadeva in his treatise \textit{Yukti-Bhasa} written in the Malayalam language \cite{oconnor_robertson_2000}.
+Madhava's infinite series expansion of the inverse tangent function can also be found in this treatise.
+R.
+C.
+Gupta, in his translation, describes this expansion as one where the arc $\theta$ is equal to the sum of a first term, "the product of the given sine and radius of the desired arc divided by the cosine of the arc"\footnotemark, followed by terms that "are obtained by a process of iteration": in which the original term is multiplied by the square of the sine and divided by the square of the cosine.
+Thereafter, each term is divided by an odd number in order (1, 3, 5, 7, \dots).
+It is then said that the arc can be found by "adding and subtracting respectively the terms of odd rank and those of even rank", definition of an alternating series.
+\cite{rc_gupta_mgseries}
 
@@
 
-\textit{Yukti-Bhasa} also describes two methods for the calculation of $\pi$. Firstly,
-Madhava's infinite series expansion for $\pi$ is described, which he obtained through
-the previous expansion for the arctangent function. And since $\arctan{1} = \frac{\pi}{4}$,
-it can be said that:
+\textit{Yukti-Bhasa} also describes two methods for the calculation of $\pi$.
+Firstly, Madhava's infinite series expansion for $\pi$ is described, which he obtained through the previous expansion for the arctangent function.
+And since $\arctan{1} = \frac{\pi}{4}$, it can be said that:
 
@@
 
-Which has come to be known as the Leibniz formula for $\pi$, after the German mathematician
-who discovered the same formula two decades later. \cite{edwards_1994} This series converges 
-to $\pi$, as seen in (Figure 1) \ref{fig:numberline}
+Which has come to be known as the Leibniz formula for $\pi$, after the German mathematician who discovered the same formula two decades later.
+\cite{edwards_1994} This series converges to $\pi$, as seen in (Figure 1) \ref{fig:numberline}
 
@@
 
-However, the method with a faster convergence can be found in a further commentary describing the findings of Madhava that states
-a method for the calculation of the circumference $c$ of a circle of diameter $d$ exists.
-The passage in question from the commentary \textit{Tantrasamgraha-vyakhya} of
-anonymous authorship states that, by following the same argument as stated before, one
-can find the circumference of a circle through a similar infinite sum. The first
-term of this sum would be "the square root of the square of the diameter multiplied by
-twelve", followed by the first term "divided by three in each successive case", and
-when these "are divided in order by the odd numbers, beginning with 1", and after the
-even terms are subtracted from sum of the odd", one is left with the circumference of the
-circle (translation from C. K. Raju). \cite{raju_2007}
+However, the method with a faster convergence can be found in a further commentary describing the findings of Madhava that states a method for the calculation of the circumference $c$ of a circle of diameter $d$ exists.
+The passage in question from the commentary \textit{Tantrasamgraha-vyakhya} of anonymous authorship states that, by following the same argument as stated before, one can find the circumference of a circle through a similar infinite sum.
+The first term of this sum would be "the square root of the square of the diameter multiplied by twelve", followed by the first term "divided by three in each successive case", and when these "are divided in order by the odd numbers, beginning with 1", and after the even terms are subtracted from sum of the odd", one is left with the circumference of the circle (translation from C.
+K.
+Raju).
+\cite{raju_2007}
 
@@
 
-François Viète approached the value of $\pi$ from a
-geometric standpoint, and found an infinite product.
-He was able to calculate $\pi$ to a place of 9 decimal points, in
-the year 1593 \cite{Kreminski}, using his method. His method is reminiscent
-of Archimedes' method, where the length of a side is calculated \cite{archimedespi},
-but differs in that it consists of finding the area of a polygon of $n$ sides in a circle of constant
-radius, rather than the circumference. As the value of $n$ is increased, the area of the $n$-gon
-tends toward the area of a circle. The geometric origin of this formula
-can be found using simple right-angle trigonometry, by first finding
-the lengths $OH$ and subsequently $BD$ in (Figure 2) \ref{fig:image1}.
+François Viète approached the value of $\pi$ from a geometric standpoint, and found an infinite product.
+He was able to calculate $\pi$ to a place of 9 decimal points, in the year 1593 \cite{Kreminski}, using his method.
+His method is reminiscent of Archimedes' method, where the length of a side is calculated \cite{archimedespi}, but differs in that it consists of finding the area of a polygon of $n$ sides in a circle of constant radius, rather than the circumference.
+As the value of $n$ is increased, the area of the $n$-gon tends toward the area of a circle.
+The geometric origin of this formula can be found using simple right-angle trigonometry, by first finding the lengths $OH$ and subsequently $BD$ in (Figure 2) \ref{fig:image1}.
 
@@
 Since the equation for the area of a polygon is defined as $A = \frac{p \cdot a}{2}$,
-where $p$ is the perimeter of the polygon and $a$ is the apothem\footnote{apothem: the segment 
-from the center of a polygon to the midpoint of one of its sides, here $OH$ for the $n$-gon and 
-$OH'$ for the $2n$-gon}, in this case $BD \cdot n$
-and $OH$ respectively, let $A_{n}$ equal the area of
-the polygon with $n$ sides (displayed in \textcolor{green}{green} in Figure 2) such that: \\
-$A_{n} = \frac{OH \cdot BD \cdot n}{2}$ \\
+where $p$ is the perimeter of the polygon and $a$ is the apothem\footnote{apothem: the segment from the center of a polygon to the midpoint of one of its sides, here $OH$ for the $n$-gon and $OH'$ for the $2n$-gon}, in this case $BD \cdot n$ and $OH$ respectively, let $A_{n}$ equal the area of the polygon with $n$ sides (displayed in \textcolor{green}{green} in Figure 2) such that:
+
+$A_{n} = \frac{OH \cdot BD \cdot n}{2}$
+
 $A_{n} = \frac{R \cos{\alpha} \cdot 2 R \sin{\alpha} \cdot n}{2} = n R^2 \sin{\alpha} \cos{\alpha}$
 
-And if n is multiplied by 2, the angle $\angle \alpha$ is divided by 2,
-and the new area becomes:
+And if n is multiplied by 2, the angle $\angle \alpha$ is divided by 2, and the new area becomes:
 
@@
 
-So it can be written that, by definition, the ratio of
-the area of an $n$-gon to one of a $2n$-gon is
+So it can be written that, by definition, the ratio of the area of an $n$-gon to one of a $2n$-gon is
 
@@
 
-Which through the trigonometric identity $\sin{2 \theta} = 2\sin{\theta} \cos{\theta}$ can
-be simplified to:
+Which through the trigonometric identity $\sin{2 \theta} = 2\sin{\theta} \cos{\theta}$ can be simplified to:
 
@@
 
-It can be then written that, through a new variable $P$, \\
-$P = \frac{A_{n}}{A_{2n}} \frac{A_{2n}}{A_{4n}} \frac{A_{4n}}{A_{8n}} \dots \frac{A_{(k-2)n}}{A_{kn}} \frac{A_{kn}}{A}$ \\
-where $A$ is the area of the circle of radius $R$
-in \ref{fig:image1}.
+It can be then written that, through a new variable $P$,
 
-So $P = \frac{A_{n}}{A}$, since the values $A_{kn}$ cancel, and
-so it is true that $P = \frac{A_{n}}{A} \Leftrightarrow A = \frac{A_{n}}{P}$. The
-value of $R = 1$ in this case, and since the area of a circle is defined
-by $A = \pi R^2$, therefore: \\
-$\pi = \frac{A_{n}}{P}$ \\
-By definition, we can say that as the value $k$ approaches infinity, the area of the $kn$-gon
-approaches that of a circle, and therefore, the value of $\pi$.
+$P = \frac{A_{n}}{A_{2n}} \frac{A_{2n}}{A_{4n}} \frac{A_{4n}}{A_{8n}} \dots \frac{A_{(k-2)n}}{A_{kn}} \frac{A_{kn}}{A}$
+
+where $A$ is the area of the circle of radius $R$ in \ref{fig:image1}.
+
+So $P = \frac{A_{n}}{A}$, since the values $A_{kn}$ cancel, and so it is true that $P = \frac{A_{n}}{A} \Leftrightarrow A = \frac{A_{n}}{P}$.
+The value of $R = 1$ in this case, and since the area of a circle is defined by $A = \pi R^2$, therefore:
+
+$\pi = \frac{A_{n}}{P}$
+
+By definition, we can say that as the value $k$ approaches infinity, the area of the $kn$-gon approaches that of a circle, and therefore, the value of $\pi$.
 
@@
 Where the value of $A_{n}$ is the area of the first polygon, with $n=4$ sides, and as such
-$A_{n} = 4 \sin{45} \cos{45} = 2$. We can define: \\
-$U_{0} = \cos{a} = \cos{45} = \frac{1}{\sqrt{2}}$ \\
-$U_{1} = cos{\frac{\alpha}{2}}$ \\
+$A_{n} = 4 \sin{45} \cos{45} = 2$. We can define:
+
+$U_{0} = \cos{a} = \cos{45} = \frac{1}{\sqrt{2}}$
+
+$U_{1} = cos{\frac{\alpha}{2}}$
+
 Which we can, through the trigonometric identity
 $\cos^2{\theta} = \frac{1}{2} + \frac{1}{2} \cos{2\theta}$, simplify as
-$U_{1} = \sqrt{\frac{1}{2} + \frac{1}{2} U_{0}}$ \\
-So it can be said that $U_{n} = \sqrt{\frac{1}{2} + \frac{1}{2} U_{n - 1}}$, which leads to
-a fully defined expression for the value of pi, using an infinite product:
+
+$U_{1} = \sqrt{\frac{1}{2} + \frac{1}{2} U_{0}}$
+
+So it can be said that $U_{n} = \sqrt{\frac{1}{2} + \frac{1}{2} U_{n - 1}}$, which leads to a fully defined expression for the value of pi, using an infinite product:
 
@@
 
-\footnotetext{where $\prod x$ signifies a product. Similar expression to $\sum x$ as in the product of 
-the terms $x$ are calculated, as opposed to the sum}
+\footnotetext{where $\prod x$ signifies a product. Similar expression to $\sum x$ as in the product of the terms $x$ are calculated, as opposed to the sum}

### 2021-03-09 research project/sections/Analysis_of_results.tex
final rp draft
rp work

@@
 
-The arithmetic mean of a decimal place was calculated for 100 trials and the data points
-in the table shown below were found (rounded to 3 significant figures). The values for the decimal place
-0 have been dropped as it would not be logical to include them.
+The arithmetic mean of a decimal place was calculated for 100 trials and the data points in the table shown below were found (rounded to 3 significant figures).
+The values for the decimal place 0 have been dropped as it would not be logical to include them.
 
@@
 
-The iterations needed for both methods were also measured, using the Python program
-(see appendix), slightly modified to increment a variable \verb|i| and subsequently
-return it, writing it in a similar manner to a \verb|.csv| file.
+The iterations needed for the approximation to a decimal place of both methods were also measured, using the Python program (see appendix), slightly modified to increment a variable \verb|i| and subsequently return it, writing it in a similar manner to a \verb|.csv| file.
 
@@
 
-Shown below is a graph demonstrating the correlation between the decimal places approximated
-and the amount of iterations needed for this specific decimal milestone. The blue markers and
-line of best-fit respresent the results received from Viète's method while the red represent those
-from Madhava's method. \\
+Shown below is a graph demonstrating the correlation between the decimal places approximated and the amount of iterations needed for this specific decimal milestone.
+The blue markers and line of best-fit represent the results received from Viète's method while the red represent those from Madhava's method.
 
@@
 
-The results gathered from this experiment show linear relationship between the amount of decimal
-places approximated and the time required for this approximation, as well as between the decimal places
-and the amount of iterations needed. The two lines of best fit for the two methods show
-similarities but also differences.
+The results gathered from this experiment show linear relationship between the amount of decimal places approximated and the time required for this approximation, as well as between the decimal places and the amount of iterations needed.
+The two lines of best fit for the two methods show similarities but also differences.
 
-Where decimal places and iterations are compared, there is a clear difference between the two
-methods: Madhava's method indeed takes more iterations to converge to a specific decimal of
-the value $\pi$. However, in the second graph comparing time, it can be noted that Madhava's
-method is marginally faster until the decimal 20, when the lines of the two methods stray apart. It could be
-assumed that this occurrence is due to differences between the types of the two methods. Although it 
-could be argued that this difference is only due to random error, the results in this graph 
-are those collected from the arithmetic mean over 100 trials, and so it can be assessed that this is not the case. Madhava's
-method is what is called an alternating series, due to its nature of alternating between values in order
-to converge to a specific value, in this case $\pi$, while Viète's method is an infinite product. The
-difference between these two methods is underlined in (Figure 3) \ref{fig:comparaison}. It can be
-deduced that at lower decimal values Madhava's method is faster as Viète's method's approximation
-accuracy firstly increases in what seems similar to a curve from an exponential function $f(x) = k^x$.
-It can also be said that while Madhava's method alternates between values such that $m_{n} > \pi, m_{n+1} < \pi$,
-Viète's method $\pi_v$ approximates pi in a manner such that $\pi_v < \pi$.
+Where decimal places and iterations are compared, there is a clear difference between the two methods: Madhava's method indeed takes more iterations to converge to a specific decimal of the value $\pi$.
+However, in the second graph comparing time, it can be noted that Madhava's method is marginally faster until the decimal 20, when the lines of the two methods stray apart.
+It could be assumed that this occurrence is due to differences between the types of the two methods.
+Although it could be argued that this difference is only due to random error, the results in this graph are those collected from the arithmetic mean over 100 trials, and so it can be assessed that this is not the case.
+Madhava's method is what is called an alternating series, due to its nature of alternating between values in order to converge to a specific value, in this case $\pi$, while Viète's method is an infinite product.
+The difference between these two methods is underlined in (Figure 3) \ref{fig:comparaison}.
+It can be deduced that at lower decimal values Madhava's method is faster as Viète's method's approximation accuracy firstly increases in what seems similar to a curve from an exponential function $f(x) = k^x$.
+It can also be said that while Madhava's method alternates between values such that $m_{n} > \pi, m_{n+1} < \pi$, Viète's method $\pi_v$ approximates pi in a manner such that $\pi_v < \pi$.
 

### 2021-03-09 research project/sections/Background.tex
final rp draft
rp work

@@
 
-And because $\frac{\sin{\theta}}{\cos{\theta}} = \tan{\theta}$, the aforementioned expression
+And because $\frac{\sin{\theta}}{\cos{\theta}} = \tan{\theta}$, the expression
 can be simplified to:
@@
 
-Since the equation for the area of a polygon is defined as $A = \frac{p \cdot a}{2}$,
-where $p$ is the perimeter of the polygon and $a$ is the apothem\footnote{apothem: the segment from the center of a polygon to the midpoint of one of its sides, here $OH$ for the $n$-gon and $OH'$ for the $2n$-gon}, in this case $BD \cdot n$ and $OH$ respectively, let $A_{n}$ equal the area of the polygon with $n$ sides (displayed in \textcolor{green}{green} in Figure 2) such that:
+Since the equation for the area of a polygon is defined as $A = \frac{p \cdot a}{2}$, where $p$ is the perimeter of the polygon and $a$ is the apothem\footnote{apothem: the segment from the center of a polygon to the midpoint of one of its sides, here $OH$ for the $n$-gon and $OH'$ for the $2n$-gon}, in this case $BD \cdot n$ and $OH$ respectively, let $A_{n}$ equal the area of the polygon with $n$ sides (displayed in \textcolor{green}{green} in Figure 2) such that:
 
@@
 
-$\frac{A_{n}}{\cos{\alpha} \cos{\frac{\alpha}{2}} \cos{\frac{\alpha}{4}} \dots}
-  \to_{k \to \infty} \pi$
+$\frac{A_{n}}{\cos{\alpha} \cos{\frac{\alpha}{2}} \cos{\frac{\alpha}{4}} \dots}\to_{k \to \infty} \pi$
 
-Where the value of $A_{n}$ is the area of the first polygon, with $n=4$ sides, and as such
-$A_{n} = 4 \sin{45} \cos{45} = 2$. We can define:
+Where the value of $A_{n}$ is the area of the first polygon, with $n=4$ sides, and as such $A_{n} = 4 \sin{45} \cos{45} = 2$. We can define:
 
@@
 
-Which we can, through the trigonometric identity
-$\cos^2{\theta} = \frac{1}{2} + \frac{1}{2} \cos{2\theta}$, simplify as
+Which we can, through the trigonometric identity $\cos^2{\theta} = \frac{1}{2} + \frac{1}{2} \cos{2\theta}$, simplify as
 
@@
 
-So it can be said that $U_{n} = \sqrt{\frac{1}{2} + \frac{1}{2} U_{n - 1}}$, which leads to a fully defined expression for the value of pi, using an infinite product:
+$U_{n} = \sqrt{\frac{1}{2} + \frac{1}{2} U_{n - 1}}$, which leads to a fully defined expression for the value of pi, using an infinite product:
 

### 2021-03-09 research project/sections/Computational_approach.tex
final rp draft
rp work

@@
 \subsection{The variables}
-The dependent variable of this experiment is the time taken $t$ by the program
-to approximate a given number $n$ of correct decimal value of the constant $\pi$.
+The dependent variable of this experiment is the time taken $t$ by the program to approximate a given number $n$ of correct decimal values in the constant $\pi$.
 
-The value $n$ will be altered in order to avoid possible similar convergence rates
-at a small amount of decimal places, and multiple trials will be run to decrease
-margin of error. Other variables of the experiment will be controlled. For example, the experiment
-will be run on a same isolated system, a virtual machine, with a minimal amount of
-processes running simultaneously to avoid any possible variance in results.
+The value $n$ will be altered in order to avoid possible similar convergence rates at a small amount of decimal places, and multiple trials will be run to decrease margin of error.
+Other variables of the experiment will be controlled.
+For example, the experiment will be run on a same isolated system, a virtual machine, with a minimal amount of processes running simultaneously to avoid any possible variance in results.
 
@@
 
-A Python application was programmed (see appendix) in order to run the two
-methods aforementioned, and manage the collection of data.
+A Python application was programmed (see appendix) in order to run the two methods, and manage the collection of data.
 
-The program assigns the time before the execution of the method to a variable \verb|t1|
-with the \verb|time.time()| Python function. At each iteration of the method, the number of valid decimal
-places of the resultant approximation are counted. Once a specified decimal place is reached,
-determined by a programming \verb|if| statement, that can be mathematically expressed as:
+The program assigns the time before the execution of the method to a variable \verb|t1| with the \verb|time.time()| Python function.
+At each iteration of the method, the number of valid decimal places of the resultant approximation are counted.
+Once a specified decimal place is reached, determined by a programming \verb|if| statement, that can be mathematically expressed as:
 
-$| \pi - \pi_{approx} | < 10^{-n}$, where $\pi_{approx}$ is the approximation by the method and 
-$n$ defines the number of correct decimal places desired.
+$| \pi - \pi_{approx} | < 10^{-n}$, where $\pi_{approx}$ is the approximation by the method and $n$ defines the number of correct decimal places desired.
 
-A new \verb|t2| time variable is assigned and the time taken, defined by $\verb|t2| - \verb|t1|$, is stored. This process is repeated for all specified decimal accuracies and
-for both methods. The times recorded were stored in a \verb|.csv| file for further analysis.
+A new \verb|t2| time variable is assigned and the time taken, defined by $\verb|t2| - \verb|t1|$, is stored.
+This process is repeated for all specified decimal accuracies and for both methods.
+The times recorded were stored in a \verb|.csv| file for further analysis.
 
-The \verb|mpmath| library was used for the floating point operations required
-for comparison between approximated values and the constant $\pi$ that wouldn't
-have been possible using standard Python libraries, due to floating-point limitations (computing 
-term for decimal numbers). \cite{mpmath}
+The \verb|mpmath| library was used for the floating point operations required for comparison between approximated values and the constant $\pi$ that wouldn't have been possible using standard Python libraries, due to floating-point limitations (computing term for decimal numbers).
+\cite{mpmath}

### 2021-03-09 research project/sections/Conclusion.tex
final rp draft
rp work

@@
 
-To conclude, even though Viète's method can seem more convoluted in practice, due to its geometric origins,
-it can be seen through the experiment conducted in this paper that this is clearly not the case
-in a computational context. His method is simplified to a point where its geometric origins are no
-longer findable from its computational implementation. And despite the assumption that could be
-made based on the types of operations needed for each method, Viète's requiring arguably more
-computationally challenging operations. These operations, such as square-root function or the product function
-used on high-precision floating point numbers, can be said to be more demanding than what could be said
-is simple summation, in the case of Madhava.
+To conclude, even though Viète's method can seem more convoluted in practice, due to its geometric origins, it can be seen through the experiment conducted in this paper that this is clearly not the case in a computational context.
+His method is simplified to a point where its geometric origins are no longer findable from its computational implementation.
+And despite the assumption that could be made based on the types of operations needed for each method, Viète's requiring arguably more computationally challenging operations.
+These operations, such as square-root function or the product function used on high-precision floating point numbers, can be said to be more demanding than what could be said is simple summation, in the case of Madhava.
 
-An example of this kind of difference in computing power 
-needed for different mathematical operations can be seen in the video-game \textit{Quake III Arena}, published in 1999. 
-The inverse square root of a number needed to be calculated, however due to performance issues, the standard 
-function could not be used. And as such, a new function for the approximation of the inverse square root of a number 
-was programmed \cite{id-software_2012}, highly more efficient than the old one, which shows that a mathematical function 
-implemented in computing can be more efficient than another.
+An example of this kind of difference in computing power needed for different mathematical operations can be seen in the video-game \textit{Quake III Arena}, published in 1999.
+The inverse square root of a number needed to be calculated, however due to performance issues, the standard function could not be used.
+And as such, a new function for the approximation of the inverse square root of a number was programmed \cite{id-software_2012}\cite{beyond3d_2006}, highly more efficient than the old one, which shows that a mathematical function implemented in computing can be more efficient than another.
 
-Ultimately, a method 
-of approximating the constant $\pi$ computationally can be more efficient than another, and to the 
-somewhat surprising extent that this method is a geometric one --- many would think an analytical one, 
-that uses simple summation as opposed to square roots, would be more efficient.
+Ultimately, a method of approximating the constant $\pi$ computationally can be more efficient than another, and to the somewhat surprising extent that this method is a geometric one --- many would think an analytical one, that uses simple summation as opposed to square roots, would be more efficient.
 
@@
 
-The method used in this paper is not perfect, and at closer inspection and in 
-retrospect, many issues can be found. Firstly, in the computational implementation of 
-Madhava's method, the value $\sqrt{12}$ is calculated at every iteration, in order to 
-determine the number of correct decimal places, which can 
-lead to efficiency losses. Instead of this, the value could have been calculated once 
-and stored in a variable, to be referenced at every iteration. However, after running 
-the program again with the modified code, the results stayed similar.
+The method used in this paper is not perfect, and at closer inspection and in retrospect, many issues can be found.
+Firstly, in the computational implementation of Madhava's method, the value $\sqrt{12}$ is calculated at every iteration, in order to determine the number of correct decimal places, which can lead to efficiency losses.
+Instead of this, the value could have been calculated once and stored in a variable, to be referenced at every iteration.
+However, after running the program again with the modified code, the results stayed similar.
 
-Another weakness in the method used in this paper is that a higher number of trials 
-could have been used, as well as the value of $\pi$ approximated to a higher 
-number of decimal places. A higher number of decimal places was disregarded during the 
-implementation of the methods as it was noticed that values over 60 would take an indeterminately 
-long time. This should not have been overlooked, as a different kind of relationship 
-could have been made apparent that way.
+Another weakness in the method used in this paper is that a higher number of trials could have been used, as well as the value of $\pi$ approximated to a higher number of decimal places.
+A higher number of decimal places was disregarded during the implementation of the methods as it was noticed that values over 60 would take an indeterminately long time.
+This should not have been overlooked, as a different kind of relationship could have been made apparent that way.

### 2021-03-09 research project/sections/Further_research.tex
final rp draft
rp work

@@
 
-As the computation of $\pi$ moved from paper to machines, as all did from accounting to
-mathematics, newer and more efficient methods were developed, notably by Ramanujan and
-the Chudnovsky brothers. These methods should also be investigated, as they 
-are arguably more pertinent in this field.
+As the computation of $\pi$ moved from paper to machines, as all did from accounting to mathematics, newer and more efficient methods were developed, notably by Ramanujan and the Chudnovsky brothers.
+These methods should also be investigated, as they are arguably more pertinent in this field.
 
@@
 
-This paper does not discuss the many advantages of efficient code in detail, 
-and a research on the speed of the mathematical functions in programming could 
-be conducted. Computer program efficiency is always a pertinent topic because 
-of a faster and faster world. An example of this could be the company Amazon, 
-whose production speed has skyrocketed during the last decade. An increasingly 
-fast demand also requires an increasingly fast computing infrastructure; to handle 
-shipments, manage customers or even conduct job interviews (an experimental system that received 
-major backlash in 2018). \cite{murad_2021}
+This paper does not discuss the many advantages of efficient code in detail, and a research on the speed of the mathematical functions in programming could be conducted.
+Computer program efficiency is always a pertinent topic because of a faster and faster world.
+An example of this could be the company Amazon, whose production speed has skyrocketed during the last decade.
+An increasingly fast demand also requires an increasingly fast and accurate computing infrastructure; to handle shipments, manage customers or even conduct job interviews (an experimental system that received major backlash in 2018).
+\cite{murad_2021}

### 2021-03-09 research project/sections/Introduction.tex
rp
rp work
final rp draft

@@
 
-The value of $\pi$ has been researched for many years, although under different names,
-and the amount of different approaches to reach the value is large. The value has been 
-found through many processes, be it analytically, the most popular, geometrically, or through 
-more obscure or convoluted methods, such as the possibility to approximate the value using 
-physics akin to those from a simple game of billiards. \cite{galperin_2003}
+The value of $\pi$ has been researched for many years, although under different names, and the amount of different approaches to reach the value is large.
+The value has been found through many processes, be it analytically, the most popular, geometrically, or through more obscure or convoluted methods, such as the possibility to approximate the value using physics akin to those from a simple game of billiards.
+\cite{galperin_2003}
 
-This paper seeks to examine the extent at which two historical methods of
-approximation of the value $\pi$, namely the approaches suggested by the aforementioned 
-mathematicians Madhava and Viète, differ in terms of computational speed and 
-speed, and explain these differences. This paper does not however, 
-suggest a method to use for computation but rather seeks to only compare 
-the speed of a geometrically derived formula and a analytically derived one.
+This paper seeks to examine the extent at which two historical methods of approximation of the value $\pi$, namely the approaches suggested by the mathematicians Madhava and Viète, differ in terms of computational speed, and explain these differences.
+This paper does not however, suggest a method to use for computation but rather seeks to only compare the speed of a geometrically derived formula and a analytically derived one.
 
-This research could prove useful in the field of computer science, as 
-there is always a demand for faster and more efficient programs in an
-ever-changing society.
-
-Furthermore, a research based on computational speed of two different 
-kinds approaches to the constant $\pi$ has not been done to date.
+This research could prove useful in the field of computer science, as there is always a demand for faster and more efficient programs in an ever-changing society.
+Furthermore, a research based on computational speed of two different kinds approaches to the constant $\pi$ has not been done to date.

### 2021-03-09 research project/sections/Conclusion.tex
rp work

@@
 This should not have been overlooked, as a different kind of relationship could have been made apparent that way.
+
+\subsection{Personal reflection}
+
+This paper has been interesting to write for me.
+The topic I chose seemed daunting in the beginning, and math was not a favorite subject of mine to say the least.
+I had to research concepts outside my understanding and the curriculum, and challenge my own programming knowledge.
+To add to this, I also managed to understand how important time management skills are for these types of tasks.
+By challenging myself like this, I think I have not only learned to like the field of mathematics but also develop important skills.
+I have also learned to use new software, notably the typesetting software \LaTeX{} and its bibliography management tool \verb|bibtex|, which proved very useful in the writing of this paper, and facilitated many boring tasks.
+To finally conclude, I think the skills I have learned through this paper (along with the probably numerous mistakes), will help me understand and organize myself for future projects of the same kind.

### 2021-03-15 math/ia_2021/paper.tex
rp work
sl oral
chem

@@
 
-$\hat \beta_1 = \frac{1806757.14}{322280.86} = 5.60$ and $\hat \beta_0 = 2257.14 - 5.61 \cdot 512.86 = -618.01$
+$\hat \beta_1 = \frac{1806757.14}{322280.86} = 5.60$ and $\hat \beta_0 = 2257.14 - 5.60 \cdot 512.86 = -618.01$
 
-So the regression line can be expressed as $\boxed{\hat y = 5.60 x -618.01}$
+So the regression line can be expressed as $\boxed{ \hat y = 5.60 x -618.01}$
 
@@
 
-However, the relationship between the distance between an earthquake and its tsunami
-and the wave height is not necessarily linear, especially if it is considered that the travel of
-sound (which is a wave, or the displacement of a fluid that can travel) in a fluid obeys the
-inverse square law. Therefore, a simple linear regression model cannot be
-used for this research. A second degree polynomial regression formula would be more
-appropriate, and a new $x^2$ term can be added to the previous formula. We get:
+However, the relationship between the distance between an earthquake and its tsunami and the wave height is not necessarily linear, especially if it is considered that the travel of sound (which is a wave, or the displacement of a fluid that can travel) in a fluid obeys the inverse square law.
+Therefore, a simple linear regression model cannot be used for this research.
+A second degree polynomial regression formula would be more appropriate, and a new $x^2$ term can be added to the previous formula.
+We get:
 
@@
 
-Even though the right side of this equation is quadratic in $x$, this is still
-linear regression as the right side is linear in the $\beta$ terms (if we were to let
-$x = 1$, we would be left with $y = a + \beta_1 + \beta_2 + e$, which is linear).
-A second degree polynomial regression would best fit our model for both independent
-variables, as seen in the graphs below, where a theoretical curve of regression is drawn
-by hand:
+Even though the right side of this equation is quadratic in $x$, this is still linear regression as the right side is linear in the $\beta$ terms (if we were to let $x = 1$, we would be left with $y = a + \beta_1 + \beta_2 + e$, which is linear).
+A second degree polynomial regression would best fit our model for both independent variables, as seen in the graphs below, where a theoretical curve of regression is drawn by hand:
 
@@
 
-In addition, our dataset has two independent variables: the distance from a particular
-earthquake event and the magnitude of that earthquake. Models such as the one explored
-in this paper, where multiple independent variables are present, are called multivariate
-and can be modeled by adding the terms of a new independent variable, as such
-for a linear model with two independent variables:
+In addition, our dataset has two independent variables: the distance from a particular earthquake event and the magnitude of that earthquake.
+Models such as the one explored in this paper, where multiple independent variables are present, are called multivariate and can be modeled by adding the terms of a new independent variable, as such for a linear model with two independent variables:
 
@@
 
-So the relationship between the two independent variables and the wave height can
-be estimated with a multivariate, polynomial regression model. We will use the Python
-scikit-learn library to aid in creating this model. \\
+So the relationship between the two independent variables and the wave height can be estimated with a multivariate, polynomial regression model.
+We will use the Python scikit-learn library to aid in creating this model.
+\\
 
-(\textit{Section not done, need to do more research on how the} \verb|PolynomialFeatures| 
-\textit{function transforms a linear array to one fit for polynomial regression})
+(\textit{Section not done, need to do more research on how the} \verb|PolynomialFeatures| \textit{function transforms a linear array to one fit for polynomial regression})
 
@@
 
-The model's fit is then calculated, with the test dataset. The \verb|accuracy_score| method
-from the \verb|sklearn.metrics| class was used for this task. This function calculates what
-is called the $R^2$ value, which explains the extent at which the variance (the amount by which numbers
-are spread from the average value) in the dependent variable can be explained by the independent
-variable(s). An $R^2$ value of 0.6 indicates that 60\% of the model's results match those of the
-dataset. This value is defined as:
+The model's fit is then calculated, with the test dataset.
+The \verb|accuracy_score| method from the \verb|sklearn.metrics| class was used for this task.
+This function calculates what is called the $R^2$ value, which explains the extent at which the variance (the amount by which numbers are spread from the average value) in the dependent variable can be explained by the independent variable(s).
+An $R^2$ value of 0.6 indicates that 60\% of the model's results match those of the dataset.
+This value is defined as:
 
@@
 
-$SS_{res} = \sum\limits_{i} e_i^2 = \sum\limits_{i} (y_i - p_i)^2$, \footnote{$\sum\limits_{i=1}^n x_i$ means that all values of
-    x are summed, beginning from $x_1$ and ending at $x_n$} where $e_i$ is the error term, equal to the difference of a data point $y_i$
-and its corresponding predicted value $p_i$.
+$SS_{res} = \sum\limits_{i} e_i^2 = \sum\limits_{i} (y_i - p_i)^2$, \footnote{$\sum\limits_{i=1}^n x_i$ means that all values of x are summed, beginning from $x_1$ and ending at $x_n$} where $e_i$ is the error term, equal to the difference of a data point $y_i$ and its corresponding predicted value $p_i$.
 
-and $SS_{tot}$ is the total sum of squares, which is proportional to the variance in the data,
-defined as:
+and $SS_{tot}$ is the total sum of squares, which is proportional to the variance in the data, defined as:
 
@@
 
-This method, like any other statistical approximation, is by no means perfect
-and for that reason should not be used in any real-world application. The results
-procured by the program are subject to many uncertainties, one of which is described
-below.
+This method, like any other statistical approximation, is by no means perfect and for that reason should not be used in any real-world application.
+The results procured by the program are subject to many uncertainties, one of which is described below.
 
-When attempting to predict the maximum wave height of a tsunami such as the one
-caused by the 2011 earthquake off the Pacific coast of Tōhoku, approximately 70km
-away from shore, which produced a 38.9m wave \cite{yomiuri_2011}. The model in this
-paper approximates this same event as only a 8m wave, when supplied the 9.0 magnitude
-of the earthquake and the 70km distance. These types of errors can be attributed to the
-fact that there are many more independent variables, such as the shape of the seafloor for
-example, that can amplify or potentially change the height of a tsunami's wave.
+When attempting to predict the maximum wave height of a tsunami such as the one caused by the 2011 earthquake off the Pacific coast of Tōhoku, approximately 70km away from shore, which produced a 38.9m wave \cite{yomiuri_2011}.
+The model in this paper approximates this same event as only a 8m wave, when supplied the 9.0 magnitude of the earthquake and the 70km distance.
+These types of errors can be attributed to the fact that there are many more independent variables, such as the shape of the seafloor for example, that can amplify or potentially change the height of a tsunami's wave.
 
@@
 
-All figures are generated through the use of TikZ (Tantau) and the Python library 
-\verb|matplotlib|. This paper is written with the aid of the typesetting software \LaTeX. 
-On a PDF viewer, the sections in the table of contents can be clicked to access that 
-section. Dataset not included due to large size (\textasciitilde 28000 columns).
+All figures are generated through the use of TikZ (Tantau) and the Python library \verb|matplotlib|.
+This paper is written with the aid of the typesetting software \LaTeX.
+On a PDF viewer, the sections in the table of contents can be clicked to access that section.
+Dataset not included due to large size (\textasciitilde 28000 columns).
 
@@
 
-Below is an example of a dataset with $x$ values and their corresponding
-Y values, the data points chosen are purely arbitrary.
+Below is an example of a dataset with $x$ values and their corresponding Y values, the data points chosen are purely arbitrary.
 
@@
 
-Listed here is the Python program through which the dataset was cleaned and
-predictions based on the statistical model were made. Run using 64-bit Python
-version 3.9.1.
+Listed here is the Python program through which the dataset was cleaned and predictions based on the statistical model were made. Run using 64-bit Python version 3.9.1.
 

### 2021-05-03 geography/ia/paper.tex
geo, chem

--- 2021-03-15 math/ia_2021/paper.tex
@@
         \vspace*{4cm}
-        A statistical approach to predicting a tsunami’s wave height \\
+        Geography IA \\
         \vspace{1cm}
@@
 
-\subsection{Background}
+Lorem ipsum dolor sit amet
 
-Naturally occurring disasters, such as tsunamis, are devastating for people, human settlements,
-buildings and other resources residing in risk areas. Tsunamis are caused
-by many different types of natural phenomena, such as landslides, collapse of
-seamount, or more rarely, the impact of a meteorite. This paper, however, focuses
-on one of the more common causes of tsunamis; earthquakes. Earthquakes cause around 72\%
-of tsunamis \cite{pacifictsunamimuseum}. The time between an earthquake and its tsunami is
-relatively long compared to events such as landslides \cite{sue_nokes_walters},
-leaving adequate time to prepare for impact. Predicting wave height, as discussed in this
-paper, can further help to assess the seriousness of the tsunami and provide help to
-prepare for the event.
+\subsection{Study area}
 
-\subsection{Personal Engagement}
+Lorem ipsum dolor sit amet
 
-I have chosen to write about this subject as I have recently become interested by the
-statistics and its implications in real-life matters. This research has proved very
-interesting for me as it has made the field of statistics more concrete, because it provides
-a real world use case. Furthermore, as I enjoy programming, the research I have done
-has helped me learn about the tools and techniques used by data analysts and other professionals.
+\subsection{Theory}
 
-\section{Statistical Model}
+Lorem ipsum dolor sit amet
 
-\subsection{Processing the dataset for use}
-%TODO: #2 Ask for help 
-The dataset, procured from the National Oceanic and Atmospheric Administration website \cite{noaa}, includes many tsunami
-runups, going back before the year 0. Due to this, the NOAA states uncertainties
-that can occur in its data. For example, reports that were recorded before the
-20th century tend to be based on written accounts, and not recorded seismic activity.
-Only since the 1960s, the installation of the Worldwide Standardized Seismograph
-Network system allowed for more accurate data in this paper. For that reason, any recorded
-activity before this date was discarded from the dataset for this paper. The dataset itself can
-be accessed through two different methods; searching the database using the online
-tool or downloading it in a somewhat obscure compressed Google Earth \verb|.kmz| format.
-The online tool supports the download of search results in a much more human-readable
-\verb|.tsv| (tab-separated) format. For this paper, all entries after the year 1900 were searched and downloaded.
+\section{Method}
 
-The data is cleaned of data points marked as "dubious" (there is an column for this parameter),
-reports before the year 1960 (see above), and of all irrelevant fields such as "Injuries"
-or "Damage \$Mil". This preliminary data cleanup
-is done through the use of the pandas Python library \cite{reback2020pandas}\cite{mckinney-proc-scipy-2010},
-which is commonly used for the handling of large datasets for its efficiency.
+Lorem ipsum dolor sit amet
 
-The data was split into a training and a test dataset. The training
-dataset was used to generate a regression model while the test dataset was used
-to calculate the general accuracy of the model, by predicting values and
-comparing those to the known value as found in the dataset. The dataset was split based on
-a fraction using the \verb|df.sample| method found in the pandas library. A split
-of 70\% for training and 30\% for test was chosen. A \verb|random_state| argument is
-set to an arbitrary value 42, as if it weren't, the dataset would be split in a different
-manner at each execution of the program. The program can be found in Appendix \ref{app:scripts}.
+\subsection{Sampling choices}
 
-\subsection{Regression}
+Lorem ipsum dolor sit amet
 
-\subsubsection{Linear model}
+\subsection{Study site choices}
 
-Regression analysis is a commonly used tool in statistics, and regression models can
-be found in numerous use cases. Regression models consist of estimating the
-relationship between one (or more) independent variables and a dependent variable. A
-simple example of regression can be seen in the linear relationship between an independent
-variable $x_{i}$, and the dependent variable $y$. The following expression shows this relationship in what is
-called linear regression, as it models a straight line that can be used for prediction of
-values:
+Lorem ipsum dolor sit amet
 
-$y =  \beta_0 a + \beta_1 x + e $,
+\section{Analysis}
 
-where the term $\beta_0$ represents the y-intercept of the line and $\beta$ represents the
-coefficient of $x$. The final term $e$ is what is known as the error or residual
-term. It is an estimate of the amount by which the predictions differ from the real
-value. An example of linear regression can be found in Appendix \ref{app:regression}, where
-a set of data points is presented on a scatter plot, and the regression line is found
-by finding the individual coefficients $\beta_n$ of $x$ (see equation above), like such:
+Lorem
 
-$\hat \beta_1 = \frac{\sum (x_i - \bar x) (y_i - \bar y)}{\sum (x_i - \bar x)^2}$, where
-$\hat \beta_n$ represents an estimated value and $\bar x$ represents the mean of the set $x$
+\subsection{Environmental and cultural sustainability}
+Lorem ipsum dolor sit amet
 
-and the $\hat \beta_0$ value can be approximated as $\hat \beta_0 = \bar y - \hat \beta_1 \bar x$
-
-For the dataset in \ref{app:regression}, the results to these equations are:
-
-$\hat \beta_1 = \frac{1806757.14}{322280.86} = 5.60$ and $\hat \beta_0 = 2257.14 - 5.60 \cdot 512.86 = -618.01$
-
-So the regression line can be expressed as $\boxed{ \hat y = 5.60 x -618.01}$
-
-\subsubsection{Special cases}
-
-However, the relationship between the distance between an earthquake and its tsunami and the wave height is not necessarily linear, especially if it is considered that the travel of sound (which is a wave, or the displacement of a fluid that can travel) in a fluid obeys the inverse square law.
-Therefore, a simple linear regression model cannot be used for this research.
-A second degree polynomial regression formula would be more appropriate, and a new $x^2$ term can be added to the previous formula.
-We get:
-
-$y = a + \beta_1 x + \beta_2 x^2 + e$
-
-Even though the right side of this equation is quadratic in $x$, this is still linear regression as the right side is linear in the $\beta$ terms (if we were to let $x = 1$, we would be left with $y = a + \beta_1 + \beta_2 + e$, which is linear).
-A second degree polynomial regression would best fit our model for both independent variables, as seen in the graphs below, where a theoretical curve of regression is drawn by hand:
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.8\textwidth]{modelshowcase.jpeg}
-    \label{fig:boat1}
-\end{figure}
-
-In addition, our dataset has two independent variables: the distance from a particular earthquake event and the magnitude of that earthquake.
-Models such as the one explored in this paper, where multiple independent variables are present, are called multivariate and can be modeled by adding the terms of a new independent variable, as such for a linear model with two independent variables:
-
-$a_1 + \beta_1 x_1 + a_2 + \beta_2 x_2 + e$,
-
-where the terms $a_n$ are the y-intercepts of the $n$th independent variable,
-
-So the relationship between the two independent variables and the wave height can be estimated with a multivariate, polynomial regression model.
-We will use the Python scikit-learn library to aid in creating this model.
-\\
-
-(\textit{Section not done, need to do more research on how the} \verb|PolynomialFeatures| \textit{function transforms a linear array to one fit for polynomial regression})
-
-\begin{figure}[H]
-    \centering
-    \subfloat[\centering Linear model]{{\includegraphics[width=7.5cm]{linear.png} }}
-    \qquad
-    \subfloat[\centering Polynomial model]{{\includegraphics[width=7.5cm]{poly.png} }}
-    \caption{2 multivariate (hence the 3 dimensions) regression models}
-    \label{fig:example}
-\end{figure}
-
-\subsubsection{Calculating fit of the model}
-
-The model's fit is then calculated, with the test dataset.
-The \verb|accuracy_score| method from the \verb|sklearn.metrics| class was used for this task.
-This function calculates what is called the $R^2$ value, which explains the extent at which the variance (the amount by which numbers are spread from the average value) in the dependent variable can be explained by the independent variable(s).
-An $R^2$ value of 0.6 indicates that 60\% of the model's results match those of the dataset.
-This value is defined as:
-
-$R^2 = 1 - \frac{SS_{res}}{SS_{total}}$,
-
-where $SS_{res}$ is the sum of the error terms, or the residual sum of squares, in a dataset of $n$
-values $y_1,\dots, y_i$, with their corresponding predicted values $p_1,\dots, p_i$ defined by
-
-$SS_{res} = \sum\limits_{i} e_i^2 = \sum\limits_{i} (y_i - p_i)^2$, \footnote{$\sum\limits_{i=1}^n x_i$ means that all values of x are summed, beginning from $x_1$ and ending at $x_n$} where $e_i$ is the error term, equal to the difference of a data point $y_i$ and its corresponding predicted value $p_i$.
-
-and $SS_{tot}$ is the total sum of squares, which is proportional to the variance in the data, defined as:
-
-$SS_{tot} = \sum\limits_{i} (y_i - \bar y)^2$, where $\bar y$ is the mean of the data $y_1,\dots, y_i$,
-or $\bar y = \frac{1}{n} \sum\limits_{i=1}^n y_i$
+\subsection{Economic sustainability}
+Lorem ipsum dolor sit amet
 
@@
 
-The research included in this paper could be useful for many life-saving scenarios,
-notably ones in which, for example, high quality and large-scale tsunami alert
-systems are not available.
-
-\textit{Conclusion and evaluation of the method are contradictory, to be corrected}
-
-\subsection{Evaluation of the method used}
-
-This method, like any other statistical approximation, is by no means perfect and for that reason should not be used in any real-world application.
-The results procured by the program are subject to many uncertainties, one of which is described below.
-
-When attempting to predict the maximum wave height of a tsunami such as the one caused by the 2011 earthquake off the Pacific coast of Tōhoku, approximately 70km away from shore, which produced a 38.9m wave \cite{yomiuri_2011}.
-The model in this paper approximates this same event as only a 8m wave, when supplied the 9.0 magnitude of the earthquake and the 70km distance.
-These types of errors can be attributed to the fact that there are many more independent variables, such as the shape of the seafloor for example, that can amplify or potentially change the height of a tsunami's wave.
+Lorem ipsum dolor sit amet
 
@@
 \label{app}
-\subsection{Example linear regression}
-\label{app:regression}
 
-Below is an example of a dataset with $x$ values and their corresponding Y values, the data points chosen are purely arbitrary.
-
-\begin{center}
-    \begin{tabular}{||c c||}
-        \hline
-        X values & Y values \\ [0.5ex]
-        \hline\hline
-        368      & 1700     \\
-        \hline
-        340      & 1500     \\
-        \hline
-        665      & 2800     \\
-        \hline
-        954      & 5000     \\
-        \hline
-        331      & 1300     \\
-        \hline
-        556      & 2200     \\
-        \hline
-        376      & 1300     \\ [1ex]
-        \hline
-    \end{tabular}
-\end{center}
-
-\pgfplotstableread{
-    X Y
-    368 1700
-    340 1500
-    665 2800
-    954 5000
-    331 1300
-    556 2200
-    376 1300
-}\table
-
-Below is the dataset plotted, along with a line of regression.
-
-\begin{center}
-    \begin{tikzpicture}[scale=1.1]
-        \begin{axis}[
-                title={Comparing X and Y},
-                xlabel={X axis},
-                ylabel={Y axis},
-                xmin=300, xmax=1000,
-                ymin=0, ymax=5500,
-                ytick={0,1000,2000,3000,4000,5000},
-                xtick={0,400,500,600,700,800,900},
-                legend pos=north west,
-                ymajorgrids=true,
-                grid style=dashed,
-                scale=1.2
-            ]
-
-            \addplot[
-                only marks,
-                color=blue,
-                mark=*,
-            ] table {\table};
-            \legend{Datapoint}
-
-            \addplot [thin, blue] table[
-                    y={create col/linear regression={y=Y}}
-                ] % compute a linear regression from the input table
-                {\table};
-            \addlegendentry{
-                Linear regression: $ x =
-                    \pgfmathprintnumber{\pgfplotstableregressiona}
-                    \cdot t
-                    \pgfmathprintnumber[print sign]{\pgfplotstableregressionb}$
-            };
-
-        \end{axis}
-    \end{tikzpicture}
-\end{center}
-
-\subsection{Python program}
-\label{app:scripts}
-
-Listed here is the Python program through which the dataset was cleaned and predictions based on the statistical model were made. Run using 64-bit Python version 3.9.1.
-
-\lstinputlisting[language=Python]{model/multregr2.py}
+\subsection{Appendix Test}
+\label{app:test}
 

### 2021-05-13 geography/ia/paper.tex
geo ia added

@@
 
-Lorem ipsum dolor sit amet
+For our investigation, the topic in question is the environmental quality.
+We shall compare the environmental quality of two areas, one meant as a residential one and one with a heavy tourist presence.
+To best represent these areas, we have chosen the XVIe and the VIIe, justifiable as the XVIe is home to many housing complexes and fosters facilities aimed at catering to the residents whereas the VIIe sees many tourists as it is home to the famed Eiffel Tower and the Seine river, prime tourist attractions of Paris.
 
@@
 
-Lorem ipsum dolor sit amet
+The method we will be using to select the sites from where we will take the Bipolar survey is Stratified which means that we are choosing the sites ourselves.
+The reason we chose this is to make sure that we got a good variety of sites that would give a good idea of the general environmental quality of that area, but also to avoid the selection of residential areas, which do in fact exist in the VIIe, which would happen if we chose the random sampling method.
+This stratified sampling approach would best represent the targeted area and yield meaningful data on tourist-heavy sites.
+
+The sampling method we used to select our sites is the stratified method. The reason for this choice is that we get a
 
@@
 
-Lorem ipsum dolor sit amet
+To best retrieve data from our sites, we shall maintain a fair test to uphold the standard and quality of our reliability by getting our data at around the same time on the same day of the week, in an attempt to best match up with the tourists’ habits.
+We have chosen the time frame of 14:00 to 16:00 on a Friday afternoon, a time frame where tourists will not be constrained by lunch or sleep.
 
@@
 
-All figures are generated through the use of TikZ (Tantau) and the Python library \verb|matplotlib|.
-This paper is written with the aid of the typesetting software \LaTeX.
-On a PDF viewer, the sections in the table of contents can be clicked to access that section.
-Dataset not included due to large size (\textasciitilde 28000 columns).
+This paper is written with the aid of the typesetting software \LaTeX. On a PDF viewer, the sections in the table of contents can be clicked to access that section.
 

### 2021-05-14 geography/ia/paper.tex
geo ia

@@
         \vspace*{4cm}
-        Geography IA \\
+        Geography\\
+        Internal Assessment \\
         \vspace{1cm}
-        Can the height of a tsunami's wave be predicted based on the distance from and
-        the magnitude of a preceding earthquake? \\
+        How does environmental quality, determined by the severity of pollution and the availability of green spaces, differ between a tourist-catering area of Paris versus a residential area? \\
         \vspace{1cm}
         \textit{Väinö-verneri Kauppila} \\
-        February 2021 \\
+        May 2021 \\
         \vspace{4cm}
         Word count: TBD \\
-        \textit{First draft, details are missing}
         \vfill
@@
 
-Lorem ipsum dolor sit amet
+\subsection{Hypotheses}
 
-\subsection{Study area}
+\begin{enumerate}
+    \item According to the Global Development Goals of the UN, 90\% of urban areas in the world had polluted air in 2016.
+\footcite{sdg_report_2020} Paris was among these countries that didn't satisfy WHO's air quality minimum \footcite{ambient_outdoor_air_pollution_2018} of 2018, with on average a 50\% higher than normal pollution density.
+In addition, the areas of tourism in Paris are, as seen in Figure 1, higher in pollution than residential areas.
+However, a counterargument to this could be that despite a higher concentration of people, tourist areas in Paris do not suffer as much from high traffic conditions from things such as typical morning rush hours, and tourists preferably using public transport or bikes.
 
-Lorem ipsum dolor sit amet
+    \item 2nd hypothesis
+    \item 3rd hypothesis
+\end{enumerate}
 
-\subsection{Theory}
-
-Lorem ipsum dolor sit amet
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.7\linewidth]{no2_map.png}
+    \caption{A map of Paris NO2 levels in 2018, as reported by AirParif, on the official Paris website \footcite{paris_air_qual}}
+\end{figure}
 
@@
 
-Lorem ipsum dolor sit amet
+Lorem ipsum dolor si amet
 
@@
 
-\printbibliography[heading=bibintoc, title=Works Cited]
+\printbibliography
 

### 2021-05-14 geography/ia/paper.tex
update geo ia. text & figures

@@
 
+\subsection{Fieldwork question}
+
+How does environmental quality, determined by the severity of pollution and the availability of green spaces, differ between a touristic area of Paris and a residential area?
+
+The fieldwork for our paper was conducted in Paris, one of the largest and most well known cities in the world.
+The city itself is arranged in a manner such that 20 districts, known locally as \textit{arrondissements}, are placed in a spiral in the city.
+It is globally known for its high number of tourists per year, equating to around 35 million tourists in the year 2019 alone.
+\footcite{statista_department_27_2020} Many of them come to visit the world-renowned Eiffel Tower, located in the VII\textsuperscript{th} \textit{arrondissement}.
+\footcite{condor_ferries} To add, Paris is home to 2.2 million people, who mostly live in the outer residential areas, from the XI\textsuperscript{th} to the XX\textsuperscript{th} \textit{arrondissements}.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.7\linewidth]{arrts.png}
+    \caption{A map of Paris \textit{arrondissements}, drawn by hand.}
+\end{figure}
+
 \subsection{Hypotheses}
@@
 \footcite{sdg_report_2020} Paris was among these countries that didn't satisfy WHO's air quality minimum \footcite{ambient_outdoor_air_pollution_2018} of 2018, with on average a 50\% higher than normal pollution density.
-In addition, the areas of tourism in Paris are, as seen in Figure 1, higher in pollution than residential areas.
+In addition, the areas of tourism in Paris are, as seen in Figure 2, higher in pollution than residential areas.
 However, a counterargument to this could be that despite a higher concentration of people, tourist areas in Paris do not suffer as much from high traffic conditions from things such as typical morning rush hours, and tourists preferably using public transport or bikes.
 
-    \item 2nd hypothesis
-    \item 3rd hypothesis
+    \item As there are more people moving about in residential areas, for example in cars for the morning commute or at noon for lunch, it can logically be theorized that noise pollution, which is obviously a function of the amount of people, would be higher in these places.
+Today it is estimated that an average noise level of $60dB$ can be found in residential areas, according to the comprehensive Bruitparif government-sponsored report.
+\footcite{bruitparif} This value largely surpasses the WHO's safe level of $53dB$.
+\footcite{who_noise_guidelines}
+
+    \item The Parisian mayor, Anne Hidalgo included the betterment of the environment in her campaign.
+The mayor has promised to make so-called "green spaces" no further than 200 meters to any person,\footcite{anne_hidalgo_2020} and as such, it should be hypothesized that green spaces, which include parks, agglomerations of trees, shall be distributed evenly with no difference between residential and touristic areas.
+The mayor emphasized on "urban forests" --- places where residents and tourists alike could enjoy the company of trees while walking along the city streets.
 \end{enumerate}
@@
 \begin{figure}[h!]
+    \begin{minipage}{\textwidth}
     \centering
     \includegraphics[width=0.7\linewidth]{no2_map.png}
-    \caption{A map of Paris NO2 levels in 2018, as reported by AirParif, on the official Paris website \footcite{paris_air_qual}}
+    \caption{A map of Paris NO2 levels in 2018, as reported by AirParif, on the official Paris website \protect\footcite{paris_air_qual}}
+    \end{minipage}
 \end{figure}
@@
 
+\subsection{Study site choices}
+
+We chose 10 sites in total to study, shown below in Figures 3, 4.
+
+\begin{figure}[H]
+    \begin{minipage}{\textwidth}
+    \centering
+    \includegraphics[width=0.7\linewidth]{no2_map.png}
+    \caption{Map of sites chosen for the residential area, the XVI\textsuperscript{th} \textit{arrondissement}. Map base layer courtesy of Google Maps, pictures seen are taken on a mobile phone.}
+    \end{minipage}
+\end{figure}
+
+\begin{figure}[H]
+    \begin{minipage}{\textwidth}
+    \centering
+    \includegraphics[width=0.7\linewidth]{no2_map.png}
+    \caption{Map of sites chosen for the residential area, the XVI\textsuperscript{th} \textit{arrondissement}. Map base layer courtesy of Google Maps, pictures seen are taken on a mobile phone.}
+    \end{minipage}
+\end{figure}
+
 \subsection{Sampling choices}
@@
 The sampling method we used to select our sites is the stratified method. The reason for this choice is that we get a
-
-\subsection{Study site choices}
-
-To best retrieve data from our sites, we shall maintain a fair test to uphold the standard and quality of our reliability by getting our data at around the same time on the same day of the week, in an attempt to best match up with the tourists’ habits.
-We have chosen the time frame of 14:00 to 16:00 on a Friday afternoon, a time frame where tourists will not be constrained by lunch or sleep.
 

### 2021-05-15 geography/ia/paper.tex
added ecole militaire picture
geo ia work, refactoring, math ia fixes

@@
     \centering
-    \includegraphics[width=0.7\linewidth]{arrts.png}
+    \includegraphics[width=0.7\linewidth]{media/arrts.png}
     \caption{A map of Paris \textit{arrondissements}, drawn by hand.}
@@
     \item As there are more people moving about in residential areas, for example in cars for the morning commute or at noon for lunch, it can logically be theorized that noise pollution, which is obviously a function of the amount of people, would be higher in these places.
-Today it is estimated that an average noise level of $60dB$ can be found in residential areas, according to the comprehensive Bruitparif government-sponsored report.
+Today it is estimated that an average noise level of $60dB$ can be found in residential areas, according to the very comprehensive Bruitparif government-sponsored report.
 \footcite{bruitparif} This value largely surpasses the WHO's safe level of $53dB$.
@@
     \centering
-    \includegraphics[width=0.7\linewidth]{no2_map.png}
+    \includegraphics[width=0.7\linewidth]{media/no2_map.png}
     \caption{A map of Paris NO2 levels in 2018, as reported by AirParif, on the official Paris website \protect\footcite{paris_air_qual}}
@@
 \end{figure}
+
+% NEW SECTION
 
@@
 
-We chose 10 sites in total to study, shown below in Figures 3, 4.
+We chose 10 sites in total to conduct a bipolar survey, shown below in Figures 3, 4.
 
@@
     \centering
-    \includegraphics[width=0.7\linewidth]{no2_map.png}
+    \includegraphics[width=0.7\linewidth]{media/16esites.png}
     \caption{Map of sites chosen for the residential area, the XVI\textsuperscript{th} \textit{arrondissement}. Map base layer courtesy of Google Maps, pictures seen are taken on a mobile phone.}
@@
 
+This site was chosen as, like stated before, it is a residential area. There are other residential areas, however this one was specifically chosen for reasons of convenience.
+
 \begin{figure}[H]
@@
     \centering
-    \includegraphics[width=0.7\linewidth]{no2_map.png}
+    \includegraphics[width=0.7\linewidth]{media/7esites.png}
     \caption{Map of sites chosen for the residential area, the XVI\textsuperscript{th} \textit{arrondissement}. Map base layer courtesy of Google Maps, pictures seen are taken on a mobile phone.}
@@
 
-\subsection{Sampling choices}
+This area in the VII\textsuperscript{th} was chosen as it is a staple of tourism in France. It is home to the famed Eiffel Tower, the Ecole Militaire building and the Seine river bank.
 
-The method we will be using to select the sites from where we will take the Bipolar survey is Stratified which means that we are choosing the sites ourselves.
-The reason we chose this is to make sure that we got a good variety of sites that would give a good idea of the general environmental quality of that area, but also to avoid the selection of residential areas, which do in fact exist in the VIIe, which would happen if we chose the random sampling method.
-This stratified sampling approach would best represent the targeted area and yield meaningful data on tourist-heavy sites.
+\subsection{Our sampling method}
 
-The sampling method we used to select our sites is the stratified method. The reason for this choice is that we get a
+The sites in the two areas were chosen using a method of stratified sampling, as seeing that the VVI\textsuperscript{th} does indeed contain residential sites and the XVI\textsuperscript{th} touristic ones, we can discard the ones that are irrelevant to our study.
+
+\subsection{Method}
+
+\subsubsection{Bipolar semantic survey}
+
+The sites will be visited to conduct a so-called "bipolar survey", which consists of rating the site based on multiple criteria.
+This process was done by selecting a single person from our group to visit one of the sites, and complete a bipolar semantic survey.
+They would grade several different aspects of the site (such as general cleanliness, noise, amount of cars), and take a set of 3 images of the site.\footnote{Images taken for our study can be found at \url{https://github.com/kinnounko/notes/tree/main/geography/ia/images}}
+
+\subsubsection{Statistical evaluations}
+
+In order to either support or disprove our third hypothesis, it is important to study how trees are scattered in Paris, which reveals how dispersed green spaces are from one another.
+A statistical test called the Nearest Neighbor Index (NNI) test can provide a reasonable answer to this: if the index is similar for both sites, it can easily be said then that our hypothesis is supported.
+This value also reveals information about the distribution of trees: clustered, random or in a regular pattern, as seen in Figure 5 below.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.2\linewidth]{media/nni1.png}
+    \includegraphics[width=0.2\linewidth]{media/nni2.png}
+    \includegraphics[width=0.2\linewidth]{media/nni3.png}
+    \caption{The NNI measures the spacial distribution of data. From a value of 0, representing a clustered pattern, to 1 (random) and to 2.15 (uniform pattern)}
+\end{figure}
+
+As individually counting trees in these areas would be a tedious task, we used the Paris OpenData platform (licensed under the permissive ODbL license\footcite{odbl}), owned by the government.
+The fact that this kind of database platform exists, shows a certain level of transparency on the government's part.
+
+We used the trees dataset procured from this library, which contains the exact coordinates of trees in the city.
+\footcite{paris_opendata} There are also other information such as species of the tree and such, however these are cleaned from the dataset for our purposes.
+
+A script, written in Python was programmed for the task of calculating the NNI value, because with a dataset of over 20,000 data points this would not be possible by hand.
+We go through each point, and find its nearest neighbor.
+In order to calculate distances between these two points, we use the Haversine Formula, as the two points would be in the form of coordinates.
+With this information the NNI can be calculated with the simple formula
+
+$$Rn = \frac{2 \bar D}{\sqrt{\frac{a}{n}}}$$
+
+where $Rn$ is the NNI index value, $\bar D$ is the mean observed distance to the nearest neighbor, $a$ is the area of the zone and $n$ is the total number of data points.
+
+% NEW SECTION
 
@@
 
+All media and information related to this IA can be found on the URL \url{https://github.com/kinnounko/notes/tree/main/geography/ia}.
+This includes images, the \LaTeX{} source code to this paper, and the Python program used to calculate the NNI value.
+
 \printbibliography

### 2021-05-15 math/ia_2021/paper.tex
geo ia
geo ia work, refactoring, math ia fixes

@@
 
-Naturally occurring disasters, such as tsunamis, are devastating for people, human settlements,
-buildings and other resources residing in risk areas. Tsunamis are caused
-by many different types of natural phenomena, such as landslides, collapse of
-seamount, or more rarely, the impact of a meteorite. This paper, however, focuses
-on one of the more common causes of tsunamis; earthquakes. Earthquakes cause around 72\%
-of tsunamis \cite{pacifictsunamimuseum}. The time between an earthquake and its tsunami is
-relatively long compared to events such as landslides \cite{sue_nokes_walters},
-leaving adequate time to prepare for impact. Predicting wave height, as discussed in this
-paper, can further help to assess the seriousness of the tsunami and provide help to
-prepare for the event.
+Naturally occurring disasters, such as tsunamis, are devastating for people, human settlements, buildings and other resources residing in risk areas.
+Tsunamis are caused by many different types of natural phenomena, such as landslides, collapse of seamount, or more rarely, the impact of a meteorite.
+This paper, however, focuses on one of the more common causes of tsunamis; earthquakes.
+Earthquakes cause around 72\% of tsunamis \cite{pacifictsunamimuseum}.
+The time between an earthquake and its tsunami is relatively long compared to events such as landslides \cite{sue_nokes_walters}, leaving adequate time to prepare for impact.
+Predicting wave height, as discussed in this paper, can further help to assess the seriousness of the tsunami and provide help to prepare for the event.
 
@@
 
-I have chosen to write about this subject as I have recently become interested by the
-statistics and its implications in real-life matters. This research has proved very
-interesting for me as it has made the field of statistics more concrete, because it provides
-a real world use case. Furthermore, as I enjoy programming, the research I have done
-has helped me learn about the tools and techniques used by data analysts and other professionals.
+I have chosen to write about this subject as I have recently become interested by the statistics and its implications in real-life matters.
+This research has proved very interesting for me as it has made the field of statistics more concrete, because it provides a real world use case.
+Furthermore, as I enjoy programming, the research I have done has helped me learn about the tools and techniques used by data analysts and other professionals.
 
@@
 \subsection{Processing the dataset for use}
-%TODO: #2 Ask for help 
-The dataset, procured from the National Oceanic and Atmospheric Administration website \cite{noaa}, includes many tsunami
-runups, going back before the year 0. Due to this, the NOAA states uncertainties
-that can occur in its data. For example, reports that were recorded before the
-20th century tend to be based on written accounts, and not recorded seismic activity.
-Only since the 1960s, the installation of the Worldwide Standardized Seismograph
-Network system allowed for more accurate data in this paper. For that reason, any recorded
-activity before this date was discarded from the dataset for this paper. The dataset itself can
-be accessed through two different methods; searching the database using the online
-tool or downloading it in a somewhat obscure compressed Google Earth \verb|.kmz| format.
-The online tool supports the download of search results in a much more human-readable
-\verb|.tsv| (tab-separated) format. For this paper, all entries after the year 1900 were searched and downloaded.
 
-The data is cleaned of data points marked as "dubious" (there is an column for this parameter),
-reports before the year 1960 (see above), and of all irrelevant fields such as "Injuries"
-or "Damage \$Mil". This preliminary data cleanup
-is done through the use of the pandas Python library \cite{reback2020pandas}\cite{mckinney-proc-scipy-2010},
-which is commonly used for the handling of large datasets for its efficiency.
+The dataset, procured from the National Oceanic and Atmospheric Administration website \cite{noaa}, includes many tsunami runups, going back before the year 0.
+Due to this, the NOAA states uncertainties that can occur in its data.
+For example, reports that were recorded before the 20th century tend to be based on written accounts, and not recorded seismic activity.
+Only since the 1960s, the installation of the Worldwide Standardized Seismograph Network system allowed for more accurate data in this paper.
+For that reason, any recorded activity before this date was discarded from the dataset for this paper.
+The dataset itself can be accessed through two different methods; searching the database using the online tool or downloading it in a somewhat obscure compressed Google Earth \verb|.kmz| format.
+The online tool supports the download of search results in a much more human-readable \verb|.tsv| (tab-separated) format.
+For this paper, all entries after the year 1900 were searched and downloaded.
 
-The data was split into a training and a test dataset. The training
-dataset was used to generate a regression model while the test dataset was used
-to calculate the general accuracy of the model, by predicting values and
-comparing those to the known value as found in the dataset. The dataset was split based on
-a fraction using the \verb|df.sample| method found in the pandas library. A split
-of 70\% for training and 30\% for test was chosen. A \verb|random_state| argument is
-set to an arbitrary value 42, as if it weren't, the dataset would be split in a different
-manner at each execution of the program. The program can be found in Appendix \ref{app:scripts}.
+The data is cleaned of data points marked as "dubious" (there is an column for this parameter), reports before the year 1960 (see above), and of all irrelevant fields such as "Injuries" or "Damage \$Mil".
+This preliminary data cleanup is done through the use of the pandas Python library \cite{reback2020pandas}\cite{mckinney-proc-scipy-2010}, which is commonly used for the handling of large datasets for its efficiency.
+
+The data was split into a training and a test dataset.
+The training dataset was used to generate a regression model while the test dataset was used to calculate the general accuracy of the model, by predicting values and comparing those to the known value as found in the dataset.
+The dataset was split based on a fraction using the \verb|df.sample| method found in the pandas library.
+A split of 70\% for training and 30\% for test was chosen.
+A \verb|random_state| argument is set to an arbitrary value 42, as if it weren't, the dataset would be split in a different manner at each execution of the program.
+The program can be found in Appendix \ref{app:scripts}.
 
@@
 
-Regression analysis is a commonly used tool in statistics, and regression models can
-be found in numerous use cases. Regression models consist of estimating the
-relationship between one (or more) independent variables and a dependent variable. A
-simple example of regression can be seen in the linear relationship between an independent
-variable $x_{i}$, and the dependent variable $y$. The following expression shows this relationship in what is
-called linear regression, as it models a straight line that can be used for prediction of
-values:
+Regression analysis is a commonly used tool in statistics, and regression models can be found in numerous use cases.
+Regression models consist of estimating the relationship between one (or more) independent variables and a dependent variable.
+A simple example of regression can be seen in the linear relationship between an independent variable $x_{i}$, and the dependent variable $y$.
+The following expression shows this relationship in what is called linear regression, as it models a straight line that can be used for prediction of values:
 
@@
 
-where the term $\beta_0$ represents the y-intercept of the line and $\beta$ represents the
-coefficient of $x$. The final term $e$ is what is known as the error or residual
-term. It is an estimate of the amount by which the predictions differ from the real
-value. An example of linear regression can be found in Appendix \ref{app:regression}, where
-a set of data points is presented on a scatter plot, and the regression line is found
-by finding the individual coefficients $\beta_n$ of $x$ (see equation above), like such:
+where the term $\beta_0$ represents the y-intercept of the line and $\beta$ represents the coefficient of $x$.
+The final term $e$ is what is known as the error or residual term.
+It is an estimate of the amount by which the predictions differ from the real value.
+An example of linear regression can be found in Appendix \ref{app:regression}, where a set of data points is presented on a scatter plot, and the regression line is found by finding the individual coefficients $\beta_n$ of $x$ (see equation above), like such:
 
-$\hat \beta_1 = \frac{\sum (x_i - \bar x) (y_i - \bar y)}{\sum (x_i - \bar x)^2}$, where
-$\hat \beta_n$ represents an estimated value and $\bar x$ represents the mean of the set $x$
+$\hat \beta_1 = \frac{\sum (x_i - \bar x) (y_i - \bar y)}{\sum (x_i - \bar x)^2}$, where $\hat \beta_n$ represents an estimated value and $\bar x$ represents the mean of the set $x$
 
@@
 
-where $SS_{res}$ is the sum of the error terms, or the residual sum of squares, in a dataset of $n$
-values $y_1,\dots, y_i$, with their corresponding predicted values $p_1,\dots, p_i$ defined by
+where $SS_{res}$ is the sum of the error terms, or the residual sum of squares, in a dataset of $n$ values $y_1,\dots, y_i$, with their corresponding predicted values $p_1,\dots, p_i$ defined by
 
-$SS_{res} = \sum\limits_{i} e_i^2 = \sum\limits_{i} (y_i - p_i)^2$, \footnote{$\sum\limits_{i=1}^n x_i$ means that all values of x are summed, beginning from $x_1$ and ending at $x_n$} where $e_i$ is the error term, equal to the difference of a data point $y_i$ and its corresponding predicted value $p_i$.
+$SS_{res} = \sum\limits_{i} e_i^2 = \sum\limits_{i} (y_i - p_i)^2$, \footnote{$\sum\limits_{i=1}^n x_i$ means that all values of $x$ are summed, beginning from $x_1$ and ending at $x_n$} where $e_i$ is the error term, equal to the difference between a data point $y_i$ and its corresponding predicted value $p_i$.
 
@@
 
-$SS_{tot} = \sum\limits_{i} (y_i - \bar y)^2$, where $\bar y$ is the mean of the data $y_1,\dots, y_i$,
-or $\bar y = \frac{1}{n} \sum\limits_{i=1}^n y_i$
+$SS_{tot} = \sum\limits_{i} (y_i - \bar y)^2$, where $\bar y$ is the mean of the data $y_1,\dots, y_i$, or $\bar y = \frac{1}{n} \sum\limits_{i=1}^n y_i$
 
@@
 
-The research included in this paper could be useful for many life-saving scenarios,
-notably ones in which, for example, high quality and large-scale tsunami alert
-systems are not available.
+The research included in this paper could be useful for many life-saving scenarios, notably ones in which, for example, high quality and large-scale tsunami alert systems are not available.
 
@@
 
-When attempting to predict the maximum wave height of a tsunami such as the one caused by the 2011 earthquake off the Pacific coast of Tōhoku, approximately 70km away from shore, which produced a 38.9m wave \cite{yomiuri_2011}.
+When attempting to predict the maximum wave height of a tsunami such as the one caused by the 2011 earthquake off the Pacific coast of Tōhoku, Japan, approximately 70km away from shore, which produced a 38.9m wave \cite{yomiuri_2011}.
 The model in this paper approximates this same event as only a 8m wave, when supplied the 9.0 magnitude of the earthquake and the 70km distance.
 These types of errors can be attributed to the fact that there are many more independent variables, such as the shape of the seafloor for example, that can amplify or potentially change the height of a tsunami's wave.
+
+It has also come to my attention that using the $R^2$ value for a non-linear relationship is not suitable, as it does not follow the same characteristics.
 

### 2021-05-15 geography/ia/paper.tex
bipolar semantic survey tallies + charts etc

@@
 
-The fieldwork for our paper was conducted in Paris, one of the largest and most well known cities in the world.
+In order to answer this question, the fieldwork for our paper was conducted in Paris, one of the largest and most well known cities in the world.
 The city itself is arranged in a manner such that 20 districts, known locally as \textit{arrondissements}, are placed in a spiral in the city.
@@
 
-\subsection{Our sampling method}
+\subsubsection{Our sampling method}
 
-The sites in the two areas were chosen using a method of stratified sampling, as seeing that the VVI\textsuperscript{th} does indeed contain residential sites and the XVI\textsuperscript{th} touristic ones, we can discard the ones that are irrelevant to our study.
+The sites in the two areas were chosen using a method of stratified sampling, as seeing that the VII\textsuperscript{th} does indeed contain residential sites and the XVI\textsuperscript{th} touristic ones, we can discard the ones that are irrelevant to our study.
 
@@
 This process was done by selecting a single person from our group to visit one of the sites, and complete a bipolar semantic survey.
-They would grade several different aspects of the site (such as general cleanliness, noise, amount of cars), and take a set of 3 images of the site.\footnote{Images taken for our study can be found at \url{https://github.com/kinnounko/notes/tree/main/geography/ia/images}}
+They would grade several different aspects of the site (such as general cleanliness, noise, amount of cars), and take a set of 3 images of the site.\footnote{Images taken for our study can be found at \url{https://github.com/kinnounko/notes/tree/main/geography/ia/images}} Our surveys were conducted during the week of the 10th of March, 2021 in the afternoon each time.
+The weather was overcast and moderately cold.
+
+Our results for the bipolar survey are shown in Appendix \ref{app:bipolar}
+
+When tallied, the total scores for each site can be expressed as a fraction over 84, as shown in Figures 5, 6.
+
+\begin{figure}
+    \begin{center}
+        \begin{tabular}{||c c||}
+            \hline
+            Site & Total score  \\ [0.5ex]
+            \hline\hline
+            Bir-Hakeim                 & 51    \\
+            \hline
+            Seine river bank           & 69    \\
+            \hline
+            Foot of the Eiffel Tower   & 45    \\
+            \hline
+            Iena Bridge                & 42    \\
+            \hline
+            Ecole Militaire            & 54    \\ [1ex]
+            \hline
+        \end{tabular}
+    \end{center}
+    \caption{Total Bipolar Semantic scores for the VII\textsuperscript{th}}
+\end{figure}
+
+\begin{figure}
+    \begin{center}
+        \begin{tabular}{||c c||}
+            \hline
+            Site & Total score  \\ [0.5ex]
+            \hline\hline
+            Rue Davioud 1       & 47    \\
+            \hline
+            Rue Davioud 2       & 53    \\
+            \hline
+            Avenue Mozart       & 31    \\
+            \hline
+            Avenue Paul Doumer  & 35    \\
+            \hline
+            Ranelagh Garden     & 77    \\ [1ex]
+            \hline
+        \end{tabular}
+    \end{center}
+    \caption{Total Bipolar Semantic scores for the XVI\textsuperscript{th}}
+\end{figure}
+
+When plotted on a map, these totals can be shown with a gradient of colors.
 
@@
 
-\subsection{Appendix Test}
-\label{app:test}
+\subsection{Bipolar survey results}
+\label{app:bipolar}
+
+Shown below are our results from our bipolar semantic survey:
+
+XVI\textsuperscript{th} residential area:
+
+\includegraphics[width=0.5\linewidth]{media/bipolar/mozart.png}
+\includegraphics[width=0.5\linewidth]{media/bipolar/davioud1.png}
+\includegraphics[width=0.5\linewidth]{media/bipolar/davioud2.png}
+\includegraphics[width=0.5\linewidth]{media/bipolar/ranelagh.png}
+\includegraphics[width=0.5\linewidth]{media/bipolar/pauldoumer.png}
+
+VII\textsuperscript{th} touristic area:
+
+\includegraphics[width=0.5\linewidth]{media/bipolar/birhakeim.png}
+\includegraphics[width=0.5\linewidth]{media/bipolar/ecolemilitaire.png}
+\includegraphics[width=0.5\linewidth]{media/bipolar/eiffel.png}
+\includegraphics[width=0.5\linewidth]{media/bipolar/iena.png}
+\includegraphics[width=0.5\linewidth]{media/bipolar/seinebank.png}
 

### 2021-05-16 geography/ia/paper.tex
geo work, ia sections a&b done

@@
         \vspace{1cm}
-        \textit{Väinö-verneri Kauppila} \\
+        \textit{Väinö-Verneri Kauppila} \\
         May 2021 \\
         \vspace{4cm}
-        Word count: TBD \\
+        Word count: \textit{TBD} \\
         \vfill
@@
 In order to answer this question, the fieldwork for our paper was conducted in Paris, one of the largest and most well known cities in the world.
-The city itself is arranged in a manner such that 20 districts, known locally as \textit{arrondissements}, are placed in a spiral in the city.
-It is globally known for its high number of tourists per year, equating to around 35 million tourists in the year 2019 alone.
+The city itself is arranged into 20 districts, known locally as \textit{arrondissements}, are placed in a spiral in the city.
+It is globally known for its high number of tourists per year, equating to around 35 million in the year 2019 alone.
 \footcite{statista_department_27_2020} Many of them come to visit the world-renowned Eiffel Tower, located in the VII\textsuperscript{th} \textit{arrondissement}.
@@
     \item According to the Global Development Goals of the UN, 90\% of urban areas in the world had polluted air in 2016.
-\footcite{sdg_report_2020} Paris was among these countries that didn't satisfy WHO's air quality minimum \footcite{ambient_outdoor_air_pollution_2018} of 2018, with on average a 50\% higher than normal pollution density.
-In addition, the areas of tourism in Paris are, as seen in Figure 2, higher in pollution than residential areas.
+\footcite{sdg_report_2020} Paris was among countries that didn't satisfy WHO's air quality minimum \footcite{ambient_outdoor_air_pollution_2018} of 2018, with on average a 50\% higher than normal pollution density.
 However, a counterargument to this could be that despite a higher concentration of people, tourist areas in Paris do not suffer as much from high traffic conditions from things such as typical morning rush hours, and tourists preferably using public transport or bikes.
 
-    \item As there are more people moving about in residential areas, for example in cars for the morning commute or at noon for lunch, it can logically be theorized that noise pollution, which is obviously a function of the amount of people, would be higher in these places.
+    \item As there are more people moving about in residential areas, for example in cars for the morning commute or at noon for lunch, it can logically be theorized that noise pollution, which is obviously a function of the amount of cars, would be higher in these places.
 Today it is estimated that an average noise level of $60dB$ can be found in residential areas, according to the very comprehensive Bruitparif government-sponsored report.
@@
 
-    \item The Parisian mayor, Anne Hidalgo included the betterment of the environment in her campaign.
+    \item The Parisian mayor, Anne Hidalgo included the improvement of environmental quality in her campaign.
 The mayor has promised to make so-called "green spaces" no further than 200 meters to any person,\footcite{anne_hidalgo_2020} and as such, it should be hypothesized that green spaces, which include parks, agglomerations of trees, shall be distributed evenly with no difference between residential and touristic areas.
@@
 
-\begin{figure}[h!]
-    \begin{minipage}{\textwidth}
-        \centering
-        \includegraphics[width=0.7\linewidth]{media/no2_map.png}
-        \caption{A map of Paris NO2 levels in 2018, as reported by AirParif, on the official Paris website \protect\footcite{paris_air_qual}}
-    \end{minipage}
-\end{figure}
+%\begin{figure}[h!]
+%    \begin{minipage}{\textwidth}
+%        \centering
+%        \includegraphics[width=0.7\linewidth]{media/no2_map.png}
+%        \caption{A map of Paris NO2 levels in 2018, as reported by AirParif, on the official Paris website \protect\footcite{paris_air_qual}}
+%    \end{minipage}
+%\end{figure}
 
@@
 For our investigation, the topic in question is the environmental quality.
-We shall compare the environmental quality of two areas, one meant as a residential one and one with a heavy tourist presence.
-To best represent these areas, we have chosen the XVIe and the VIIe, justifiable as the XVIe is home to many housing complexes and fosters facilities aimed at catering to the residents whereas the VIIe sees many tourists as it is home to the famed Eiffel Tower and the Seine river, prime tourist attractions of Paris.
+We will compare the environmental quality of two areas, one meant as a residential one and one with a heavy tourist presence.
+To best represent these areas, we have chosen the XVIe and the VIIe.
+The XVIe is home to many housing complexes and fosters facilities aimed at catering to the residents whereas the VIIe sees many tourists as it is home to the famed Eiffel Tower and the Seine river, prime tourist attractions of Paris.
 
@@
 
-We chose 10 sites in total to conduct a bipolar survey, shown below in Figures 3, 4.
+We chose 10 sites in total to conduct a bipolar survey, shown below in Figures 2, 3.
 
@@
 
-This site was chosen as, like stated before, it is a residential area. There are other residential areas, however this one was specifically chosen for reasons of convenience.
+This site was chosen as, like stated before, it is a residential area. There are other residential areas in Paris, however this one was specifically chosen for reasons of convenience.
 
@@
 
-The sites in the two areas were chosen using a method of stratified sampling, as seeing that the VII\textsuperscript{th} does indeed contain residential sites and the XVI\textsuperscript{th} touristic ones, we can discard the ones that are irrelevant to our study.
+The sites in the two areas were chosen using a method of stratified sampling.
+Considering the VII\textsuperscript{th} does indeed contain residential sites and the XVI\textsuperscript{th} touristic ones, we can discard the ones that are irrelevant to our study.
 
@@
 
-The sites will be visited to conduct a so-called "bipolar survey", which consists of rating the site based on multiple criteria.
+The sites were visited to conduct a so-called "bipolar survey", which consisted of rating the site based on multiple criteria.
 This process was done by selecting a single person from our group to visit one of the sites, and complete a bipolar semantic survey.
@@
 
-When tallied, the total scores for each site can be expressed as a fraction over 84, as shown in Figures 5, 6.
+When tallied, the total scores for each site can be expressed as a fraction over 84, as shown in Figures 4, 5.
 
-\begin{figure}
+\begin{figure}[H]
     \begin{center}
@@
 
-\begin{figure}
+\begin{figure}[H]
     \begin{center}
@@
 
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.7\linewidth]{media/7eColors.png}
+    \caption{Each site in the VVI\textsuperscript{th}, with its respective total score expressed as a percentage for clarity, and a color assigned based on the gradient in Figure 8.}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.7\linewidth]{media/16eColors.png}
+    \caption{Each site in the XVI\textsuperscript{th}, with a percentage. Color assigned with gradient in Figure 8.}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.4\linewidth]{media/bipolar/gradient.png}
+    \caption{Gradient from red (hex color \#FF0000) for 0\%, to green (hex color \#00FF00) for 100\%. }
+\end{figure}
+
 \subsubsection{Statistical evaluations}
@@
 
-We used the trees dataset procured from this library, which contains the exact coordinates of trees in the city.
+We used the trees dataset procured from the library, which contains the exact coordinates of trees in the city.
 \footcite{paris_opendata} There are also other information such as species of the tree and such, however these are cleaned from the dataset for our purposes.
 
-A script, written in Python was programmed for the task of calculating the NNI value, because with a dataset of over 20,000 data points this would not be possible by hand.
+A script, written in Python was tasked to calculate the NNI value, because with a dataset of over 20,000 data points this would not be possible by hand.
 We go through each point, and find its nearest neighbor.
@@
 
-Lorem
+\textit{Not done}
 
 \subsection{Environmental and cultural sustainability}
-Lorem ipsum dolor sit amet
+\textit{Not done}
 
 \subsection{Economic sustainability}
-Lorem ipsum dolor sit amet
+\textit{Not done}
 
@@
 
-Lorem ipsum dolor si amet
+\textit{Not done}
 

### 2021-05-17 geography/ia/paper.tex
automatic wordcount for vscode. requires extension with id "emeraldwalk.runonsave"

@@
         \vspace{4cm}
-        Word count: \textit{TBD} \\
+        \input{wordcount.txt}
         \vfill
@@
 \section{Introduction}
+\label{sec:intro}
 
@@
 \section{Method}
+\label{sec:method}
 
@@
 
-This site was chosen as, like stated before, it is a residential area. There are other residential areas in Paris, however this one was specifically chosen for reasons of convenience.
+This site was chosen as, like stated before, it is a residential area.
+There are other residential areas in Paris, however this one was specifically chosen for reasons of convenience.\footnote{We had 2 members of our group residing in this area.}
 
@@
 This value also reveals information about the distribution of trees: clustered, random or in a regular pattern, as seen in Figure 5 below.
+These values will be analysed in the later section \ref{sec:analysis} Analysis.
 
@@
 \section{Analysis}
+\label{sec:analysis}
 
@@
 \section{Conclusion}
+\label{sec:conclusion}
 

### 2021-05-17 math/ia_2021/paper.tex
ia 2021 final....

@@
 \renewcommand{\footrulewidth}{0.4pt}
+
+\pgfplotstableread{
+    X Y
+    368 1700
+    340 1500
+    665 2800
+    954 5000
+    331 1300
+    556 2200
+    376 1300
+}\table
 
@@
         \vspace{4cm}
-        Word count: TBD \\
-        \textit{First draft, details are missing}
+        \input{wordcount.txt} \\
         \vfill
@@
 
-The dataset, procured from the National Oceanic and Atmospheric Administration website \cite{noaa}, includes many tsunami runups, going back before the year 0.
+The dataset, procured from the National Oceanic and Atmospheric Administration website \cite{noaa}, includes many tsunami runups, going back to the year 2000 BC.
 Due to this, the NOAA states uncertainties that can occur in its data.
@@
 
-$y =  \beta_0 a + \beta_1 x + e $,
+$y =  \beta_0 x + \beta_1 x + e $,
 
@@
 It is an estimate of the amount by which the predictions differ from the real value.
-An example of linear regression can be found in Appendix \ref{app:regression}, where a set of data points is presented on a scatter plot, and the regression line is found by finding the individual coefficients $\beta_n$ of $x$ (see equation above), like such:
+An example of linear regression can be found in Appendix \ref{app:regression}.
+The set of data points is presented here on a scatter plot (along with its line of regression, explained in the following paragraph):
+
+\begin{center}
+    \begin{tikzpicture}[scale=1.1]
+        \begin{axis}[
+                title={Comparing X and Y},
+                xlabel={X axis},
+                ylabel={Y axis},
+                xmin=300, xmax=1000,
+                ymin=0, ymax=5500,
+                ytick={0,1000,2000,3000,4000,5000},
+                xtick={0,400,500,600,700,800,900},
+                legend pos=north west,
+                ymajorgrids=true,
+                grid style=dashed,
+                scale=1.2
+            ]
+
+            \addplot[
+                only marks,
+                color=blue,
+                mark=*,
+            ] table {\table};
+            \legend{Datapoint}
+
+            \addplot [thin, blue] table[
+                    y={create col/linear regression={y=Y}}
+                ] % compute a linear regression from the input table
+                {\table};
+            \addlegendentry{
+                Linear regression: $ x =
+                    \pgfmathprintnumber{\pgfplotstableregressiona}
+                    \cdot t
+                    \pgfmathprintnumber[print sign]{\pgfplotstableregressionb}$
+            };
+
+        \end{axis}
+    \end{tikzpicture}
+\end{center}
+
+The regression line is found by finding the individual coefficients $\beta_n$ of $x$ (see equation above), like such:
 
@@
 So the relationship between the two independent variables and the wave height can be estimated with a multivariate, polynomial regression model.
-We will use the Python scikit-learn library to aid in creating this model.
-\\
-
-(\textit{Section not done, need to do more research on how the} \verb|PolynomialFeatures| \textit{function transforms a linear array to one fit for polynomial regression})
+The Python scikit-learn library was used to aid in creating this model.
+The model itself can be found in Appendix \ref{app:scripts}.
 
@@
 
+This function is implemented in the sklearn library, and when used on our model, we get the result of $R^2 = 0.36$ (rounded to 2 s.f.).
+
+This result is not optimal, as it means that only $\frac{1}{3}$ of the data from our model fits the test dataset.
+However, the $R^2$ value does not explain anything about causation, which is obviously present in this data, as it could be said that logically the distance from a wave is proportional to its observed height.
+
 \section{Conclusion}
 
-The research included in this paper could be useful for many life-saving scenarios, notably ones in which, for example, high quality and large-scale tsunami alert systems are not available.
-
-\textit{Conclusion and evaluation of the method are contradictory, to be corrected}
+The research included in this paper would not be very useful for life-saving scenarios, because of its low $R^2$ value.
+However, further research by professionals on this subject could potentially reveal relationships between these variables and as such create better models.
+A low score such as the one in this paper can be due to many things, but mainly I think this was due to the fit of the model.
+An exponential regression model would fit better.
 
@@
 
-It has also come to my attention that using the $R^2$ value for a non-linear relationship is not suitable, as it does not follow the same characteristics.
+It has also come to my attention that using the $R^2$ value for a non-linear relationship is not suitable, as it assumes the relationship is linear \cite{frost_2018}.
 
@@
 
-\pgfplotstableread{
-    X Y
-    368 1700
-    340 1500
-    665 2800
-    954 5000
-    331 1300
-    556 2200
-    376 1300
-}\table
-
-Below is the dataset plotted, along with a line of regression.
-
-\begin{center}
-    \begin{tikzpicture}[scale=1.1]
-        \begin{axis}[
-                title={Comparing X and Y},
-                xlabel={X axis},
-                ylabel={Y axis},
-                xmin=300, xmax=1000,
-                ymin=0, ymax=5500,
-                ytick={0,1000,2000,3000,4000,5000},
-                xtick={0,400,500,600,700,800,900},
-                legend pos=north west,
-                ymajorgrids=true,
-                grid style=dashed,
-                scale=1.2
-            ]
-
-            \addplot[
-                only marks,
-                color=blue,
-                mark=*,
-            ] table {\table};
-            \legend{Datapoint}
-
-            \addplot [thin, blue] table[
-                    y={create col/linear regression={y=Y}}
-                ] % compute a linear regression from the input table
-                {\table};
-            \addlegendentry{
-                Linear regression: $ x =
-                    \pgfmathprintnumber{\pgfplotstableregressiona}
-                    \cdot t
-                    \pgfmathprintnumber[print sign]{\pgfplotstableregressionb}$
-            };
-
-        \end{axis}
-    \end{tikzpicture}
-\end{center}
-
 \subsection{Python program}

### 2021-05-17 math/ia_2021/model/multregr2.py

@@ -81,7 +81,7 @@ if __name__ == "__main__":
     # The mean squared error
     print("Mean squared error: %.2f" % np.mean((y_pred - y_test) ** 2))
     # Explained variance score: 1 is perfect prediction
-    print('Variance score: %.2f' % regressor.score(predict_, y_test))
+    print('R^2: %.2f' % regressor.score(predict_, y_test))
 
     # Code for generating 3d plot with prediction scatter 
     # only generates one graph, rerun with PolynomialFeatures(degree=n)

### 2021-05-26 geography/ia/paper.tex
typo fix
analysis & data start (work on this later pls)

@@
         \includegraphics[width=0.7\linewidth]{media/7esites.png}
-        \caption{Map of sites chosen for the residential area, the XVI\textsuperscript{th} \textit{arrondissement}. Map base layer courtesy of Google Maps, pictures seen are taken on a mobile phone.}
+        \caption{Map of sites chosen for the touristic area, the VII\textsuperscript{th} \textit{arrondissement}. Map base layer courtesy of Google Maps, pictures seen are taken on a mobile phone.}
     \end{minipage}
@@
 
-\section{Analysis}
+\section{Data \& Analysis}
 \label{sec:analysis}
 
-\textit{Not done}
+Once our data collected, we can analyse certain patterns that appear
 
-\subsection{Environmental and cultural sustainability}
-\textit{Not done}
+\subsection{Data collected}
 
-\subsection{Economic sustainability}
-\textit{Not done}
+\subsection{Analysis}
 
@@
 
-\textit{Not done}
+\subsection{Evaluation}
+
+There are some processes in our paper that would be done differently if done now. For example:
+
+\begin{itemize}
+
+    \item The computer program does not seem to take into account an area of greenery in the VII\textsuperscript{th}, which can change the NNI value found for the VII\textsuperscript{th}.
+This area includes parks and such, however the map (generated also with the computer program) does not show trees in this location.
+This should have been investigated, however due to constraints related to time, I could not.
+    \item
+
+\end{itemize}
 

### 2021-05-26 geography/ia/nearest_neighbor_index.py

@@ -67,8 +67,11 @@ if __name__ == "__main__":
 
     parser.add_argument(
         '-b', '--bounding_box', help='choose an area bounded by 2 coordinates in comma-separated form: lat1,lon1,lat2,lon2. If no value is supplied the entire area is chosen', type=lambda s: [float(item) for item in s.split(',')])
+
     parser.add_argument(
         '-j', '--json', help='export bounded area to a json file JSON')
+
+    parser.add_argument('-v', '--verbose', dest='verbose', action='store_true', help='verbose output (show number of points etc)')
 
     args = parser.parse_args()
 
@@ -116,6 +119,8 @@ if __name__ == "__main__":
         npbounded, nni_querytree), box_area(boundingbox), len(npbounded))
 
     print("Calculated NNI value for", args.dataset, "=", nni)
+    if args.verbose is True: 
+        print("n: ", len(npbounded))
 
 
     #

### 2021-05-27 geography/ia/paper.tex
geo ia work -> toward final draft

@@
 
-Our results for the bipolar survey are shown in Appendix \ref{app:bipolar}
+\subsubsection{Statistical evaluations}
 
-When tallied, the total scores for each site can be expressed as a fraction over 84, as shown in Figures 4, 5.
+In order to either support or disprove our third hypothesis, it is important to study how trees are scattered in Paris, which reveals how dispersed green spaces are from one another.
+A statistical test called the Nearest Neighbor Index (NNI) test can provide a reasonable answer to this: if the index is similar for both sites, it can easily be said then that our hypothesis is supported.
+This value also reveals information about the distribution of trees: clustered, random or in a regular pattern, as seen in Figure 4 below.
+These values will be presented and analysed in the later section \ref{sec:analysis}.
+Data \& Analysis.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.2\linewidth]{media/nni1.png}
+    \includegraphics[width=0.2\linewidth]{media/nni2.png}
+    \includegraphics[width=0.2\linewidth]{media/nni3.png}
+    \caption{The NNI measures the spacial distribution of data. From a value of 0, representing a clustered pattern, to 1 (random) and to 2.15 (uniform pattern)}
+\end{figure}
+
+As individually counting trees in these areas would be a tedious task, we used the Paris OpenData platform (licensed under the permissive ODbL license\footcite{odbl}), owned by the government.
+The fact that this kind of database platform exists, shows a certain level of transparency on the government's part.
+
+We used the trees dataset procured from the library, which contains the exact coordinates of trees in the city.
+\footcite{paris_opendata} There are also other information such as species of the tree and such, however these are cleaned from the dataset for our purposes.
+
+A script, written in Python was tasked to calculate the NNI value, because with a dataset of over 20,000 data points this would not be possible by hand.
+We go through each point, and find its nearest neighbor.
+In order to calculate distances between these two points, we use the Haversine Formula, as the two points would be in the form of coordinates.
+With this information the NNI can be calculated with the simple formula
+
+$$Rn = \frac{2 \bar D}{\sqrt{\frac{a}{n}}}$$
+
+where $Rn$ is the NNI index value, $\bar D$ is the mean observed distance to the nearest neighbor, $a$ is the area of the zone and $n$ is the total number of data points.
+
+\subsubsection{Questionnaire}
+
+We also collected data from a questionnaire sent out to many people.
+This questionnaire contained the two questions \textit{"Do you believe that the Passy area has more green spaces than the area around the Eiffel tower?
+(Not including the champ de mars)"} and \textit{"How much litter is in the Passy area compared to around the Eiffel tower?"}.
+The data collected from this survey is listed in Section \ref{sec:data}.
+We surveyed our year, which has people that live in and frequent these two areas.
+
+% NEW SECTION
+
+\section{Data \& Analysis}
+\label{sec:analysis}
+
+Once our data collected, we can analyse certain patterns or appearances that do not follow a general trend. Our data is shown below in Subsection \ref{sec:data}.
+
+\subsection{Data collected}
+\label{sec:data}
+
+Our results for the bipolar survey are shown in Appendix \ref{app:bipolar} (not shown here due to large size).
+When tallied, the total scores for each site can be expressed as a fraction over 84, as shown in Figures 5, 6.
 
@@
 
-When plotted on a map, these totals can be shown with a gradient of colors.
+When plotted on a map, these totals can be shown as percentages ($\frac{s}{84} \cdot 100$, where $s$ is the total score) with a gradient of colors, as seen below in Figures 7, 8:
 
@@
     \includegraphics[width=0.7\linewidth]{media/7eColors.png}
-    \caption{Each site in the VVI\textsuperscript{th}, with its respective total score expressed as a percentage for clarity, and a color assigned based on the gradient in Figure 8.}
+    \caption{Each site in the VVI\textsuperscript{th}, with its respective total score expressed as a percentage for clarity, and a color assigned based on the gradient in Figure 9.}
 \end{figure}
@@
     \includegraphics[width=0.7\linewidth]{media/16eColors.png}
-    \caption{Each site in the XVI\textsuperscript{th}, with a percentage. Color assigned with gradient in Figure 8.}
+    \caption{Each site in the XVI\textsuperscript{th}, with a percentage. Color assigned with gradient in Figure 9.}
 \end{figure}
@@
 
-\subsubsection{Statistical evaluations}
+As for images collected during our inspections of the sites, for the sake of keeping the file size of this paper down, I will not attach the images collected and they will be available for download or viewing at the URL \url{https://github.com/kinnounko/notes/tree/main/geography/ia}.
 
-In order to either support or disprove our third hypothesis, it is important to study how trees are scattered in Paris, which reveals how dispersed green spaces are from one another.
-A statistical test called the Nearest Neighbor Index (NNI) test can provide a reasonable answer to this: if the index is similar for both sites, it can easily be said then that our hypothesis is supported.
-This value also reveals information about the distribution of trees: clustered, random or in a regular pattern, as seen in Figure 5 below.
-These values will be analysed in the later section \ref{sec:analysis} Analysis.
+The dataset collected from the questionnaire is shown in the tables in Figures 10, 11.
 
 \begin{figure}[H]
-    \centering
-    \includegraphics[width=0.2\linewidth]{media/nni1.png}
-    \includegraphics[width=0.2\linewidth]{media/nni2.png}
-    \includegraphics[width=0.2\linewidth]{media/nni3.png}
-    \caption{The NNI measures the spacial distribution of data. From a value of 0, representing a clustered pattern, to 1 (random) and to 2.15 (uniform pattern)}
+    \begin{center}
+        \begin{tabular}{||c c||}
+            \hline
+            Answer             & Number of people that agree \\ [0.5ex]
+            \hline\hline
+            Significantly Less & 4                           \\
+            \hline
+            Less               & 8                           \\
+            \hline
+            About the same     & 8                           \\
+            \hline
+            More               & 10                          \\
+            \hline
+            Significantly more & 0                           \\ [1ex]
+            \hline
+        \end{tabular}
+    \end{center}
+    \caption{\textit{Do you believe that the Passy area has more green spaces than the area around the Eiffel tower? (Not including the champ de mars)}}
 \end{figure}
 
-As individually counting trees in these areas would be a tedious task, we used the Paris OpenData platform (licensed under the permissive ODbL license\footcite{odbl}), owned by the government.
-The fact that this kind of database platform exists, shows a certain level of transparency on the government's part.
-
-We used the trees dataset procured from the library, which contains the exact coordinates of trees in the city.
-\footcite{paris_opendata} There are also other information such as species of the tree and such, however these are cleaned from the dataset for our purposes.
-
-A script, written in Python was tasked to calculate the NNI value, because with a dataset of over 20,000 data points this would not be possible by hand.
-We go through each point, and find its nearest neighbor.
-In order to calculate distances between these two points, we use the Haversine Formula, as the two points would be in the form of coordinates.
-With this information the NNI can be calculated with the simple formula
-
-$$Rn = \frac{2 \bar D}{\sqrt{\frac{a}{n}}}$$
-
-where $Rn$ is the NNI index value, $\bar D$ is the mean observed distance to the nearest neighbor, $a$ is the area of the zone and $n$ is the total number of data points.
-
-% NEW SECTION
-
-\section{Data \& Analysis}
-\label{sec:analysis}
-
-Once our data collected, we can analyse certain patterns that appear
-
-\subsection{Data collected}
+\begin{figure}[H]
+    \begin{center}
+        \begin{tabular}{||c c||}
+            \hline
+            Answer             & Number of people that agree \\ [0.5ex]
+            \hline\hline
+            Significantly Less & 3                           \\
+            \hline
+            Less               & 10                           \\
+            \hline
+            About the same     & 12                           \\
+            \hline
+            More               & 5                          \\
+            \hline
+            Significantly more & 0                           \\ [1ex]
+            \hline
+        \end{tabular}
+    \end{center}
+    \caption{\textit{How much litter is in the Passy area compared to around the Eiffel tower?}}
+\end{figure}
 

### 2021-05-27 geography/ia/paper.tex
added graphs for questionnaire -> toward final draft

@@
 \subsubsection{Questionnaire}
+\label{sec:questionnaire}
 
@@
 
-Once our data collected, we can analyse certain patterns or appearances that do not follow a general trend. Our data is shown below in Subsection \ref{sec:data}.
+Once our data collected, we can analyse certain patterns or appearances that do not follow a general trend. Our data is shown below in Section \ref{sec:data}.
 
@@
 
-The dataset collected from the questionnaire is shown in the tables in Figures 10, 11.
+The dataset collected from the questionnaire (mentioned in Section \ref{sec:questionnaire}) is shown in the tables in Figures 10, 11.
 
@@
 
+This data can be shown in the form of a bar graph (Figure 12), in order to visualize the distribution and for further analysis.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=1\linewidth]{media/graphs.png}
+    \caption{Graphs representing aforementioned data in Figures 10, 11}
+\end{figure}
+
 \subsection{Analysis}
@@
 
-There are some processes in our paper that would be done differently if done now. For example:
+There are some processes in our paper that would be done differently if done now. To list the ones I managed to find at the time of writing:
 

### 2021-05-27 geography/ia/paper.tex
geo ia final draft

@@
 A script, written in Python was tasked to calculate the NNI value, because with a dataset of over 20,000 data points this would not be possible by hand.
-We go through each point, and find its nearest neighbor.
-In order to calculate distances between these two points, we use the Haversine Formula, as the two points would be in the form of coordinates.
+We go through each point, and find its nearest neighbor, using an efficient complex data structure called a $k$d-tree.\footnote{$k$-dimensional tree.
+In this case, this "tree" is 2-dimensional.
+More information at: \cite{bentley_kdtree}}  In order to calculate distances between these two points, we use the Haversine Formula, as the two points would be in the form of coordinates.
 With this information the NNI can be calculated with the simple formula
@@
 
+\subsubsection{Bipolar survey}
+
+The results from our bipolar semantic survey seem to suggest a few things.
+Firstly, for the VII\textsuperscript{th}, the mean of the data series seems to be an overall environmental percentage value of $62.5\%$.
+We can notice an outlier in this data: the Seine river bank, with a score of $82\%$.
+This higher score is due to the nature of this site: no cars, low vandalism and no odor (due to its proximity to a body of water).
+In general however, it seems that the more tourist-heavy areas are more prone to lower scores, such as the foot of the Eiffel Tower and the Iena Bridge.
+
+In the XVI\textsuperscript{th}, the results differ from the ones in the touristic area.
+Again, the mean of the data series is $58\%$, considerably lower than that of the touristic area.
+This does include the obvious outlier however, the Ranelagh garden at $92\%$.
+Considering the sites with higher car counts and noise levels (such as Avenue Mozart) are in this residential area, this does indeed agree with our 2\textsuperscript{nd} hypothesis.
+
+In general, it can be said with this information that statistically, our bipolar semantic survey agrees with the 1\textsuperscript{st} and 2\textsuperscript{nd} hypotheses made at the beginning of our paper: the residential area does indeed show higher levels of noise and a lower environmental score.
+
+\subsubsection{Nearest Neighbor Index}
+
+We must analyse the NNI values from the Nearest Neighbor statistical test in order to either prove or disprove the 3\textsuperscript{rd} hypothesis we made.
+With the \verb|arbres| dataset downloaded, we can use the \verb|nearest_neighbor_index.py| \footnote{The program can be found at \url{https://github.com/kinnounko/notes/blob/main/geography/ia/nearest_neighbor_index.py} or Appendix \ref{app:program}} program in order to calculate the NNI values of the XVI\textsuperscript{th} and the VII\textsuperscript{th}.
+We must use the following syntax:
+
+\begin{lstlisting}[breaklines]
+    python nearest_neighbor_index.py -b x1,y1,x2,y2 /folder/dataset.csv
+\end{lstlisting}
+
+Or in practice, for the XVI\textsuperscript{th}:
+\begin{lstlisting}[breaklines]
+    python nearest_neighbor_index.py -b 48.85373,2.26728,48.86042,2.27865 /home/octo/trees.csv
+\end{lstlisting}
+
+And for the VII\textsuperscript{th}:
+\begin{lstlisting}[breaklines]
+    python nearest_neighbor_index.py -b 48.84881,2.28708,48.86075,2.30742 /home/octo/trees.csv
+\end{lstlisting}
+
+This will output the NNI values for the XVI\textsuperscript{th} and the VII\textsuperscript{th}, respectively. They are shown in the table below in Figure 13.
+
+\begin{figure}[H]
+    \begin{center}
+        \begin{tabular}{||c c||}
+            \hline
+            Area                    & NNI value (no unit, rounded to 3 s.f.) \\ [0.5ex]
+            \hline\hline
+            XVI\textsuperscript{th} & 0.610                       \\
+            \hline
+            VII\textsuperscript{th} & 0.534                       \\ [1ex]
+            \hline
+        \end{tabular}
+    \end{center}
+    \caption{Computed NNI values}
+\end{figure}
+
+This information allows us to reach a conclusion about the similarity of the dispersion of trees in the two areas: both values are between the "Clustered" value of 0.00 and the "Random" value of 1.00, and as such are both dispersed spatially in a clustered-random manner.
+We are however not interested in their manner of dispersion but the similarity between the two values.
+And as it can be seen, these values are very similar.
+
+In fact, if we observe the graph below in Figure 14, we can see that our NNI values express very "significant element of clustering", as our datasets have a number of points of 1277 and 5231 (values that are not included in the $x$-axis of this graph due to their large size):
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.7\linewidth]{media/certaintynn.png}
+    \caption{A graph showing the certainty of a NNI value result}
+\end{figure}
+
+And as such, to conclude, the 3\textsuperscript{rd} hypothesis of our research is confirmed (at least for the two study areas).
+We can with confidence say that trees, and as a result green spaces, are evenly spread in the two areas.
+They also exhibit signs of clustering, which signifies the presence of these trees in groups, or parks.
+As the mayor has stated, it does seem that green spaces are available to everyone.
+
+\subsubsection{Questionnaire}
+
+As seen in Figure 12, our questionnaire brought some mixed opinions.
+Many believed that the XVI\textsuperscript{th} has more green spaces surrounding it, as shown by the negatively skewed bar chart.
+People also believed that the residential area in the VII\textsuperscript{th} was less littered than the touristic one, as shown by again a negatively skewed graph.
+
+This information does go against our hypotheses, but as this is an opinionated survey and because I believe some people have not visited one of the sites, I will not allocate much weight to these results and rather rely on empirical evidence.
+
 \section{Conclusion}
@@
 
+With this information analysed, we can reach a sensible conclusion.
+Paris as a city is generally not very favorable environmentally: with high air and noise pollution in residential areas and other issues, the most habited places of the city are not very nice to be in.
+However touristic areas do show a small but noticeably higher level of cleanliness.
+
+There is however a visible effort in order to fix some of these more unfavorable areas.
+It could be for press-related issues, with the upcoming 2024 Paris Olympics, or by pressure from international organizations with their pollution guidelines, but there is a noticeable effort being done: promises of parks for everyone by the mayor seem to have been fulfilled and are valid, and more regulations are being put into place as evidenced by the 88-page manifesto-like government sponsored brochure.
+\footcite{bruitparif} These improvements are not uniform in the city, and personally they seem to be focused in touristic and publicized areas.
+Throughout my time in Paris, I have seen many improvements to environmental quality, albeit predominately in these touristic areas, and zero to no effort in residential areas such as the one I live in.
+In order to promote environmental sustainability in this city, there must be some action taken in residential areas, and not only in tourist-frequented ones.
+
 \subsection{Evaluation}
 
-There are some processes in our paper that would be done differently if done now. To list the ones I managed to find at the time of writing:
+There are some processes in our paper that would be done differently if done now. To list the ones I have managed to find at the time of writing:
 
@@
 This should have been investigated, however due to constraints related to time, I could not.
-    \item
+I suspect this was due to a preliminary cleanup of the data.
+I do not think this error has affected the outcome of our research by a lot, as with these parks the value would be around the same, as they also show signs of clustering.
+
+          \begin{figure}[H]
+              \centering
+              \includegraphics[width=0.5\linewidth]{media/boundingerror.png}
+              \caption{A map showing the trees in the dataset used for the VII\textsuperscript{th}, along with some parks in the bottom left quadrant, missing their trees.}
+          \end{figure}
+
+    \item Our research was done during a time of lock-down due to the SARS-COVID-19 pandemic in France.
+Due to travel restrictions, international --- even national tourist counts were considerably down.
+This causes many discrepancies in the data as, well, we are in fact studying a touristic area.
 
 \end{itemize}
+
+To conclude the evaluation, it can be said that despite some disparity in our data, our experimental method was correct, computational issues aside.
+Once the SARS-COVID-19 pandemic over, this research could be repeated once again for more accurate results.
 
@@
 All media and information related to this IA can be found on the URL \url{https://github.com/kinnounko/notes/tree/main/geography/ia}.
-This includes images, the \LaTeX{} source code to this paper, and the Python program used to calculate the NNI value.
+This includes images, the \LaTeX{} source code to this paper, and the Python program used to calculate the NNI value (also found in Appendix \ref{app:program}).
 
@@
 
+\subsection{NNI computation program}
+\label{app:program}
+\lstinputlisting[breaklines]{nearest_neighbor_index.py}
+
 \end{document}
